### 2019-05-13 paper/paper.tex
Exported to TeX; can't be emacs+latex =[

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage[margin=1in]{geometry}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Solution of Differential Algebraic Equations with Manifold Constraints on the Latent Space of an Autoencoder},
            pdfauthor={Alejandro Francisco Queiruga},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Solution of Differential Algebraic Equations with Manifold Constraints
on the Latent Space of an Autoencoder}
\author{Alejandro Francisco Queiruga}
\date{2018-9-20}

\begin{document}
\maketitle
\begin{abstract}
We train an autoencoder to determine the lower-dimensional
representation of the EOS surface using modern deep learning techniques
without using any phase labels. The decoder end of the model is then
plugged back into the balance laws, eliminating the constrain equation.
\end{abstract}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{header-n3216}{%
\section{Introduction}\label{header-n3216}}

Representing complex physical systems with multiple speicies, phases,
chemical reactions, and heterogenous scale effects poses an
intractability problem to modeling. How do we analyze complex
experimental data and \emph{ab initio} simulations and derive new
macroscale relations to describe the systems? How do we then implement
these relations in the rest of our analyses and production-scale
computer simulations? A single material with multiple phase transitions
is a challenge in of itself, and the descriptions become combinatorially
more complex as species and phases are added. This results in
combinatorially more human labor and combinatorially more code which
tends to involve complex branching logic. The descriptions then pose
numerical challenges when incorporated into macroscale simulations due
to obstacles such as discrete phase state machines and primary variable
switching.

Motivations for this work are subsurface energy systems such as shale
and oceanic natural gas reservoirs in shales or engineered geothermal
systems.

This work is only considered with the "single beaker" problem; the flow
and transport aspects have similar challenges. (The next stage of this
work will "connect the beakers".)

This work proposes a new methodology that steps away from the
traditional approaches

In this approach, we directly \emph{look for solvable representations}
instead of representations that carry human intuition. We value
simulation runtime and robustness. Further, we must remember the value
of the human effort involved: beyond improving computational aspects, we
are fully automating part of the human involvement in programming the
simulator.

Part of the new modern software engineering framework involves replacing
human-written code with computer-written code. We are applying this
mindset from the computer-vision and advertisement dominated Artificial
Intelligence space into scientific computing.

The actual meaning of the parameterization is not important, so long as
it can be effectively solved and then \emph{decoded} back to the
physically meaningful quantities.

The original redescribe our problem as a differential algebraic equation
with a constraint: the balances of mass and energy are constrained by
the EOS. Because the EOS was originally derived from data, a data-driven
approach lends itself in which the EOS is a set of points in
temperature-pressure-density-enthalpy space. (In the multicomponent
cases there are more dimensions.)

We demonstrate this approach to solving constrained differential
equations on the pendulum, before demonstrating two EOSes of water: one
with just liquid-gas phases, and one with liquid-gas-ice-supercritical
regimes. The notion of phases does not appear in this multiphase
simulation. We show a model architecture that is able to learn phase
labels in the unsupervised autoencoder training process, such
human-scientist interpretable results are obtained from the machine
learning process. We compare this and other model architectures,
including common deep neural networks, based on autoencoder error and
simulation success metrics. We stress that the presented approach is
only utilizing machine learning on material relations derived from
empirical data: the simulation is indeed solving physical balance laws.

Our proposed approach is to build upon modern machine learning
techniques to aid in equation discovery. An autoencoder is trained on
the original material dataset to learn a lower-dimensional
representation of the underlying manifold. The model is designed to be
directly incorporated into a macroscale simulation of physical balance
laws, where now the traditional partial differential equations are being
solved on the learned latent space. This provides a more robust
numerical algorithm, while putting the onus of programming part of the
simulation itself onto the computer: tens of thousands of lines of human
written code consisting of empirical fits and branching logic are
replaced by code generated by the learned model. The methodology will
first be demonstrated on datasets based on thermodynamic intensive
variables, which are already challenging to analyze, before being
extended to richer datasets such as experimental images or molecular
dynamics. We stress that the proposed approach is only utilizing machine
learning on material data: the simulation is indeed solving physical
balance laws. Through deep learning approaches combined with
domain-specific goals, we can learn feature sets that are both
interpretable and directly apply to analyzing larger systems.

Consider a general differential algebraic equation of the form

\begin{equation}
z(\dot{u},u) = 0
\end{equation}

We are interested in equations that have the form

\begin{align}
\frac{\mathrm{d}u}{\mathrm{d}t} & = v \\
0 & = c(u,v)
\end{align}

Where do such equations arise? A simple example is the pendulum, which
will be explored in Section 2.

\begin{align}
\text{Solve for}\, x(t), \, y(t), \, f(t) \, \text{satisfying:} \\
m \dot{x} & = f x/L \\
m \dot{y} & = f y/L + m g \\
x^2 + y^2 & = L^2
\end{align}

\hypertarget{header-n3234}{%
\section{Multiphase Equations of State}\label{header-n3234}}

We also consider the phase equilibria; at a larger scale it is possible
for a control volume to be a mixture of two phase coexisting, such as
boiling with some mixture saturation \(S_{gas}=1-S_{liq}\).

For this preliminary work, we consider only a single-beaker problem
without flow. Let \(\rho\) be the material density, \(T\) be the
temperature, \(p\) be the material pressure, \(h\) be the material
enthalpy, and \(u\) be the internal energy. There are two equations for
balance of mass and energy.

\begin{align}
\partial_t \rho & = \mathbf{k}_p(p_\infty - p) + r\\
\partial_t \rho u & = \mathbf{k}_T(T_\infty-T) + s
\end{align}

The beaker is exposed to a reservoir with \(p_\infty\) and \(T_\infty\)
via a mixed boundary condition straw with transport and conduction
coefficients \(\mathbf{k}_p\), \(\mathbf{k}_T\). Mass and heating
sources and sinks are the terms \(r\) and \(s\).\footnote{We express the
  material by enthalpy instead of internal energy; they are related by
  the relation \(\rho u = \rho h - p\). Future work may work directly
  with internal energy.}

The typical methodology is to use the empirical relations for density
and enthalpy \(h\) as a function of pressure and temperature and solve
the DAE for \(p\) and \(T\) implicitly:

\begin{align}
\text{Solve for}\, p(t)\, \text{and}\, T(t)\, \text{such that:}\\
\partial_t \rho(p,T) & = \nabla \cdot \mathbf{k}\nabla p + r\\
\partial_t \rho u(p,T) & = \nabla \cdot \mathbf{k'}\nabla T + s
\end{align}

The complicatation is that the functions \(\rho(p,T)\) and \(u(p,T)\)
are not well defined functions due to the presence of phase changes that
yield sharp discontinuities in \(p,T\), as shown in Figure 1 for water.
The material can exist on this section of the surface as a mixture of
the two phases.

\[\rho = X_{H_2O}\left(S_{gas} \rho_{gas} + S_{liq} \rho_{liq} + S_{ice} \rho_{ice} + S_{crit} \rho_{crit} \right)\]

We will also note right now that applicaitons also involve continuum
mechanics and chemical reactions! The interplay of deformations,
reactions, and phase transitions makes the equations extremely difficult
to solve.

One possible definition of phase boundaries:

\begin{quote}
Phase transitions occur when the thermodynamic free energy of a system
is non-analytic for some choice of thermodynamic variables (cf. phases).

https://en.wikipedia.org/wiki/Phase\_transition
\end{quote}

\begin{figure}
\centering
\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/phase_diagram.png}
\caption{}
\end{figure}

\begin{figure}
\centering
\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/water_eos.png}
\caption{}
\end{figure}

\hypertarget{header-n3253}{%
\subsection{State Machine logic}\label{header-n3253}}

Formulate balance equations as

\[R(X ; phase) = 0\]

Solve the differentiable part:
\[\frac{\partial R}{\partial X}\Delta X^{k+1/2} = -R(X^k ; phase^k)\]
and iterate on the nondifferential part:
\[phase^{k+1},X^{k+1} = statemachine(X^{k+1/2} ; phase^k)\]
The phase switching logic has the form of a switch case:

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{switch}\NormalTok{ phase_old:}
  \ControlFlowTok{case}\NormalTok{ gas:}
    \ControlFlowTok{if}\NormalTok{ p,T crossing boundary:}
\NormalTok{      phase_new = liquid_gas}
  \ControlFlowTok{case}\NormalTok{ liquid:}
    \ControlFlowTok{if}\NormalTok{ p,T crossing boundary:}
\NormalTok{      phase_new = liquid_gas}
  \ControlFlowTok{case}\NormalTok{ liquid_gas:}
    \ControlFlowTok{if}\NormalTok{ S_gas >= }\DecValTok{1}\NormalTok{:}
\NormalTok{      phase_new = gas}
    \ControlFlowTok{if}\NormalTok{ S_liquid >= }\DecValTok{1}\NormalTok{:}
\NormalTok{      phase_new = gas}
\end{Highlighting}
\end{Shaded}

\hypertarget{header-n3262}{%
\subsection{A constrained Differential Algebraic
Equation}\label{header-n3262}}

We argue that the relations that are used are not the \emph{ground
truth} in of themselves. Beyond ideal gases and simple linear fluids,
the functions \(\rho(p,T)\) or \(p(\rho,T)\) are simply complicated fits
that were derived in an ad-hoc piecewise fashion from experimental data.
This is not a bad thing---it is the fundamental nature of the problem.
When we look at the corpus of literature on a material as water, what we
actually have is a decision branch to multiple different complicated
fits for each material branch.

We instead replace our problem with two balance laws and one unknown
constraint:

\begin{align}
\text{Solve for}\, \rho(t), \, h(t), \, p(t),\, \text{and}\, T(t)\, \text{satisfying:}\\
\partial_t \rho & = \nabla \cdot \mathbf{k}\nabla p + r\\
\partial_t \rho h - p & = \nabla \cdot \mathbf{k'}\nabla T + s\\
\text{such that they lie on the material EOS,}\\
EOS(\rho,h,p,T) & = 0
\end{align}

The \emph{ground truth} is the experimental data in the first place; the
branching curve fit is only one realization of representing the data. We
have demonstrated that it is not possible to represent any two of these
in terms of the other two; and even if so, it is not necessarily a good
and robust fit. What if we search for a new fit?

\hypertarget{header-n3267}{%
\section{The Pendulum}\label{header-n3267}}

An archetypical constraint problem the pendulum. Posed as dynamics, it
has three unknowns wtih two second-order equations and a constraint,

\begin{align}
\text{Solve for}\, x(t), \, y(t), \, f(t) \, \text{satisfying:} \\
m \ddot{x} & = f x/L \\
m \ddot{y} & = f y/L + m g \\
x^2 + y^2 & = L^2
\end{align}

In our methodology, we do not have an explicit equation for the
constraint, but a dataset of (x,y) pairs that are on the manifold. For
our proof-of-concept test, we manufacture this data for the pendulum.
Note that we are considering the nonlinear pendulum.

There are many ways to solve this cannonical equation: Lagrange
multipliers (\(f(t)\) is the multiplier in the way we have written it
now), penalties, change of variables, etc. The change of variables
method is to introduce a new variable \(\theta\) that parameterizes
\(x=L\cos\theta\) and \(y=L\sin\theta\):

\begin{align}
\text{Solve for}\, \theta(t) \text{and} \, f(t) \, \text{satisfying:} \\
m \ddot{x}(\theta) & = f x(\theta)/L \\
m \ddot{y}(\theta) & = f y(\theta)/L + m g \\
\end{align}

In this example, this lets us collapse the problem into a
single-component Ordinary Differential Equation.

The manifold for this simple problem is smooth, so two nested
polynomials forms the autoencoder:

\[\left\{\begin{array}{c} x\\y\end{array}\right\}
\rightarrow W^1 x^n \rightarrow q \rightarrow W^2 x^n \rightarrow 
\left\{\begin{array}{c} x\\y\end{array}\right\}\]

(We explicitly write the bias \(+b\); it is equivalent to including
\(x^0=1\) in the polynomial set.) The Lagrangian for the pendulum is

\begin{equation}
L = \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right) - m g y
\end{equation}

with the constraint that \(x\) and \(y\) lie on the path. The equations
of motion given some parameter space that satisfies the constraint is

\begin{equation}
\frac{\mathrm{d}q}{\mathrm{d}t}\frac{\partial L}{\partial \dot{q}} = \frac{\partial L}{\partial q}.
\end{equation}

We'll use the autoencoder to generate the parameterization of \(x(q)\)
and \(y(q)\). By the chain rule, the velocities are related to the
feature space by

\begin{equation}
\dot{x} = \frac{\mathrm{d}x(q)}{\mathrm{d}t} = \frac{\partial x}{\partial q}\frac{\mathrm{d}q}{\mathrm{d}t}
\end{equation}

and similarly for the \(y\) direction.

Taking the partial derivatives with respect to the feature variable and
its rate, the components to the equation of motion are

\begin{equation}
\frac{\partial L}{\partial q} = m g \frac{\partial y}{\partial q}
\end{equation}

When we want to solve the dynamics, we can treat the autoencoder the
same as \(x(\theta)\), plugging in

\[L(x(q),v(q,\dot{q}))\]

and build the DAE we know from physics:

\[\frac{\mathrm{d}}{\mathrm{d}t}\frac{\partial L}{\partial \dot{q}} = \frac{\partial L}{\partial q}\]

where we get the velocity with the chain rule:

\[v = \frac{\partial x}{\partial q} \dot{q}\]

Then plug it into a Runge Kutta (trapezoidal) to solve with Newton's
method:

\[\frac{\partial L}{\partial \dot{q}}_i - \Delta t \frac{\partial L}{\partial q}_i = \frac{\partial L}{\partial \dot{q}}_0 + \Delta t \frac{\partial L}{\partial q}_0\]

\begin{figure}
\centering
\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/pendulum_q.png}
\caption{}
\end{figure}

\hypertarget{header-n3294}{%
\section{Representation Learning for
Reparameterization}\label{header-n3294}}

The central idea of this work is to obtain two new variables solve for
those directly,

\begin{align}
\text{Solve for}\, u(t)\, \text{and}\, v(t)\, \text{such that:}\\
\partial_t \rho(u,v) & = \nabla \cdot \mathbf{k} \nabla p(u,v) + r \\
\partial_t \rho e(u,v) & = \nabla \cdot \mathbf{k'}\nabla T(u,v) + s
\end{align}

where \(u\) and \(v\) have no physical meaning other than simply being a
parameterization of the equation of state constraint. We want to select
\(u\) and \(v\) to require no auxilary phase index to define the system
and no additional logic in the code. The sharp kinks and possible
discontinuities can be a part of the functions
\(\rho(u,v), p(u,v), e(u,v)\) and \(T(u,v)\).

Can we automatically discover this parameterization? (For a such a
simple equation, we do not think the proposed methodology will be
better, but the pendulum will instead be our \emph{unit test}.) Our
proposed method is motivated by the problems of multiphase reaction and
transport, where the underlying constraints are the ill-behaved material
properties which are difficult to describe. We change our perspective on
these problems in the next section.

\hypertarget{header-n3299}{%
\subsection{Latent Space}\label{header-n3299}}

Instead of the painstaking work of parameterizing phase boundaries and
determining curve fits by hand, we rephrase the problem of representing
the equation of state by learning and autoencoder. We have an encoding
phase, \(E(\rho,p,e,T; a)\) and a decoding phase, \(D(u,v; b)\) that
forms an identity function with a compressed subspace:

\begin{equation}
\left\{ \rho, p, e, T \right\} \rightarrow  E \rightarrow \left\{ u,v \right\} \rightarrow D \rightarrow \left\{ \rho, p, e, T \right\}
\end{equation}

The autoencoder is solved for by optimizing its parameters using the
goal

\begin{equation}
\min_a \sum_x \left( x - D(E(x;a);a) \right)^2
\end{equation}

Note that we are not yet considering constraints \(c\) with partial
differential equation components in space. These types of material
equations of states are only enforced pointwise. The time components of
the equations can contain spatial derivatives; i.e. this method fits
easily inside a finite volume simulation with little change.

\hypertarget{header-n3305}{%
\subsection{Training}\label{header-n3305}}

We use a Euclidean norm loss function with a contractive penalty:

\[L(x)=\left\|x-D(E(x))\right\|^2_2+\lambda\left\|\frac{\partial E}{\partial x}\right\|_F^2\]

The gradient on the encoder is a \(4\times2\) matrix in our case, and
\(\|\|_F\) denotes the Frobenius norm. The contractive penalty
\(\lambda\) smooths out and evenly distributes the mapping to \(q\).

The variational autoencoder is a little more complicated to implement,
and we do not necessarily want to match a Gaussian interior. Because we
generated our datasets artificially (both in this manufactured setting
and in an actual laboratory setting), the dataset does not reflect draws
from a distribution, so the KL divergence against a gaussian is not
quite applicable.

We note that penalizing the decoder would smooth out the sharp kinks at
phase boundaries.

We train in two phases. The first step is minibatched stochastic
gradient descent on the entire deep network. After a few epochs of this,
we do a step of linear regression on the last layer of the network,
denote by \(W^{dec}\), which directly controls the values. The linear
system that needs to be solved is easily generated by assembling the
Hessian of the total loss function,
\begin{align}
 \mathbf{G} &= \sum_{x\in X}\frac{\partial L}{\partial W^{dec}} \\
\mathbf{H} &= \sum_{x\in X}\frac{\partial^2 L}{\partial (W^{dec})^2}
\end{align}
and solving the linear system for an update to the last parameters.

\[[\mathbf{H}] \Delta W^{dec} = -\{\mathbf{G}\}\]

To form the hessian and gradient for this step, we sum over all of the
training data. (The bias \(b^{dec}\) is included in \(W^{dec}\).) (The
contractive penalty term drops out because it does not depend on
\(W^{dec}\).)

In future work, we would want to continue appending new \(q\)s in a
systemic way to derive, e.g., an \(H_2O + NaCl\) equation of state with
3 degrees of freedom, bootstraping from the pre-trained \(H_2O\)
architecture with 2 degrees of freedom.

\hypertarget{header-n3312}{%
\subsection{Encoder Initializations}\label{header-n3312}}

The properties \(p,T\) are the usual prefered choice for primary
variables, but as discussed above, the equations of state are not
analytic for any choice. (Indeed, this idea first spawned when the
author was thinking about systematically rewriting equations in
\(\rho,h\).) Ideally, this system would be fully automated, but due to
the nonlinearilty of the autoencoder problem there are many "bad"
options the trainer can fall into. Multiple starts are possibly needed
to get a good solution when starting from purely randomized newtork
parameters. Global optimization is truly a necessity when trying to
optimize networks with so few parameters. We incorporate a little bit of
human intuition and preference by trying three options to start off the
autoencoder: 1) fully random, 2) \(p,T\) prefered, 3) \(\rho,h\)
prefered. This is implemented by added a bias to the initialization of
the encoder coefficients, e.g. for \(p,T\),
\begin{equation}
 W^{init}_{enc} = \left[\begin{array}{ccccc}
1 & 0 & 0 & 0 & ... \\
0 & 1 & 0 & 0 & ...
                        \end{array}\right]+[\sigma]
\end{equation}
where \([\sigma]\) is the random initialization and \(…\) denotes
padding the coefficients to the nonlinear terms by 0s. Recall that the
values for \(s\) have already been rescaled to be around 0 with a range
of 1. The random initialization part is decided by
\(\sigma \sim \mathcal{N}(0,0.1)\) .

\hypertarget{header-n3317}{%
\subsection{Unsupervised Phase Classification}\label{header-n3317}}

\begin{figure}
\centering
\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_network.png}
\caption{The architecture of the classifier network designed for
  unsupervised labeling of phases.}
\end{figure}

\hypertarget{header-n3321}{%
\section{Simulation}\label{header-n3321}}

Solve for \(q_1(t)\) and \(q_2(t)\) such that:

\begin{align}
\partial_t \rho(q_1,q_2) & = \nabla \cdot \mathbf{k} \nabla p(q_1,q_2) + r \\
\partial_t \rho h(q_1,q_2)-p & = \nabla \cdot \mathbf{k'}\nabla T(q_1,q_2) + s
\end{align}

where \(\rho(q_1,q_2)\) etc. are the back of an autoencoder:

\begin{equation}
\left\{ \begin{array}{c}
T\\ p\\ \rho\\ h
\end{array}\right\} \rightarrow  E \rightarrow 
\left\{ \begin{array}{c} q_1\\q_2 \end{array} \right\}\rightarrow D \rightarrow 
\left\{ \begin{array}{c}
T\\ p\\ \rho\\ h
\end{array}\right\}
\end{equation}

Using the autoencoder to automatically remove the local constraint:

\begin{equation}
eos(D(q_1,q_2)) = 0 \, \forall \,q_1,q_2
\end{equation}

\textbf{Only training the material representation, not the balance
laws.}

First, we had a constrained DAE:

\begin{align}
\frac{\mathrm{d}}{\mathrm{d}t} m(T,p,\rho,h) &= r(T,p,\rho,h)\\
eos(T,p,\rho,h) &= 0
\end{align}

with no good representation of \(eos\) everywhere.

The autoencoder parameterizes the constraint :

\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t} m(D(q)) = r(D(q))
\end{equation}

We can parameterize this equation with any time stepping scheme, .e.g. a
Runge-Kutta scheme. For subsurface flow, backward Euler is popular due
to its simplicity and L-stability, which we follow since the topic of
this paper is not exploring new time stepping schemes. This yields the
discretizated algebraic equation for \(q_i\) at \(t+\Delta t\) given the
value \(q^0\) at \(t\),

\begin{equation}
0 = R(q^i) = m(q^{i}) - m(q^{0}) - \Delta t \, r(q^i)
\end{equation}

where we have called the quation \(R(q)\). By design \(R(q)\) is smooth
enough to solve with Newton's method without any statemachine logic:

\begin{equation}
\frac{\partial R}{\partial q} \Delta q^{k+1} = R(q^k)
\end{equation}

The function \(R(q)\) has complex logic encoded in it; if-then
statements appear if the model used activations such as rectifiers.

\hypertarget{header-n3339}{%
\subsection{Initialization}\label{header-n3339}}

There is a caveat with representing the degrees of freedom of the system
on a learned space: the user does not know the meaning of \(q\) as an
input. Further, the values of \(q\) vary from network to network and
depend on the parameterization. The user needs to be able to specify the
"real" physical quantities.

The user specifies any two of the values,

Note that the encoder only necessarily works when all four values are
correct. Thus, it is necessary to actually solve the nonlinear problem
on the decoder:

\[D(q^0)_i = s^0_i \quad \text{for specified}\; i\]

This equation has two rows. It is solved with Newton's method until
convergence.

\[K_ij = \frac{\partial D_i}{\partial q_j}  \quad \text{for specified}\; i, \text{for}\;j=1,2\]

\[K\Delta q = s^0_i-D(q)_i\]

The problem is not gaurunteed to converge, however, and often does not.
E.g., if the initial guess was in the liquid phase and the target point
in the EOS space is in a solid, the iteration will fail. To address this
issue, we perform multiple starts with an initial condition in each
phase and equilibrium, discard points that diverge, and then pick the
\(q\) that is closest to the two specified components.

It is possible that none of the initial guesses for this iteration
converge. There are two possiblities: the target state point is
incorrectly specified or out-of-range (e.g., negative density), or the
network is incapable of representing the desired material properties
(i.e., poorly trained).

The weakness of this necessity means that more error checking is needed
in verifying user inputs, but this is necessary in any
multiphase/multicomponent reservoir simulator. LatentSim does not yet
detect and report errors, and only reports divergence. In our study, the
testing problem specifications are all correct, but not all network
realizations are acceptable. Failure to determine a \(q\) for one of the
test initial conditions is a common mode of detecting unacceptable
network realizations; this is marked as a testing error during training
which invalidates a model as a potential realization.

The current implementation LatentSim does not support inputting via
saturations on the equilibria, a commonly used primary variable, but
this problem could be solved with future engineering work for a
non-experimental simulator.

\hypertarget{header-n3493}{%
\subsection{Methodology Summary}\label{header-n3493}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Make a database of \(T,p,\rho,h,X_1,...\)

  \begin{itemize}
  \item
    Piece together empirical fits for each phase from literature
  \item
    (\emph{Experimental data in the future})
  \end{itemize}
\item
  Normalize (and \(\log(p)\)) the database, shuffle it

  \begin{itemize}
  \item
    \(\log(p)\) distributes the low-\(p\) phases evenly w.r.t.
    high-\(p\) phases
  \end{itemize}
\item
  Train the autoencoder on batch-generated architectures. Note that
  phase labels are not used.
\item
  Load the models and generate physics code
\item
  Verify and grade architectures on tests

  \begin{itemize}
  \item
    Need more than autoencoder mean-squared-error
  \item
    Differentiability and numerical stability in simulation
  \item
    Evaluation speed (billions of times in a simulation!)
  \end{itemize}
\item
  Pick best one to package into production code
\end{enumerate}

\hypertarget{header-n3353}{%
\subsection{Pseudo-code}\label{header-n3353}}

\begin{verbatim}
load network from database
build DAE graph
input problem properties and initial value
solve for q0: D(q0) = s0
loop t=0 to t_max:
	solve for q[i]: R(q[i],q[i-1])=0
decode s[i] = D(q[i])
\end{verbatim}

\hypertarget{header-n3356}{%
\section{Implementation}\label{header-n3356}}

The training framework and simulations are implemented in TensorFlow
using custom defined operations. The various visualizations are built
using Matplotlib, Plotly, and Visdom. The accompanying video was
generated using Paraview.

The codebase is released open source at
\url{https://github.com/afqueiruga/latentsim}. Helper methods that the
author reuses in various TensorFlow-based projects are factored out in
the dully-named \url{https://github.com/afqueiruga/afqstensorflowutils}.
The implementation of the empirical equations of state for water are
factored out at
\href{https://github.com/afqueiruga/equations_of_state}{https://github.com/afqueiruga/equations\emph{of}state}.

\hypertarget{header-n3359}{%
\section{Simulation Evaluation}\label{header-n3359}}

Multiple hyperparameters are trained.

Each architecture trained on the autoencoder task is then

Three different extents:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Linear Equation of State

  \begin{itemize}
  \item
    \textbf{Test!}
  \item
    Reduces to single phase Darcy's law problem (slide 1)
  \item
    \( p = 10^5+[-10^3, 10^3] Pa,\quad T = [ 19, 21 ] ^o C\)
  \end{itemize}
\item
  Water Liquid-Gas Regime

  \begin{itemize}
  \item
    One phase boundary 
  \item
    \( p = [100,5\times 10^5] Pa, \quad T = [274,594] K\)
  \end{itemize}
\item
  Water Solid-Liquid-Gas-Supercritical Regimes

  \begin{itemize}
  \item
    No linear mapping to latent space 
  \item
    Entire span
  \item
    \( p = [6\times 10^{-6},3\times 10^8]Pa, \quad T = [150,1000] K \)
  \end{itemize}
\end{enumerate}

\hypertarget{header-n3418}{%
\subsection{Benchmarks}\label{header-n3418}}

The simulator was benchmarke against the long standing TOUGH+ multiphase
multicomponent

Single-grid block simulations were set up for most of the cases. TOUGH+
does not handle supercritical \(H_2O\), so these test problems were not
used to benchmark. (The supercritical phase region is not practically
relevant for water, but it is very important for other fluids such as
\(CO_2\).) (Note that, as a major motivation of this work, adding
supercritical states to the existing Fortran simulator would require
coding of one new logical state with two new state transitions with
primary variable switches; for the approach of this paper, the dataset
just needed to be extended.)

\hypertarget{header-n3421}{%
\section{Conclusions}\label{header-n3421}}

\begin{itemize}
\item
  Deep learning to replace and improve hand-baked equations and
  algorithms
\item
\item
  \textbf{Only training the material representation, not the balance
  laws}
\item
  Extend to more complicated materials
\item
  Put into a flow simulation
\item
  Close loop on testing with reinforcement learning
\end{itemize}

The ultimate goal for this work is to automate integration of data
analysis into simulation.

We hypothesize this end-to-end approach will allow the system to develop
\emph{better} simulations \emph{faster}.

We demonstrated a specially crafted architecture whose parameters can be
human interpretted and perform unsupervised phase/equilibria
classification.

We illustrated a programming model that is able to incorporate machine
learning models into existing well known computational methodologies.
This allows us to enforce physically known quanities, such as the
balance laws.

The TensorFlow system is not quite suited for this type of workflow -\/-
the resulting simualtions are extremely slow due to repeated calls to
session.run and a very complicated graph structure for the polynomials.
Currently, the system is being reimplemented in Julia using Flux and
Zygote to seemlessly reincorporate into a fully-fledged new reservoir
simulator. The pendulum has been reimplemented in Julia at .

We are also demonstrating a new programming paradigm applied to
scientific computing. The entire system has another level of complexity
to it. An offline training phase searches for a new representation, i.e.
for new equations with new unknown degrees of freedom, that are good to
solve.

Each implemented equation of state (three options in this work spanning
different extents of \(H_2O\)) has a database of potential realizations.
The entries are model architectures, parameters, and training
environment configurations and logs for each equation-of-state dataset.
This database of models replaces a version controlled repository of user
written modules. Hundreds of trials for novel primary variable
configurations happen in a matter of minutes on a workstation in place
of pain-staking trial-and-error labor performed with human "intuition"
and experience. We did encode a little bit of human intuition into the
model architectures and tricks for initializing the networks.

The larger simulation code links into the trained database to load the
computer-written model into the computation graph for the simulation,
deriving all needed ingredients for the mainloop using automatic
differentiation.

A fully automated testing system can verify which realizations worked
(relate this to an independent review board comparing the results the
independently developed codes) and then select the fastest and most
robust one.

\end{document}

### 2019-05-13 iapws95.py
these were factored out

@@ -1,163 +1,0 @@
-n_ideal = [
-    -8.32044648201,
-    6.6832105268,
-    3.00632,
-    0.012436,
-    0.97315,
-    1.27950,
-    0.96956,
-    0.24873
-]
-
-gamma = [
-    None,
-    None,
-    None,
-    1.28728967,
-    3.53734222,
-    7.74073708,
-    9.24437796,
-    27.5075105
-    ]
-ni = [
-    0.12533547935523e-1,
-    0.78957634722828e1,
-    -0.87803203303561e1,
-    0.31802509345418,
-    -0.26145533859358,
-    -0.78199751687981e-2,
-    0.88089493102134e-2,
-    -0.66856572307965,
-    0.20433810950965,
-    -0.66212605039687e-4,
-    -0.19232721156002,
-    -0.25709043003438,
-    0.16074868486251,
-    -0.40092828925807e-1,
-    0.39343422603254e-6,
-    -0.75941377088144e-5,
-    0.56250979351888e-3,
-    -0.15608652257135e-4,
-    0.11537996422951e-8,
-    0.36582165144204e-6,
-    -0.13251180074668e-11,
-    -0.62639586912454e-9,
-    -0.10793600908932,
-    0.17611491008752e-,1
-    0.22132295167546,
-    -0.40247669763528,
-    0.58083399985759,
-    0.49969146990806e-2,
-    -0.31358700712549e-1,
-    -0.74315929710341,
-    0.47807329915480,
-    0.20527940895948e-1,
-    -0.13636435110343,
-    0.14180634400617e-1,
-    0.83326504880713e-2,
-    -0.29052336009585e-1,
-    0.38615085574206e-1,
-    -0.20393486513704e-1,
-    -0.16554050063734e-2,
-    0.19955571979541e-2,
-    0.15870308324157e-3,
-    -0.16388568342530e-4,
-    0.43613615723811e-1,
-    0.34994005463765e-1,
-    -0.76788197844621e-1,
-    0.22446277332006e-1,
-    -0.62689710414685e-4,
-    -0.55711118565645e-9,
-    -0.19905718354408,
-    0.31777497330738,
-    -0.11841182425981,
-]
-
[…]

### 2019-05-13 iapws97.py

@@ -1,270 +1,0 @@
-import numpy as np
-import sympy
-from sympy import log, exp
-R = 0.461526e3
-
-triple_point   = T_t,p_t = (273.16, 611.657)
-critical_point = T_c,p_c = (647.096, 22.064e6)
-rho_critical   = rho_c = 322.0
-
-import sympy
-def density_enthalpy(gibbs):
-    T,p = sympy.symbols('T p')
-    g = gibbs(T,p)
-    density = 1/g.diff(p)
-    enthalpy = g - T * g.diff(T)
-    rhovec = np.vectorize(sympy.lambdify([T,p],density))
-    hvec = np.vectorize(sympy.lambdify([T,p],enthalpy))
-                   
-    return (lambda x,y : np.real(rhovec(x,y))),\
-           (lambda x,y : np.real(  hvec(x,y)))
-
-def gibbs_region1(T,p):
-    p1_star = 1.653e7
-    T1_star  = 1.386e3
-    n1 = [ 0.14632971213167e00, -0.84548187169114e00,
-          -3.7563603672040e+00,  3.3855169168385e+00, 
-          -0.95791963387872e00,  0.15772038513228e00,
-          -1.6616417199501e-02,  8.1214629983568e-04, 
-           2.8319080123804e-04, -6.0706301565874e-04,
-          -1.8990068218419e-02, -3.2529748770505e-02, 
-          -2.1841717175414e-02, -5.2838357969930e-05,
-          -4.7184321073267e-04, -3.0001780793026e-04, 
-           4.7661393906987e-05,
-          
-          -4.4141845330846e-06,
-          -7.2694996297594e-16, -3.1679644845054e-05, 
-          -2.8270797985312e-06, -8.5205128120103e-10,
-          -2.2425281908000e-06, -6.5171222895601e-07, 
-          -1.4341729937924e-13, -4.0516996860117e-07,
-          -1.2734301741641e-09, -1.7424871230634e-10, 
-          -6.8762131295531e-19,  1.4478307828521e-20,
-           2.6335781662795e-23, -1.1947622640071e-23, 
[…]

### 2019-05-14 paper/paper.tex
more writing; ignoring latex temps

@@
 \author{Alejandro Francisco Queiruga}
-\date{2018-9-20}
+\date{\today}
 
@@
 without using any phase labels. The decoder end of the model is then
-plugged back into the balance laws, eliminating the constrain equation.
+plugged back into the balance laws, eliminating the constraint equation.
 \end{abstract}
@@
 simulation runtime and robustness. Further, we must remember the value
-of the human effort involved: beyond improving computational aspects, we
-are fully automating part of the human involvement in programming the
-simulator.
+of the human effort involved: beyond improving computational aspects,
+we are fully automating part of the human involvement in programming
+the simulator.
 
@@
 human-written code with computer-written code. We are applying this
-mindset from the computer-vision and advertisement dominated Artificial
+mindset from industry problems in the Artificial
 Intelligence space into scientific computing.
@@
 
-The original redescribe our problem as a differential algebraic equation
+We redescribe our problem as a differential algebraic equation
 with a constraint: the balances of mass and energy are constrained by
@@
 
-We train in two phases. The first step is minibatched stochastic
-gradient descent on the entire deep network. After a few epochs of this,
-we do a step of linear regression on the last layer of the network,
-denote by \(W^{dec}\), which directly controls the values. The linear
+The dataset is split into a standard $2/3-1/6-1/6$ division into
+training-testing-validation groups. We note that we can always
+manufacture more datapoints. We train by alternating phases. The first
+step is minibatched stochastic descent on the entire deep network, assembling the losses on minibatches of the data, $X_{mini}$,
+\begin{equation}
+  L(X_{mini}) = \sum_{x\in X_{mini}} L(x)
+  \end{equation}
+to perform the out-of-the-box ADAM optimizer on all of the
+parameters in the networks. (The box is
+TensorFlow.) After a few epochs of this,
+the training algorithm alternates to a few steps step of linear regression on the last layer of the network,
+denoted by \(W^{dec}\), which directly controls the magnitudes of the values.
+(The bias $b^{dec}$ is included in $W^{dec}$.) The linear
 system that needs to be solved is easily generated by assembling the
-Hessian of the total loss function,
+gradient and Hessian of the total loss function,
 \begin{align}
- \mathbf{G} &= \sum_{x\in X}\frac{\partial L}{\partial W^{dec}} \\
-\mathbf{H} &= \sum_{x\in X}\frac{\partial^2 L}{\partial (W^{dec})^2}
+ \mathbf{G}\left(X_{train}\right) &= \sum_{x\in X_{train}}\frac{\partial L}{\partial W^{dec}} \\
+\mathbf{H}\left(X_{train}\right) &= \sum_{x\in X_{train}}\frac{\partial^2 L}{\partial (W^{dec})^2}
 \end{align}
@@
 To form the hessian and gradient for this step, we sum over all of the
-training data. (The bias \(b^{dec}\) is included in \(W^{dec}\).) (The
-contractive penalty term drops out because it does not depend on
-\(W^{dec}\).)
+training data. (The contractive penalty term drops out because it does
+not depend on \(W^{dec}\).)
 
@@
 \end{equation}
-
-The function \(R(q)\) has complex logic encoded in it; if-then
+This equation is deceptively simple: the function \(R(q)\) has complex logic encoded in it; if-then
 statements appear if the model used activations such as rectifiers.
+However, the complex material behavior which traditionally requires
+the vast majority of programming was implemented automatically by the
+models; implementing equations XX, XX, and XX require only a few lines
+of human-written code.
 
@@
 
-The user specifies any two of the values,
+Since the original manifold is 2D, the user only needs to specify two
+initial values. In a multiphase solver such as TOUGH, the state needs
+to be specified with the two primary variables plus the state tag: 
+\begin{verbatim}
+9e5 20 Aqu
+0.5 20 AqG
+\end{verbatim}
+Note that the first primary variable means pressure in the first line
+for a pure-liquid state,
+but saturation in the second line for the liquid-gas equilibrium.
+This makes sense inuitively, as this is how humans are taught, but
+still leads to bugs and occasionally requires looking up values in
+textbooks or the manual. (Is it $S_{gas}$ or $S_{aqu}$ on the second
+line?) A fully fledged solver needs to check the user specified
+initial conditions for erroneous inputs.
+
+For LatentSim, the input system is not yet as robust as a
+fully-fledged user-facing simulator, but the inputs are specified in
+the $T p \rho h$ space instead of the $q$ space.
+The user (or test problem) specifies any two of the four values in a
+dictionary format,
+\begin{verbatim}
+p=9e5 T=20
+rho=500 T=20
+\end{verbatim}
+Then, the simulator must determine $q(t=0)$ from these two
+values. This sets up a system of two equations.
 
@@
 on the decoder:
+\[D(q^0)_i = s^0_i \quad \text{for specified}\; i\]
+This equation has two rows indexing into the four state variables;
+e.g. only the rows associated with $p$ and $T$ are solved if those are
+specified, which is usually the case. The above equation is solved
+with Newton's method. The two corresponding rows are cut out of
+the tangent matrix to $D$,
+\[K_{ij} = \frac{\partial D_i}{\partial q_j}  \quad \text{for specified}\; i, \text{for}\;j=1,2\]
+and the following system of equations is solved for an update to $q$,
+\[K\Delta q = s^0_i-D(q)_i\]
+Updating $q$ is repeated until convergence.
 
-\[D(q^0)_i = s^0_i \quad \text{for specified}\; i\]
+The initial condition problem is not gaurunteed to converge, however,
+and often does not.
+For example, if the initial guess was in the liquid phase and the target point
+in the EOS space is in a solid, the iteration will fail.
+To address this issue, we perform multiple starts with an initial condition in each
+phase and equilibrium, discard points that diverge, and then pick the
+\(q\) that is closest to the two specified components. The multiple
+start points are specified at EOS development time to form a set of
+values that represent middle points for each phase,
+\begin{equation}
+\left\{s_{guess}\right\} = \left\{ \left(\begin{array}{c}
+\bar{T}_{gas}\\
+\bar{p}_{gas}\\
+\bar{\rho}_{gas}\\
+\bar{h}_{gas}
+\end{array}\right),\left(\begin{array}{c}
+\bar{T}_{liq}\\
+\bar{p}_{liq}\\
+\bar{\rho}_{liq}\\
+\bar{h}_{liq}
+\end{array}\right),\left(\begin{array}{c}
+\bar{T}_{ice}\\
+\bar{p}_{ice}\\
+\bar{\rho}_{ice}\\
+\bar{h}_{ice}
+\end{array}\right)... \right\}
+  \end{equation}
+The multistart method could be specified without any prior knowledge
+with random points, but this is a sufficiently robust method for
+now. Each of the possible guess points is used to fill in the two
+unspecified values as guesses and then passed into the encoder.
+E.g. if $p$ and $T$ were specified, each possible $\bar{\rho}$ and
+$\bar{h}$ combination from the above sets are filled in to provide
+multiple guesses for $q$,
+\begin{equation}
+s^0 = \left(\begin{array}{c}
+T^*(0)\\
+p^*(0)\\
+\bar{\rho}_{guess}\\
+\bar{h}_{guess}
+\end{array}\right);\quad q^0_{guess}=E\left(s^0\right)
+ \end{equation}
+ These sets of $q^0_{guess}$ are then iterated upon using the Newton's
+ method procedure. We further add
+ under-relaxation to the iteration,
+ \begin{equation}
+q^0_{k+1} =q^0_{k} + \alpha \Delta q_{k}
+\end{equation}
+with the under relaxation parameter usually set to $\alpha \approx
+0.1$.
 
-This equation has two rows. It is solved with Newton's method until
-convergence.
-
-\[K_ij = \frac{\partial D_i}{\partial q_j}  \quad \text{for specified}\; i, \text{for}\;j=1,2\]
-
-\[K\Delta q = s^0_i-D(q)_i\]
-
-The problem is not gaurunteed to converge, however, and often does not.
-E.g., if the initial guess was in the liquid phase and the target point
-in the EOS space is in a solid, the iteration will fail. To address this
-issue, we perform multiple starts with an initial condition in each
-phase and equilibrium, discard points that diverge, and then pick the
-\(q\) that is closest to the two specified components.
-
-It is possible that none of the initial guesses for this iteration
+ It is further possible that none of the initial guesses for this iteration
 converge. There are two possiblities: the target state point is
@@
 
+The simulation inputs are the name of the equation of state, .e.g,
+``water\_lg'' or ``water\_slgc'', and the name of the architecture
+hyperparameters, e.g. ``Classifying\_pT\_1.0\_1,3,6,12,Sigmoid''.
+In a mature code, the name of the architecture loaded would be hidden
+from the end-user and would be a setting which the doman scientist and developer freezes
+after checking the training process.
+
 A fully automated testing system can verify which realizations worked

### 2019-05-14 paper/paper.tex
writing more

@@
 these relations in the rest of our analyses and production-scale
-computer simulations? A single material with multiple phase transitions
+computer simulations?
+
+Accurately simulating multispecies reaction and transport with
+undergoing phase transitions is a critical aspect to a variety of
+problems.
+Motivations for this work are subsurface energy systems such as shale
+and oceanic natural gas reservoirs in shales or engineered geothermal
+systems.
+
+A single material with multiple phase transitions
 is a challenge in of itself, and the descriptions become combinatorially
@@
 
-Motivations for this work are subsurface energy systems such as shale
-and oceanic natural gas reservoirs in shales or engineered geothermal
-systems.
-
-This work is only considered with the "single beaker" problem; the flow
+The system of equations we are looking at are, for only two components,
+\begin{align}
+  \partial_t \rho_{H_2O} =& \nabla \cdot flux_{H_2O} + s \\
+  \partial_t \rho_{CO_2} =& \nabla \cdot flux_{CO_2} + s \\
+  \partial_t e& = \nabla \cdot flux_{heat} + s
+\end{align}
+More species add more rows, and applicaitons also involve continuum
+mechanics and chemical reactions! The interplay of deformations,
+reactions, and phase transitions makes the equations extremely difficult
+to solve. \cite{my millstone paper}
+
+This work only considers the "single beaker" problem; the flow
 and transport aspects have similar challenges. (The next stage of this
@@
 
-This work proposes a new methodology that steps away from the
-traditional approaches
-
-In this approach, we directly \emph{look for solvable representations}
-instead of representations that carry human intuition. We value
-simulation runtime and robustness. Further, we must remember the value
-of the human effort involved: beyond improving computational aspects,
-we are fully automating part of the human involvement in programming
-the simulator.
+In section 2, we will cast the above equation into a basic form for a
+differential algebraic equation,
+\begin{align}
+  z( u, \partial_t u) &= 0 \\
+  c(u) & = 0
+\end{align}
+We will introduce a new way of thinking about constraints as sets of
+solutions or observations of possible $u$:
+\begin{equation}
+c(u)=0 \leftrightarrow \left\{ u | c(u)=0 \;\text{or}\; u\; \text{observed} \right\}
+\end{equation}
+We will use this set of possible values viewpoint to train an
+autoencoder on the set of $u$ for a latent space parameterized by $q$:
+\begin{equation}
+u = D(q), \; q=E(u) \;\forall u \in U
+\end{equation}
+which satisifies
+\begin{equation}
+  D(q)\in U \forall q.
+\end{equation}
+We can then write the original equation by
+\begin{equation}
+z(q,\partial_t q)=0
+\end{equation}
 
@@
 
-Consider a general differential algebraic equation of the form
-
-\begin{equation}
-z(\dot{u},u) = 0
-\end{equation}
-
-We are interested in equations that have the form
-
-\begin{align}
-\frac{\mathrm{d}u}{\mathrm{d}t} & = v \\
-0 & = c(u,v)
-\end{align}
-
-Where do such equations arise? A simple example is the pendulum, which
-will be explored in Section 2.
-
-\begin{align}
-\text{Solve for}\, x(t), \, y(t), \, f(t) \, \text{satisfying:} \\
-m \dot{x} & = f x/L \\
-m \dot{y} & = f y/L + m g \\
-x^2 + y^2 & = L^2
-\end{align}
+This work proposes a new methodology that steps away from the
+traditional approaches. 
+In this approach, we directly \emph{look for solvable representations}
+instead of representations that carry human intuition. We value
+simulation runtime and robustness. Further, we must remember the value
+of the human effort involved: beyond improving computational aspects,
+we are fully automating part of the human involvement in programming
+the simulator.
+
+In section 2, the balance of mass and energy will be described for a
+single multiphase material. We will discuss the current state of the
+art algorithm for deriving relationships for multiphase materials, and
+then the primary-variable switching algorithm that is used in
+state-of-the art solvers. The equation will then be recast as a
+constraint problem in th above form.
+In section 3, we will demonstrate applying the autoencoder learning
+methodology to solve the pendulum,
+which can be considered the archetypical constraint problem.
+In section 4, the autoencoder methodology wil lbe applied to the
+equation of state of water. In section 5, the new simulation process
+will be described, wherein the learned decoders are placed into the
+balance laws, forming newly discovered equations on learned
+representations. The software implementation of the training and
+simulation programs are detailed in section 6. The new simulation
+framework on sets of learned representations is evaluated on
+multiple datasets of varying extents of water with a
+hyperparameter search on a suite of test problems to prouce the discussion of section 7, as
+well as a systematic automated testing procedure of a
+machine-dominated programming methodology.
+Active research and software development directions are
+concluded in the final section.
+
+The notebooks and software implementations are released open source to
+accompany this work. They can be found at the github repository
+located at: \href{https://github.com/afqueiruga/latentsim}
+Certain components are factored into submodule repositories which are
+each released under the LGPL at seperate repositories referenced in
+the implementation details.
 
@@
 \section{Multiphase Equations of State}\label{header-n3234}}
-
-We also consider the phase equilibria; at a larger scale it is possible
-for a control volume to be a mixture of two phase coexisting, such as
-boiling with some mixture saturation \(S_{gas}=1-S_{liq}\).
 
@@
 are not well defined functions due to the presence of phase changes that
-yield sharp discontinuities in \(p,T\), as shown in Figure 1 for water.
+yield sharp discontinuities in \(p,T\), as shown in Figure 1 for
+water.
+
+We also consider the phase equilibria; at a larger scale it is possible
+for a control volume to be a mixture of two phase coexisting, such as
+boiling with some mixture saturation \(S_{gas}=1-S_{liq}\).
+(For a very fine scale simulation, bubbles would be explicitly modeled.)
 The material can exist on this section of the surface as a mixture of
@@
 \[\rho = X_{H_2O}\left(S_{gas} \rho_{gas} + S_{liq} \rho_{liq} + S_{ice} \rho_{ice} + S_{crit} \rho_{crit} \right)\]
-
-We will also note right now that applicaitons also involve continuum
-mechanics and chemical reactions! The interplay of deformations,
-reactions, and phase transitions makes the equations extremely difficult
-to solve.
 
@@
 \includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/phase_diagram.png}
-\caption{}
+\caption{Phase boundaries for water }
 \end{figure}
@@
 \includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/water_eos.png}
-\caption{}
+\caption{\label{fig:pTrho} The $\log{p}$, $T$, $\rho$ equation of
+  state surface for water. This plot spans four phases, three
+  equilibria, and the triple point, which are each given their own color.
+  There is a fourth dimension to the surface of enthalpy $h$ that exhibits
+  similar discontinuities; the $\log{p}$, $T$, $h$ surface is plotted
+  in the Appendix.}
 \end{figure}
+
+There are even more known phases of crystal ice which for water
+For water, the supercritical phase and exotic ices only occur in
+laboratory conditions or naturally near the
+crust \cite{}. For other materials these exotic phases can be
+applicable. For example, in geoengineered systems using $CO_2$
+such as geothermal plants are carbon sequestration facilities, the
+working fluid is usually in the supercritical regime at depth.
 
@@
 
-An archetypical constraint problem the pendulum. Posed as dynamics, it
+An archetypical constraint problem the pendulum. Posed as dynamics
+problem with a Lagrange-multipler-esque centripetal force, it
 has three unknowns wtih two second-order equations and a constraint,
@@
 \text{Solve for}\, x(t), \, y(t), \, f(t) \, \text{satisfying:} \\
-m \ddot{x} & = f x/L \\
-m \ddot{y} & = f y/L + m g \\
-x^2 + y^2 & = L^2
+m \ddot{x} & = f x/\sqrt{x^2+y^2} \\
+m \ddot{y} & = f y/\sqrt{x^2+y^2} - m g \\
+x^2 + y^2 & = R^2
 \end{align}
 
-In our methodology, we do not have an explicit equation for the
-constraint, but a dataset of (x,y) pairs that are on the manifold. For
-our proof-of-concept test, we manufacture this data for the pendulum.
 Note that we are considering the nonlinear pendulum.
@@
 multipliers (\(f(t)\) is the multiplier in the way we have written it
-now), penalties, change of variables, etc. The change of variables
-method is to introduce a new variable \(\theta\) that parameterizes
-\(x=L\cos\theta\) and \(y=L\sin\theta\):
-
+above), penalties, change of variables, etc. The change of variables
+method exploits the Lagrangian mechanics formulation, where we state
+the Lagrangian,
+\begin{equation}
+L = \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right) - m g y
+\end{equation}
+with the constraint that \(x\) and \(y\) lie on the path.
+We can introduce a new variable \(\theta\) that parameterizes
+\(x=R\cos\theta\) and \(y=R\sin\theta\) and rederive the equations of
+motion with the familiar,
 \begin{align}
-\text{Solve for}\, \theta(t) \text{and} \, f(t) \, \text{satisfying:} \\
-m \ddot{x}(\theta) & = f x(\theta)/L \\
-m \ddot{y}(\theta) & = f y(\theta)/L + m g \\
+\text{Solve for}\, \theta(t) \ \text{satisfying:} \\
+\frac{\mathrm{d}}{\mathrm{d}t} \frac{\partial L}{\partial
+  \dot{\theta}} =
+  \frac{\partial L}{\partial \theta}
 \end{align}
-
 In this example, this lets us collapse the problem into a
-single-component Ordinary Differential Equation.
+single-component second-roder differential equation.
+
+In our methodology, we do not have an explicit equation for the
+constraint, but a dataset of $X$ of $(x,y)$ pairs that are on the manifold. For
+our proof-of-concept test, we manufacture this data for the circular
+pendulum to verify against the analytical solution:
+\begin{equation}
+X = \left\{ (R\cos\theta_i, R\sin\theta_i)\; | \;\theta_i \in [-5\pi/4,\pi/4] \right\}
+\end{equation}
+We also try a ``goofier'' looking curve with no known closed form
+solution to exercise the methodology,
+
+(For the equation of state problem, this is analogous to the dataset
+of $p,T,\rho,h$ 4-tuples that make the plot in Figure \ref{fig:pTrho}.)
 
@@
 polynomials forms the autoencoder:
-
 \[\left\{\begin{array}{c} x\\y\end{array}\right\}
-\rightarrow W^1 x^n \rightarrow q \rightarrow W^2 x^n \rightarrow 
+\rightarrow W_{enc} x^n \rightarrow q \rightarrow W_{dec} x^n \rightarrow 
 \left\{\begin{array}{c} x\\y\end{array}\right\}\]
@@
 (We explicitly write the bias \(+b\); it is equivalent to including
-\(x^0=1\) in the polynomial set.) The Lagrangian for the pendulum is
-
-\begin{equation}
-L = \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right) - m g y
-\end{equation}
-
-with the constraint that \(x\) and \(y\) lie on the path. The equations
+\(x^0=1\) in the polynomial set.)
+The equations
 of motion given some parameter space that satisfies the constraint is

### 2019-05-15 batch_test_latent_sim.py
added two more tets

@@ -1,6 +1,6 @@
 from __future__ import print_function
 import sys,inspect
-import iapws97
+from equations_of_state import iapws97
 
 class Linear_Liquid():
     t_max = 1.0
@@ -64,6 +64,28 @@ class Transition_L2G():
     @staticmethod
     def schedule(sim,t):
         sim.set_params(T_inf=350,p_inf=5.0e3)
+    answer = dict(T=350,p=5.0e3,
+        rho= iapws97.density_region2(350.0,5.0e3),
+        h  =iapws97.enthalpy_region2(350.0,5.0e3) )
+    
+class Transition_L2G_Drain():
+    t_max = 10.0
+    initial = dict(T=350,p=5.0e5, phase="Liquid")
+    params =  dict(mass_source=-0.1,k_p=0.0,k_T=0.0, Dt=t_max/1000.0)
+    @staticmethod
+    def schedule(sim,t):
+        pass
+    answer = dict(T=350,p=5.0e3,
+        rho= iapws97.density_region2(350.0,5.0e3),
+        h  =iapws97.enthalpy_region2(350.0,5.0e3) )
+    
+class Liquid_Drain():
+    t_max = 2000.0
+    initial = dict(T=373.0,p=3.0e5, phase="Liquid")
+    params =  dict(mass_source=-0.1,k_p=0.0,k_T=0.0, Dt=t_max/1000.0)
+    @staticmethod
+    def schedule(sim,t):
+        pass
     answer = dict(T=350,p=5.0e3,
         rho= iapws97.density_region2(350.0,5.0e3),
         h  =iapws97.enthalpy_region2(350.0,5.0e3) )

### 2019-05-15 latent_sim.py
Adding enthalpy transport

@@ -84,8 +84,8 @@ class LatentSim():
             None: s0_none,
             "Gas": np.array([[450,1.0e2, 0.00048149,2835269.40]]),
             "Liquid": np.array([[350,1.0e7,978.09,329726.06]]),
-            "Solid": np.array([[250,1.0e5,919.87,-379930.33]]),
-            "Supercritical": np.array([[900,132167913.14,900,2964049.86]])
+#             "Solid": np.array([[250,1.0e5,919.87,-379930.33]]),
+#             "Supercritical": np.array([[900,132167913.14,900,2964049.86]])
         }
         # One guess:
 #         s0 = guesses[phase]
@@ -205,9 +205,10 @@ class LatentSim():
         
         m = tf.concat([rho,
                        rho*h-p],axis=-1)
+        mflux = k_p*(p_inf-p)+mass_source
         rate = tf.concat([
-            k_p*(p_inf-p)+mass_source,
-            k_T*(T_inf-T)+heat_source],axis=-1)
+            mflux,
+            h*mflux + k_T*(T_inf-T)+heat_source],axis=-1)
         return m, rate
     
     def set_params(self, **kwargs):
@@ -215,7 +216,7 @@ class LatentSim():
         for k,v in kwargs.items():
             self._vars[k].load(v,self._sess)
     
-    def solve_a_time_step(self, q0):
+    def solve_a_time_step(self, q0, under_relax=0.1):
         """Solve one timestep. Note that LatentSim is stateless in this 
         regard."""
         qi = q0.copy()
@@ -227,7 +228,8 @@ class LatentSim():
             nDq = np.linalg.norm(Dq)
             if nDq<1.0e-14: 
                 break # It might be 0, we're done
-            step = min(0.001/nDq,1.0)*Dq
+            # step = min(0.001/nDq,1.0)*Dq
+            step = under_relax*Dq
             # TODO line search
             qi[:] += step
             if nDq<2.0e-7: break

### 2019-05-15 training_water_lg.py
refactored training parameters and architecture parameters

@@ -3,51 +3,27 @@ if __name__=="__main__":
     data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
 #     data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
     dataset = "water_lg_sharded/*.csv"
-    
-    sets_to_try = [
-#         {'type':'Classifying','args':[1,1, 6,12,'tanh']},
-#         {'type':'Classifying','args':[1,1, 6,12,'sigmoid']},
-#         {'type':'Classifying','args':[1,3, 6,12,'tanh'],'ini':'rand','cae':0},
-#         {'type':'Classifying','args':[1,3, 6,12,'tanh'],'ini':'pT','cae':0},
-#         {'type':'Classifying','args':[1,3, 6,12,'tanh'],'ini':'rhoh','cae':0},
-#         {'type':'Classifying','args':[1,3, 6,12,'tanh'],'ini':'rand','cae':1.0},
-#         {'type':'Classifying','args':[1,3, 6,12,'tanh'],'ini':'pT','cae':1.0},
-#         {'type':'Classifying','args':[1,3, 6,12,'tanh'],'ini':'rhoh','cae':1.0},
-        {'type':'Classifying','args':[1,3, 6,12,'sigmoid'],'ini':'rand','cae':0},
-        {'type':'Classifying','args':[1,3, 6,12,'sigmoid'],'ini':'pT','cae':0},
-        {'type':'Classifying','args':[1,3, 6,12,'sigmoid'],'ini':'rhoh','cae':0},
-        {'type':'Classifying','args':[1,3, 6,12,'sigmoid'],'ini':'rand','cae':1.0},
-        {'type':'Classifying','args':[1,3, 6,12,'sigmoid'],'ini':'pT','cae':1.0},
-        {'type':'Classifying','args':[1,3, 6,12,'sigmoid'],'ini':'rhoh','cae':1.0},
-#         {'type':'Classifying','args':[1,4, 6,12,'sigmoid']},
-#         {'type':'Classifying','args':[1,5, 6,12,'sigmoid']},
-#         {'type':'Classifying','args':[1,1, 6,12,'relu']},
-#         {'type':'Deep','args':[1,[],1,[8,8,8]]},
-#         {'type':'Deep','args':[1,[],2,[8,8,8]]},
-#         {'type':'Deep','args':[1,[],3,[8,8,8]]},
[…]

### 2019-05-15 paper/paper.tex
added in bibliography

@@
 \usepackage{ifxetex,ifluatex}
+\usepackage{chemmacros}
+\usepackage{natbib}
+\usepackage{lineno}
 \usepackage[margin=1in]{geometry}
@@
 
+% Patch equations to properly generate line numbers for review
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+  \renewenvironment{#1}%
+     {\linenomath\csname old#1\endcsname}%
+     {\csname oldend#1\endcsname\endlinenomath}}%
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+  \patchAmsMathEnvironmentForLineno{#1}%
+  \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
+
 % set default figure placement to htbp
@@
 reactions, and phase transitions makes the equations extremely difficult
-to solve. \cite{my millstone paper}
+to solve. \cite{queiruga2018millstone}
 
@@
 \end{equation}
+
+ \cite{wu_physics-informed_2018}
+\cite{xie_tempogan:_2018}
 
@@
 
-Solve for \(q_1(t)\) and \(q_2(t)\) such that:
+With the trained model, we solve for \(q_1(t)\) and \(q_2(t)\) such that:
 
@@
 
-\textbf{Only training the material representation, not the balance
-laws.}
+Only training the material representation, not the balance
+laws.
 
@@
 
-The autoencoder parameterizes the constraint :
+The decoder part of the network is inserted into the balance law
 
@@
 
+Technique ssuch as L1 regularization would be needed to search for low
+cost models that also satisfy our desire for Occam's razor
+simplifications. (The IAPWS relations used a prelimary form of this
+type of simplification, but are not optimal in the author's opinion.)
+
 A fully automated testing system can verify which realizations worked
@@
 
+\bibliographystyle{plainnat}
+\bibliography{zotero}
+
 \end{document}

### 2019-05-16 autoencoder_trainer.py
trying to get more information out of the training

@@ -94,7 +94,11 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
         loghook = tf.train.SummarySaverHook(
             summary_op=[tf.summary.scalar("goaltrain",ae.goal_all),
                         tf.summary.scalar("goaltest",ae.goal_test)],
-            save_steps=25,output_dir=training_dir)
+            save_steps=1,output_dir=training_dir)
+        # Print to screen hook
+        @DoStuffHook(freq=1)
+        def printgoalhook(ctx,run_values):
+            print(ctx.session.run(ae.goal_all))
         stophook = tf.train.StopAtStepHook(num_steps=n_epoch)
         saverhook = SaveAtEndHook(training_dir+"/final_variables")
         # Make a closure into a hook
@@ -104,7 +108,7 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
             onum_val = onum.eval(session=ctx.session)
             onum_val+=1
             onum.load(onum_val, session=ctx.session)
-            print(ae.goal.eval(session=ctx.session))
+            print("extra:",ae.goal.eval(session=ctx.session))
             header="T, p, rho, h"
             ae.save_fit(training_dir+"/surf_{0}.csv".format(onum_val),
                         header,sess=ctx.session)
@@ -113,16 +117,21 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
         def newthook(ctx,run_values):
             # Now do the Hessian step on the last layer
             print("Running the newton step")
-            for i in range(5):
+            for i in range(1):
+                print("newt:",ctx.session.run(ae.goal_all))
                 ctx.session.run(ae.newt_step)
+            print("newt:",ctx.session.run(ae.goal_all))
+
         # set up the session
         session = tf.train.MonitoredTrainingSession(
             checkpoint_dir=training_dir,
-            hooks=[loghook,stophook,saverhook,extrahook,newthook])
+            hooks=[loghook,printgoalhook,saverhook,extrahook,newthook,stophook])
         # train away
         with session as sess:
             # Do the SGD rounds
             while not sess.should_stop():
+                print("loop:",sess.run(ae.goal_all))
                 sess.run(ae.train_step)
+                
 
     

### 2019-05-16 training_water_lg.py

@@ -3,22 +3,25 @@ if __name__=="__main__":
     data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
 #     data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
     dataset = "water_lg_sharded/*.csv"
-    inis = [ 'rand', 'pT', 'rhoh' ]
-    caes = [ 0.0, 1.0 ]
-    extend_train_params = lambda arch : [ dict(ini=ini,cae=cae,**arch) for ini in inis for cae in caes ]
+#     inis = [ 'rand', 'pT', 'rhoh' ]
+#     caes = [ 0.0, 1.0 ]
+    inis = [ 'rand', ]# 'pT', 'rhoh' ]
+    caes = [ 0.0 ]
+    extend_train_params = lambda arch : \
+        [ dict(ini=ini,cae=cae,**arch) for ini in inis for cae in caes ]
     archs_to_try = [
+        {'type':'Classifying','args':[1,3, 3,3,'sigmoid']},
 #         {'type':'Classifying','args':[1,3, 6,12,'sigmoid']},
-        {'type':'Classifying','args':[1,1, 6,12,'tanh']},
-        {'type':'Classifying','args':[1,1, 6,12,'sigmoid']},
-        {'type':'Poly','args':[1,4]},
+#         {'type':'Classifying','args':[1,1, 6,12,'tanh']},
+#         {'type':'Classifying','args':[1,1, 6,12,'sigmoid']},
+#         {'type':'Poly','args':[1,4]},
 #         {'type':'Poly','args':[1,5]},
     ]
     sets_to_try = []
     for arch in archs_to_try:
         sets_to_try.extend(extend_train_params(arch))
 
-
-    n_epoch = 1000
+    n_epoch = 100
 
     import multiprocessing as multi
     import itertools
@@ -29,7 +32,7 @@ if __name__=="__main__":
                           training_dir=training_dir,
                           data_dir = data_dir,
                           n_epoch = n_epoch)
-    p = multi.Pool(processes=4)
-    p.map( job, sets_to_try )
-#     for S in sets_to_try:
-#         job(S)
+#     p = multi.Pool(processes=1)
+#     p.map( job, sets_to_try )
+    for S in sets_to_try:
+        job(S)

### 2019-05-17 autoencoder_trainer.py
The callbacks happene every _run_ call. How odd

@@ -130,8 +130,8 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
         with session as sess:
             # Do the SGD rounds
             while not sess.should_stop():
-                print("loop:",sess.run(ae.goal_all))
-                sess.run(ae.train_step)
+                print("loop:",sess.run([ae.goal_all,ae.train_step]))
+
                 
 
     

### 2019-05-17 paper/paper.tex
More writing about the paper

@@
 \text{Solve for}\, \rho(t), \, h(t), \, p(t),\, \text{and}\, T(t)\, \text{satisfying:}\\
-\partial_t \rho & = \nabla \cdot \mathbf{k}\nabla p + r\\
-\partial_t \rho h - p & = \nabla \cdot \mathbf{k'}\nabla T + s\\
+\partial_t \rho &  = \mathbf{k}_p(p_\infty - p) + r\\
+\partial_t (\rho h-p) & = \mathbf{k}_T(T_\infty-T) + s \\
 \text{such that they lie on the material EOS,}\\
-EOS(\rho,h,p,T) & = 0
+\rho,h,p,T & \in \{\text{EOS}\}
 \end{align}
@@
 feature space by
-
 \begin{equation}
@@
 \end{equation}
-
 and similarly for the \(y\) direction.
-
-Taking the partial derivatives with respect to the feature variable and
-its rate, the components to the equation of motion are
-
-\begin{equation}
-\frac{\partial L}{\partial q} = m g \frac{\partial y}{\partial q}
-\end{equation}
 
@@
 same as \(x(\theta)\), plugging in
-
 \[L(x(q),v(q,\dot{q}))\]
-
-and build the DAE we know from physics:
+Taking the partial derivatives with respect to the feature variable and
+its rate, the components to the equation of motion are build the DAE we know from physics:
 
 \[\frac{\mathrm{d}}{\mathrm{d}t}\frac{\partial L}{\partial \dot{q}} = \frac{\partial L}{\partial q}\]
-
 where we get the velocity with the chain rule:
-
 \[v = \frac{\partial x}{\partial q} \dot{q}\]
@@
 method:
-
 \[\frac{\partial L}{\partial \dot{q}}_i - \Delta t \frac{\partial L}{\partial q}_i = \frac{\partial L}{\partial \dot{q}}_0 + \Delta t \frac{\partial L}{\partial q}_0\]

### 2019-06-09 paper/paper.tex
writing up edits

@@
 \hypersetup{
-            pdftitle={Solution of Differential Algebraic Equations with Manifold Constraints on the Latent Space of an Autoencoder},
+            pdftitle={Solution of Differential Algebraic Equations with Local Constraints on the Latent Space of an Autoencoder},
             pdfauthor={Alejandro Francisco Queiruga},
@@
 \begin{abstract}
-We train an autoencoder to determine the lower-dimensional
-representation of the EOS surface using modern deep learning techniques
-without using any phase labels. The decoder end of the model is then
-plugged back into the balance laws, eliminating the constraint equation.
-\end{abstract}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\hypertarget{header-n3216}{%
-\section{Introduction}\label{header-n3216}}
-
-Representing complex physical systems with multiple speicies, phases,
-chemical reactions, and heterogenous scale effects poses an
-intractability problem to modeling. How do we analyze complex
-experimental data and \emph{ab initio} simulations and derive new
-macroscale relations to describe the systems? How do we then implement
-these relations in the rest of our analyses and production-scale
-computer simulations?
-
-Accurately simulating multispecies reaction and transport with
-undergoing phase transitions is a critical aspect to a variety of
-problems.
-Motivations for this work are subsurface energy systems such as shale
-and oceanic natural gas reservoirs in shales or engineered geothermal
-systems.
-
-A single material with multiple phase transitions
-is a challenge in of itself, and the descriptions become combinatorially
-more complex as species and phases are added. This results in
-combinatorially more human labor and combinatorially more code which
-tends to involve complex branching logic. The descriptions then pose
-numerical challenges when incorporated into macroscale simulations due
-to obstacles such as discrete phase state machines and primary variable
-switching.
-
-The system of equations we are looking at are, for only two components,
-\begin{align}
-  \partial_t \rho_{H_2O} =& \nabla \cdot flux_{H_2O} + s \\
-  \partial_t \rho_{CO_2} =& \nabla \cdot flux_{CO_2} + s \\
-  \partial_t e& = \nabla \cdot flux_{heat} + s
-\end{align}
-More species add more rows, and applicaitons also involve continuum
-mechanics and chemical reactions! The interplay of deformations,
-reactions, and phase transitions makes the equations extremely difficult
-to solve. \cite{queiruga2018millstone}
-
-This work only considers the "single beaker" problem; the flow
-and transport aspects have similar challenges. (The next stage of this
-work will "connect the beakers".)
-
-In section 2, we will cast the above equation into a basic form for a
-differential algebraic equation,
-\begin{align}
-  z( u, \partial_t u) &= 0 \\
-  c(u) & = 0
-\end{align}
-We will introduce a new way of thinking about constraints as sets of
-solutions or observations of possible $u$:
-\begin{equation}
-c(u)=0 \leftrightarrow \left\{ u | c(u)=0 \;\text{or}\; u\; \text{observed} \right\}
-\end{equation}
-We will use this set of possible values viewpoint to train an
-autoencoder on the set of $u$ for a latent space parameterized by $q$:
-\begin{equation}
-u = D(q), \; q=E(u) \;\forall u \in U
-\end{equation}
-which satisifies
-\begin{equation}
-  D(q)\in U \forall q.
-\end{equation}
-We can then write the original equation by
-\begin{equation}
-z(q,\partial_t q)=0
-\end{equation}
-
- \cite{wu_physics-informed_2018}
-\cite{xie_tempogan:_2018}
-
-Part of the new modern software engineering framework involves replacing
-human-written code with computer-written code. We are applying this
-mindset from industry problems in the Artificial
-Intelligence space into scientific computing.
-
-The actual meaning of the parameterization is not important, so long as
-it can be effectively solved and then \emph{decoded} back to the
-physically meaningful quantities.
-
-We redescribe our problem as a differential algebraic equation
-with a constraint: the balances of mass and energy are constrained by
-the EOS. Because the EOS was originally derived from data, a data-driven
-approach lends itself in which the EOS is a set of points in
-temperature-pressure-density-enthalpy space. (In the multicomponent
-cases there are more dimensions.)
-
-We demonstrate this approach to solving constrained differential
-equations on the pendulum, before demonstrating two EOSes of water: one
-with just liquid-gas phases, and one with liquid-gas-ice-supercritical
-regimes. The notion of phases does not appear in this multiphase
-simulation. We show a model architecture that is able to learn phase
-labels in the unsupervised autoencoder training process, such
-human-scientist interpretable results are obtained from the machine
-learning process. We compare this and other model architectures,
-including common deep neural networks, based on autoencoder error and
-simulation success metrics. We stress that the presented approach is
-only utilizing machine learning on material relations derived from
-empirical data: the simulation is indeed solving physical balance laws.
+Multiphase simulation is hard.
 
@@
 
+We train an autoencoder to determine the lower-dimensional
+representation of the EOS surface using modern deep learning techniques
+without using any phase labels. The decoder end of the model is then
+plugged back into the balance laws, eliminating the constraint equation.
+\end{abstract}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\hypertarget{header-n3216}{%
+\section{Introduction}\label{header-n3216}}
+
+Representing complex physical systems with multiple speicies, phases,
+chemical reactions, and heterogenous scale effects poses an
+intractability problem to modeling. How do we analyze complex
+experimental data and \emph{ab initio} simulations and derive new
+macroscale relations to describe the systems? How do we then implement
+these relations in the rest of our analyses and production-scale
+computer simulations?
+
+Accurately simulating multispecies reaction and transport with
+undergoing phase transitions is a critical aspect to a variety of
+problems.
+This work is motivated by advancing simulation techniques for subsurface energy systems such as shale
+and oceanic natural gas reservoirs in shales or engineered geothermal
+systems.
+
+The system of equations we are looking at are, for only two components,
+\begin{align}
+  \partial_t \rho_{H_2O} =& \nabla \cdot F_{H_2O} + s \\
+  \partial_t \rho_{CO_2} =& \nabla \cdot F_{CO_2} + s \\
+  \partial_t e& = \nabla \cdot F_{heat} + s
+\end{align}
+More species add more balance laws (rows of equations). The applications also involve continuum
+mechanics and chemical reactions, which are not considered. The interplay of deformations,
+reactions, and phase transitions makes the equations extremely difficult
+to solve. \cite{queiruga2018millstone}
+(In this paper, we use use $\partial_t$, $\mathrm{d}/\mathrm{d}t$,
+and $\dot{()}$ to denote differentiation with respect to time
+interchangably depending on context.
+The complixity is hidden inside the relationships between temperature
+$T$, and pressure,$p$, density, $\rho$, and internal energy $e$ (or
+enthalpy, $h$, s.t. $e=\rho h - p$), 
+for the materials in question.
+
+% {\bf RESERVOIR PHYSICS LIT REVIEW HERE; Other applications;
+%   plasticity, }
+
+A single material with multiple phase transitions
+is a challenge in of itself, and the descriptions become combinatorially
+more complex as species and phases are added. This results in
+combinatorially more human labor and combinatorially more code which
+tends to involve complex branching logic. The descriptions then pose
+numerical challenges when incorporated into macroscale simulations due
+to obstacles such as discrete phase state machines and primary variable
+switching.
+This work only considers the "single beaker" problem; the flow
+and transport aspects have similar challenges. (The next stage of this
+work will "connect the beakers".)
+
+We redescribe our problem as a differential algebraic equation
+with a constraint: the balances of mass and energy are constrained by
+the EOS. We can cast the above equation into a basic form for a
+differential algebraic equation,
+\begin{align}
+  z( u, \partial_t u) &= 0 \\
+  c(u) & = 0
+\end{align}
+where $u=\{T,p,\rho,h\}$. We will introduce a new way of thinking about constraints as sets of
+solutions or observations of possible $u$:
+\begin{equation}
+c(u)=0 \leftrightarrow \left\{ u | c(u)=0 \;\text{or}\; u\; \text{observed} \right\}
+\end{equation}
+This maps directly to the original experimental procedure for
+determining a constraint equation.
+
+Because the EOS was originally derived from data, a data-driven
+approach lends itself in which the EOS is a set of points in
+temperature-pressure-density-enthalpy space. (In the multicomponent
+cases there are more dimensions.)
+We will use this set of possible values viewpoint to train an
+autoencoder on the set of $u$ for a latent space parameterized by $q$:
+\begin{equation}
+u = D(q), \; q=E(u) \;\forall u \in U
+\end{equation}
+which satisifies
+\begin{equation}
+  D(q)\in U \forall q.
+\end{equation}
+We can then write the original equation by
+\begin{equation}
+z(q,\partial_t q)=0
+\end{equation}
+The actual meaning of the parameterization is not important, so long as
+it can be effectively solved and then \emph{decoded} back to the
+physically meaningful quantities.
+
+% {\bf ML LIT REVIEW HERE} % TODO
+\cite{wu_physics-informed_2018}
+\cite{xie_tempogan:_2018}
+
+ A program synthesis approach was used to search for and evaluate possible methods of Peridynamics in Queiruga, 2017 \cite{queiruga2017numerical}.
+(An open source implementation of the study can be found Periflakes)\cite{queiruga2018periflakes}.
+
+Part of the new modern software engineering framework involves replacing
+human-written code with computer-written code. We are applying this
+mindset from industry problems in the Artificial
+Intelligence space into scientific computing.
+
 This work proposes a new methodology that steps away from the
@@
 the simulator.
+
+We demonstrate this approach to solving constrained differential
+equations on the pendulum, before demonstrating two EOSes of water: one
+with just liquid-gas phases, and one with liquid-gas-ice-supercritical
+regimes. The notion of phases does not appear in this multiphase
+simulation. We show a model architecture that is able to learn phase
+labels in the unsupervised autoencoder training process, such
+human-scientist interpretable results are obtained from the machine
+learning process. We compare this and other model architectures,
+including common deep neural networks, based on autoencoder error and
+simulation success metrics. We stress that the presented approach is
+only utilizing machine learning on material relations derived from
+empirical data: the simulation is indeed solving physical balance laws.
+
+We build a test dataset off of the IAPWS solutions. The proposed
+methodology applies to experimental data directly, or preexisting
+compendiums of empirical fits and theoretical relations.
+Note that, even if a relation can be theoretical derived, that does
+not mean that the analytical relationship yields the best program
+implementation: it is often the case that numerial approximations are
+superior to closed-form mathematical solutions in computational
+efficiency, accuracy, and numerical stability.
 
@@
 accompany this work. They can be found at the github repository
-located at: \href{https://github.com/afqueiruga/latentsim}
+located at: \href{https://github.com/afqueiruga/latentsim}.
 Certain components are factored into submodule repositories which are
@@
 
-\hypertarget{header-n3234}{%
-\section{Multiphase Equations of State}\label{header-n3234}}
+\section{Multiphase Equations of State}
+
+% \begin{figure}
+%    CSTR
+%   \caption{ \label{fig:cstr} Illustration of the Continuous stirred-tank
+%   reactor. The material in the reactor is exposed to a reservoir with pressure
+%   $p_\inf$ and temperature $T_\inf$ via a conduit with permeability
+%   and thermal conductivities $k_p$ and $k_T$, direct forced mass
+%   injection/withdrawl $r$, or heating and cooling $s$.}
+% \end{figure}
 
 For this preliminary work, we consider only a single-beaker problem
-without flow. Let \(\rho\) be the material density, \(T\) be the
+without flow, a continuous stirred-tank reactor (CSTR), as illustrated
+in Figure \ref{fig:cstr}. Let \(\rho\) be the material density, \(T\) be the
 temperature, \(p\) be the material pressure, \(h\) be the material
@@
 balance of mass and energy.
-
 \begin{align}
@@
 \end{align}
-
 The beaker is exposed to a reservoir with \(p_\infty\) and \(T_\infty\)
@@
   with internal energy.}
+We are neglecting upstreaming in the connection between the reservoir
+and the container.
+
+We also consider the phase equilibria: beyond the scale of resolving
+individual bubbles, it is possible
+for a control volume to be a mixture of two phase coexisting.
+(For a very fine scale simulation, e.g. at the pore scale, bubbles
+would be explicitly modeled.)
+We introduce a saturation $S_{phase}\in[0,1]$ for each phase;.
+The material can exist on this section of the surface as a mixture of
+the two phases. Thus, its density is calculated by a mixing the
+properties of all phases at that point in the phase diagram:
+\begin{equation}
+  \rho = S_{gas} \rho_{gas} + S_{liq} \rho_{liq} + S_{ice}
+    \rho_{ice} + S_{crit} \rho_{crit}
+  \end{equation}
+Each phase
+and equilibrium surface constrains the possible saturations. For example, on the boiling equilibrium surface there are two phases with some mixture
+saturations \(S_{gas}=1-S_{liq}\) and $S_{ice}=S_{crit}=0$. 
+For water, three phases can exist in equilibrium. Given more species, it is possible for an arbitrary number of phases
+to be in equilibrium; there is a quadruple point in methane-water
+hydrates.
+
+But, what is a phase boundary? The traditional explanation is shown in
+Figure \ref{fig:phaseboundaries}. The use of intensive properties breaks down:\footnote{https://en.wikipedia.org/wiki/Phase\_transition}
+
+\begin{quote}
+Phase transitions occur when the thermodynamic free energy of a system
+is non-analytic for some choice of thermodynamic variables (cf. phases).
+\end{quote}
+
+We can also assemble the surface of possible $T,p,\rho$ coordinates,
+shown in Figure \ref{fig:pTrho}.
+\begin{figure}
+\centering
+\includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/phase_diagram.png}
+\caption{\label{fig:phaseboundaries}Phase boundaries for water on the $T-p$ diagram. The standard
+interpretation is that the properties of the material change
+drastically when crossing this boundary. Note how it is possible to form a path from a
+definitively gas state to a definitively liquid state without crossing
+a phase boundary. The supercritical regime has no clear boundaries,
+but requires a seperate curve fit in practice due to numerical challenges.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/water_eos.png}
+\caption{\label{fig:pTrho} The $\log{p}$, $T$, $\rho$ equation of
+  state surface for water in units of $\log{\text{Pa}}$, K, and \si{}{kg/m$^2$} This plot spans four phases, three
+  equilibria, and the triple point, which are each given their own color.
+  There is a fourth dimension to the surface of enthalpy $h$ that exhibits
+  similar discontinuities; the $\log{p}$, $T$, $h$ surface is plotted
+  in the Appendix. (Every point on this plot is a point in the
+  generated dataset used for the optimization)}
+\end{figure}
+
+{\bf Remark:}There are even more known phases of crystal ice which for water
+For water, the supercritical phase and exotic ices only occur in
+laboratory conditions or naturally near the
+crust \cite{}. For our case, we only handle the Ice I and cutoff the
+regime at \si{10^8}{Pa}. For other materials these exotic phases can be
+applicable. For example, in geoengineered systems using $CO_2$
+such as geothermal plants are carbon sequestration facilities, the
+working fluid is usually in the supercritical regime at depth. cite
+curt. cite hydrates.
+When binary mixtures are involved, the number of possible phases and
+equilibria increases again.
+
+{\bf Remark:} This interpretation breaks down with the existenceof super-saturated phases in non-stable regions: e.g., it is possible to
+heat water beyond it's boiling point in a microwave. We are concerned
+with environments where this would not happen (the very rough and
+heterogenous) pores of rocks, so we have not thought of how to
+integrate this problem.
+
+\hypertarget{header-n3253}{%
+\subsection{State Machine logic}\label{header-n3253}}
 
@@
 the DAE for \(p\) and \(T\) implicitly:
-
 \begin{align}
@@
 \end{align}
-
 The complicatation is that the functions \(\rho(p,T)\) and \(u(p,T)\)
@@
 
-We also consider the phase equilibria; at a larger scale it is possible
-for a control volume to be a mixture of two phase coexisting, such as
-boiling with some mixture saturation \(S_{gas}=1-S_{liq}\).
-(For a very fine scale simulation, bubbles would be explicitly modeled.)
-The material can exist on this section of the surface as a mixture of
-the two phases.
-
-\[\rho = X_{H_2O}\left(S_{gas} \rho_{gas} + S_{liq} \rho_{liq} + S_{ice} \rho_{ice} + S_{crit} \rho_{crit} \right)\]
-
-One possible definition of phase boundaries:
-
-\begin{quote}
-Phase transitions occur when the thermodynamic free energy of a system
-is non-analytic for some choice of thermodynamic variables (cf. phases).
-
-https://en.wikipedia.org/wiki/Phase\_transition
-\end{quote}
-
+\begin{tabular}{l|c|c}
+  State & $X_1$ & $X_2$ \\
+  \hline
+  Ice & $p$ & $T$ \\
+  Liquid & $p$ & $T$ \\
+  Gas & $p$ & $T$ \\
+  Supercritical & $\rho$ & $T$ \\
+  Ice-Liquid & $S_{ice}$ & $T$ \\
+  Ice-Gas & $S_{ice}$ & $T$ \\
+  Gas-Liquid & $S_{gas}$ & $T$ \\
+  Ice-Gas-Liquid & $S_{ice}$ & $S_{gas}$ 
+\end{tabular}
+
+  This requires a statemachine. For the complete states of water, it is shown in Figure \ref{fig:statemachine}.
   \begin{figure}
-\centering
-\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/phase_diagram.png}
-\caption{Phase boundaries for water }
+    \includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/phase_4.png}
+    \caption{\label{fig:statemachine}States and possible transitions for water.}
 \end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/water_eos.png}
-\caption{\label{fig:pTrho} The $\log{p}$, $T$, $\rho$ equation of
-  state surface for water. This plot spans four phases, three
-  equilibria, and the triple point, which are each given their own color.
-  There is a fourth dimension to the surface of enthalpy $h$ that exhibits
-  similar discontinuities; the $\log{p}$, $T$, $h$ surface is plotted
-  in the Appendix.}
-\end{figure}
-
-There are even more known phases of crystal ice which for water
-For water, the supercritical phase and exotic ices only occur in
-laboratory conditions or naturally near the
-crust \cite{}. For other materials these exotic phases can be
-applicable. For example, in geoengineered systems using $CO_2$
-such as geothermal plants are carbon sequestration facilities, the
-working fluid is usually in the supercritical regime at depth.
-
-\hypertarget{header-n3253}{%
-\subsection{State Machine logic}\label{header-n3253}}
-
 Formulate balance equations as
-
-\[R(X ; phase) = 0\]
-
+\begin{equation}
+R(X ; phase) = 0
+\end{equation}
 Solve the differentiable part:
@@
 constraint:
-
 \begin{align}
@@
 \end{align}
-
 The \emph{ground truth} is the experimental data in the first place; the
@@
 
-An archetypical constraint problem the pendulum. Posed as dynamics
+An archetypical constraint problem is the pendulum. Posed as dynamics
 problem with a Lagrange-multipler-esque centripetal force, it
@@
 \end{align}
-
 Note that we are considering the nonlinear pendulum.
@@
 \[\left\{\begin{array}{c} x\\y\end{array}\right\}
-\rightarrow W_{enc} x^n \rightarrow q \rightarrow W_{dec} x^n \rightarrow 
+\rightarrow W_{enc} x^n +b_{enc}\rightarrow q \rightarrow W_{dec} x^n +b_{dec}\rightarrow 
 \left\{\begin{array}{c} x\\y\end{array}\right\}\]
-
+where $x^n$ denotes a polynomial basis set, e.g. for two variables $x^n=\{x,y,x^2,xy,y^2,..y^n\}$.
 (We explicitly write the bias \(+b\); it is equivalent to including
 \(x^0=1\) in the polynomial set.)
-The equations
-of motion given some parameter space that satisfies the constraint is
-
-\begin{equation}
-\frac{\mathrm{d}q}{\mathrm{d}t}\frac{\partial L}{\partial \dot{q}} = \frac{\partial L}{\partial q}.
-\end{equation}
 
 We'll use the autoencoder to generate the parameterization of \(x(q)\)
-and \(y(q)\). By the chain rule, the velocities are related to the
-feature space by
-\begin{equation}
-\dot{x} = \frac{\mathrm{d}x(q)}{\mathrm{d}t} = \frac{\partial x}{\partial q}\frac{\mathrm{d}q}{\mathrm{d}t}
-\end{equation}
-and similarly for the \(y\) direction.
+and \(y(q)\). To do this, we optimize the loss function,
+\begin{equation}
+L(x) = \left\| x-D(E(x)) \right\|
+\end{equation}
 
@@
 its rate, the components to the equation of motion are build the DAE we know from physics:
-
-\[\frac{\mathrm{d}}{\mathrm{d}t}\frac{\partial L}{\partial \dot{q}} = \frac{\partial L}{\partial q}\]
-where we get the velocity with the chain rule:
-\[v = \frac{\partial x}{\partial q} \dot{q}\]
-
+\[\frac{\mathrm{d}}{\mathrm{d}t}\frac{\partial L}{\partial \dot{q}} = \frac{\partial L}{\partial q}\].
+By the chain rule, the velocities are related to the
+feature space by
+\begin{equation}
+\dot{x} = \frac{\mathrm{d}x(q)}{\mathrm{d}t} = \frac{\partial x}{\partial q}\frac{\mathrm{d}q}{\mathrm{d}t}
+\end{equation}
+and similarly for the \(y\) direction.
 Then plug it into a Runge Kutta (trapezoidal) to solve with Newton's
@@
 
+Soltutions are shown in Figure \ref{fig:pendulum}.
 \begin{figure}
 \centering
-\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/pendulum_q.png}
-\caption{}
+\includegraphics[width=5in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/pendulum_q.png}
+\caption{\label{fig:pendulum}Solution of the pendulum on an arbitrary latent space.}
 \end{figure}
@@
 
+\section*{Acknowlegements}
+
+This work was supported by Department of Energy, Office of Fossil
+Energy contracts
+1. Hydrates
+2. Shales
+
+The author would like to thank his colleagues Matt Reagan and George
+Moridis for all their instruction on chemsitry. 
 \bibliographystyle{plainnat}

### 2019-06-10 autoencoder.py
Added ability to reset optimizer; realized that after_run fuses with the sess.run somehow?

@@ -74,12 +74,23 @@ class Autoencoder(object):
             return loss
 
     def _make_train_step(self, data):
-        opt = tf.train.AdamOptimizer(1e-2)
+#         opt = tf.train.AdamOptimizer(1e-2)
+#         opt = tf.train.GradientDescentOptimizer(1e-3)
+        opt = tf.train.RMSPropOptimizer(1e-2)
         loss = self.make_goal(data)
-        ts = opt.minimize(loss,global_step=tf.train.get_or_create_global_step())
+        var_list = tf.get_collection(tf.GraphKeys.TRAINABLE_VARIABLES)
+        ts = opt.minimize(loss,global_step=tf.train.get_or_create_global_step(),var_list=var_list)
         self.sgd_opt = opt
         self.sgd_step = ts
-        self.sgd_reset = tf.variables_initializer(opt.variables())
+        opt_vars = [opt.get_slot(var, name) for name in opt.get_slot_names() for var in var_list]
+        try:
+            opt_vars += opt._get_beta_accumulators() # only for adam
+        except AttributeError:
+            print("Wasn't ADAM")
+            pass
+        opt_vars = [ v for v in opt_vars if isinstance(v, tf.Variable) ]
+        print(opt_vars)
+        self.sgd_reset = tf.variables_initializer(opt_vars)
         return ts
     
     def _get_hess_vars(self):

### 2019-06-10 autoencoder_trainer.py

@@ -54,7 +54,7 @@ class DoStuffHook(tf.train.SessionRunHook):
         return self
     
 def train_autoencoder(name, dataname, outerdim, innerdim, 
-                      hyper=default_hyper,
+                          hyper=default_hyper,
                       training_dir='',data_dir='',n_epoch=5000, image_freq=1500):
     
     hyperpath = string_identifier(hyper)
@@ -113,26 +113,40 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
             ae.save_fit(training_dir+"/surf_{0}.csv".format(onum_val),
                         header,sess=ctx.session)
         # Do hessian steps here and there
+        def newtstep(sess):
+            print("Running the newton step")
+            for i in range(1):
+                print("newt:",sess.run(ae.goal_all) )
+                for v in ae._get_hess_vars():
+                    print(sess.run(v))
+                sess.run(ae.newt_step)
+            print("newt:",sess.run([ae.goal_all]) )
+            for v in ae._get_hess_vars():
+                print(sess.run(v))
+            sess.run(ae.sgd_reset)
         @DoStuffHook(freq=n_epoch/2)
         def newthook(ctx,run_values):
             # Now do the Hessian step on the last layer
-            print("Running the newton step")
-            for i in range(1):
-                print("newt:",ctx.session.run(ae.goal_all))
-                ctx.session.run(ae.newt_step)
-            print("newt:",ctx.session.run(ae.goal_all))
-            ctx.session.run(ae.sgd_reset)
+            newtstep(ctx.session)
 
         # set up the session
+        # replace with MonitoredSession with manually crafted
+        # checkpointhook
         session = tf.train.MonitoredTrainingSession(
             checkpoint_dir=training_dir,
-            hooks=[loghook,printgoalhook,saverhook,extrahook,
-                   newthook,
+            hooks=[loghook,saverhook,extrahook,
+                   #newthook,
                    stophook])
         # train away
         with session as sess:
             # Do the SGD rounds
+            i=0
             while not sess.should_stop():
-                print("loop:",sess.run([ae.goal_all,ae.train_step]))
+                if i%(n_epoch/2)==n_epoch/2-1:
+                    newtstep(sess)
+                else:
+                    print("loop:",sess.run([ae.goal_all,ae.train_step]
+                                       +list(ae._get_hess_vars())) )
+                i+=1
 
 

### 2019-06-10 paper/paper.tex
edited the paper

@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
+%\setcounter{secnumdepth}{0}
 % Redefines (sub)paragraphs to behave more like sections
@@
   Multiphase simulation is hard.
+
+Our
+proposed method is motivated by the problems of multiphase reaction and
+transport, where the underlying constraints are the ill-behaved material
+properties which are difficult to describe.
 
@@
   \partial_t \rho_{CO_2} =& \nabla \cdot F_{CO_2} + s \\
-  \partial_t e& = \nabla \cdot F_{heat} + s
+  \partial_t e =& \nabla \cdot F_{heat} + s
 \end{align}
@@
 \end{align}
-where $u=\{T,p,\rho,h\}$. We will introduce a new way of thinking about constraints as sets of
+where, for the equation above, the variable is $u=\{T,p,\rho,h\}$. We will introduce a new way of thinking about constraints as sets of
 solutions or observations of possible $u$:
 \begin{equation}
-c(u)=0 \leftrightarrow \left\{ u | c(u)=0 \;\text{or}\; u\; \text{observed} \right\}
+c(u)=0 \longleftrightarrow \left\{ u | c(u)=0 \;\text{or}\; u\; \text{observed} \right\}
 \end{equation}
@@
 The central idea of this work is to obtain two new variables solve for
-those directly,
-
-\begin{align}
-\text{Solve for}\, u(t)\, \text{and}\, v(t)\, \text{such that:}\\
-\partial_t \rho(u,v) & = \nabla \cdot \mathbf{k} \nabla p(u,v) + r \\
-\partial_t \rho e(u,v) & = \nabla \cdot \mathbf{k'}\nabla T(u,v) + s
-\end{align}
-
-where \(u\) and \(v\) have no physical meaning other than simply being a
-parameterization of the equation of state constraint. We want to select
-\(u\) and \(v\) to require no auxilary phase index to define the system
+those directly which have no physical meaning other than simply being a
+parameterization of the equation of state constraint. We want to select a
+\(q_1\) and \(q_2\) to require no auxilary phase index to define the system
 and no additional logic in the code. The sharp kinks and possible
 discontinuities can be a part of the functions
-\(\rho(u,v), p(u,v), e(u,v)\) and \(T(u,v)\).
-
-Can we automatically discover this parameterization? (For a such a
-simple equation, we do not think the proposed methodology will be
-better, but the pendulum will instead be our \emph{unit test}.) Our
-proposed method is motivated by the problems of multiphase reaction and
-transport, where the underlying constraints are the ill-behaved material
-properties which are difficult to describe. We change our perspective on
-these problems in the next section.
+\(\rho(q_1, q_2), p(q_1, q_2), h(q_1, q_2)\) and \(T(q_1, q_2)\).
 
@@
 forms an identity function with a compressed subspace:
-
-\begin{equation}
-\left\{ \rho, p, e, T \right\} \rightarrow  E \rightarrow \left\{ u,v \right\} \rightarrow D \rightarrow \left\{ \rho, p, e, T \right\}
-\end{equation}
-
+\begin{equation}
+\left\{ \begin{array}{c}
+T\\ p\\ \rho\\ h
+\end{array}\right\} \rightarrow  E \rightarrow 
+\left\{ \begin{array}{c} q_1\\q_2 \end{array} \right\}\rightarrow D \rightarrow 
+\left\{ \begin{array}{c}
+T\\ p\\ \rho\\ h
+\end{array}\right\}
+\end{equation}
 The autoencoder is solved for by optimizing its parameters using the
@@
 To form the hessian and gradient for this step, we sum over all of the
-training data. (The contractive penalty term drops out because it does
+training data. (The contractive penalty term drops out of the this step because it does
 not depend on \(W^{dec}\).)
@@
 \(\rho,h\).) Ideally, this system would be fully automated, but due to
-the nonlinearilty of the autoencoder problem there are many "bad"
-options the trainer can fall into. Multiple starts are possibly needed
+the nonlinearilty of the autoencoder problem there are many poor local minima.
+Multiple starts are possibly needed
 to get a good solution when starting from purely randomized newtork
@@
 
+The architecture, illustrated in Figure \ref{fig:classifyingnetwork}, has the form:
+\begin{align}
+  q &= W_{enc} x^n + b_{enc}\\
+  f &= W_{dec} q^m + b_{dec} \\
+  h &= \sigma\left(W_{bound}q^m+b_{bound}\right) \\
+  p &= \text{softmax}\left(W_{phase}q^m+b_{bound[phase}\right) \\
+  \bar{x}&= f * p
+\end{align}
+
 \begin{figure}
@@
 \includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_network.png}
-\caption{The architecture of the classifier network designed for
+\caption{\label{fig:classifyingnetwork}The architecture of the classifier network designed for
   unsupervised labeling of phases.}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_good.jpg}
+  \caption{\label{fig:classified}Result of the unsupervised phase classifier. Compare this plot to the original labeled dataset Figure \ref{fig:pTrho}.}
+\end{figure}
+Because the identity of the phases,
+For comparison purposes, we group together liquid-gas-supercritical as one contiguous label.
+
 \hypertarget{header-n3321}{%
@@
 With the trained model, we solve for \(q_1(t)\) and \(q_2(t)\) such that:
-
 \begin{align}
@@
 \end{align}
-
 where \(\rho(q_1,q_2)\) etc. are the back of an autoencoder:
-
-\begin{equation}
-\left\{ \begin{array}{c}
-T\\ p\\ \rho\\ h
-\end{array}\right\} \rightarrow  E \rightarrow 
-\left\{ \begin{array}{c} q_1\\q_2 \end{array} \right\}\rightarrow D \rightarrow 
-\left\{ \begin{array}{c}
-T\\ p\\ \rho\\ h
-\end{array}\right\}
-\end{equation}
 
@@
 First, we had a constrained DAE:
-
 \begin{align}
@@
 \end{align}
-
-with no good representation of \(eos\) everywhere.
+where $m=\{\rho,\rho h -p\}$ and $r$ is the right hand side, with no good representation of \(eos\) everywhere.
+
+We obtain four new functions with two unknowns. We're actually solving the manifold of Figure X.
+\begin{figure}
+  \centering
+  \includegraphics{../slides/water_four_surfaces_plot.png}
+  \caption{\label{fig:foursurfaces}The four 2D surfaces that make up the new parameterization.
+The learned surface captures the sharp kinks and vertical fronts in Figure \ref{fig:classifyingresult}, but is ananalytic and smooth function of the two learned unknowns.}
+  \end{figure}
 
@@
 this paper is not exploring new time stepping schemes. This yields the
-discretizated algebraic equation for \(q_i\) at \(t+\Delta t\) given the
+discretizated algebraic equation for \(q_i\) and \(t+\Delta t\) given the
 value \(q^0\) at \(t\),
@@
 \end{equation}
-
-where we have called the quation \(R(q)\). By design \(R(q)\) is smooth
+where we have called the equation \(R(q)\). By design \(R(q)\) is smooth
 enough to solve with Newton's method without any statemachine logic:
-
-\begin{equation}
-\frac{\partial R}{\partial q} \Delta q^{k+1} = R(q^k)
-\end{equation}
+\begin{equation}
+\frac{\partial R}{\partial q} \Delta q^{k+1} = - R(q^k)
+\end{equation}
+with the update
+\begin{equation}
+q^{k+1} = q^k + \alpha\Delta q^{k+1}
+\end{equation}
+where $\alpha$ is an under relaxation parameter.
+
 This equation is deceptively simple: the function \(R(q)\) has complex logic encoded in it; if-then
@@
 the vast majority of programming was implemented automatically by the
-models; implementing equations XX, XX, and XX require only a few lines
+models; implementing equations XX, XX, and XX requires only a few lines
 of human-written code.
@@
 initial values. In a multiphase solver such as TOUGH, the state needs
-to be specified with the two primary variables plus the state tag: 
+to be specified with the two primary variables plus the state tag (sans input file formatting): 
 \begin{verbatim}
@@
 and the following system of equations is solved for an update to $q$,
-\[K\Delta q = s^0_i-D(q)_i\]
-Updating $q$ is repeated until convergence.
-
+\[K_{ij}\Delta q_j = s^0_i-D(q)_i\]
+The iteration u $q$ is repeated until convergence, with under-relaxation used again to update
+ \begin{equation}
+q^0_{k+1} =q^0_{k} + \alpha \Delta q_{k}
+\end{equation}
+with the under-relaxation parameter usually set to $\alpha \approx
+0.1$.
 The initial condition problem is not gaurunteed to converge, however,
@@
  These sets of $q^0_{guess}$ are then iterated upon using the Newton's
- method procedure. We further add
- under-relaxation to the iteration,
- \begin{equation}
-q^0_{k+1} =q^0_{k} + \alpha \Delta q_{k}
-\end{equation}
-with the under relaxation parameter usually set to $\alpha \approx
-0.1$.
+ method procedure.
 
@@
 factored out at
-\href{https://github.com/afqueiruga/equations_of_state}{https://github.com/afqueiruga/equations\emph{of}state}.
+\href{https://github.com/afqueiruga/equations_of_state}{https://github.com/afqueiruga/equations\_of\_state}.
 
@@
 \hypertarget{header-n3418}{%
-\subsection{Benchmarks}\label{header-n3418}}
-
+\subsection{Validation and Testing}\label{header-n3418}}
+
+The verification problems are listed in Table \ref{tab:problems}.
+\begin{table}
+  \caption{\label{tag:problems}Enumeration of the problems used to test and evaluate each fo the generaed architectures}
+\begin{tabular}{l|c|c|c}
+  Name & start & path & end\\
+\hline
+\end{tabular}
+\end{table}
 The simulator was benchmarke against the long standing TOUGH+ multiphase
@@
 
+We can also think about this work as offline preconditioning.
+
+In TOUGH, for one instance, all of the calculations are performmed in SI units because it is simply too much labor to normalize everything!
+Adverse effects to iterative algorithms and floating point error propogation occur, especially because numerical differentiation is used.
+The learned primary variables are normalized to a unit range around zero, and the scaling matrix is explicitly in one step.
+The scaling matrix could be a way to systematically nondimensionalize the entire problem.
+
 The larger simulation code links into the trained database to load the
@@
 
-Technique ssuch as L1 regularization would be needed to search for low
+Techniquessuch as L1 regularization would be needed to search for low
 cost models that also satisfy our desire for Occam's razor

### 2019-06-11 autoencoder.py
managed to unfurl it all

@@ -75,8 +75,8 @@ class Autoencoder(object):
 
     def _make_train_step(self, data):
 #         opt = tf.train.AdamOptimizer(1e-2)
-#         opt = tf.train.GradientDescentOptimizer(1e-3)
-        opt = tf.train.RMSPropOptimizer(1e-2)
+        opt = tf.train.GradientDescentOptimizer(1e-3)
+#         opt = tf.train.RMSPropOptimizer(1e-2)
         loss = self.make_goal(data)
         var_list = tf.get_collection(tf.GraphKeys.TRAINABLE_VARIABLES)
         ts = opt.minimize(loss,global_step=tf.train.get_or_create_global_step(),var_list=var_list)
@@ -100,9 +100,10 @@ class Autoencoder(object):
         
     def _make_hess_train_step(self,data):
         loss = self.make_goal(data)
-        newt = atu.NewtonsMethod(loss, self._get_hess_vars())
-        self.newt_step = newt
-        return newt
+        ops = atu.NewtonsMethod_pieces(loss, self._get_hess_vars())
+        self.newt_ops = ops
+        self.newt_step = ops[0]
+        return self.newt_step
     
     def eval_q(self, i_x):
         return sess.eval( self.o_q, feed_dict={self.i_x:i_x} )
@@ -286,7 +287,7 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         h_select = tf.tensordot(h_bound,W_select, axes=[-1,0],
                                 name=(name if not self.softmax_it else None))
         if self.softmax_it:
-            h_select = tf.nn.softmax(h_select, name=name)
+            h_select = tf.nn.softmax(h_select)
         return h_select
     
     def decode(self, q, name=None):
@@ -310,4 +311,4 @@ class ClassifyingPolyAutoencoder(Autoencoder):
     def _get_hess_vars(self):
         """These are variables used for the final fitting
         phase."""
-        return (self.vars["dec_W_curve"], self.vars["dec_b_curve"])
+        return (self.vars["dec_W_curve"],)# self.vars["dec_b_curve"])

### 2019-06-11 autoencoder_trainer.py

@@ -98,7 +98,7 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
         # Print to screen hook
         @DoStuffHook(freq=1)
         def printgoalhook(ctx,run_values):
-            print(ctx.session.run(ae.goal_all))
+            print(ctx.sessiormn.run(ae.goal_all))
         stophook = tf.train.StopAtStepHook(num_steps=n_epoch)
         saverhook = SaveAtEndHook(training_dir+"/final_variables")
         # Make a closure into a hook
@@ -119,6 +119,7 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
                 print("newt:",sess.run(ae.goal_all) )
                 for v in ae._get_hess_vars():
                     print(sess.run(v))
+                from IPython import embed ; embed()
                 sess.run(ae.newt_step)
             print("newt:",sess.run([ae.goal_all]) )
             for v in ae._get_hess_vars():

### 2019-06-11 training_water_lg.py

@@ -1,21 +1,29 @@
 if __name__=="__main__":
     training_dir = "/Users/afq/Google Drive/networks/"
-    data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
-#     data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
+#    data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
+    data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
     dataset = "water_lg_sharded/*.csv"
-#     inis = [ 'rand', 'pT', 'rhoh' ]
 #     caes = [ 0.0, 1.0 ]
     inis = [ 'rand', ]# 'pT', 'rhoh' ]
-    caes = [ 0.0 ]
+    caes = [ 0.0, ]
     extend_train_params = lambda arch : \
         [ dict(ini=ini,cae=cae,**arch) for ini in inis for cae in caes ]
     archs_to_try = [
-        {'type':'Classifying','args':[1,3, 3,3,'sigmoid']},
-#         {'type':'Classifying','args':[1,3, 6,12,'sigmoid']},
+#          {'type':'Classifying','args':[1,1, 3,3,'sigmoid']},
+#          {'type':'Classifying','args':[1,1, 6,12,'sigmoid']},
+        #{'type':'Poly','args':[1,2]},
+        #{'type':'Poly','args':[1,3]},
+        #{'type':'Poly','args':[1,5]},
+        {'type':'Classifying','args':[1,1, 3,2,'sigmoid']},
+        
+        
+        # {'type':'Classifying','args':[1,3, 6,12,'sigmoid']},
+#         {'type':'Classifying','args':[1,4, 6,12,'sigmoid']},
+        #         {'type':'Classifying','args':[1,4, 3,3,'sigmoid']},
+#         {'type':'Classifying','args':[1,4, 6,12,'sigmoid']},
 #         {'type':'Classifying','args':[1,1, 6,12,'tanh']},
 #         {'type':'Classifying','args':[1,1, 6,12,'sigmoid']},
-#         {'type':'Poly','args':[1,4]},
-#         {'type':'Poly','args':[1,5]},
+
     ]
     sets_to_try = []
     for arch in archs_to_try:
@@ -27,12 +35,15 @@ if __name__=="__main__":
     import itertools
     def job(S):
         # Import of tensorflow has to happen *after* forking
[…]

### 2019-06-12 paper/paper.tex
Added a note

@@
 
+For this step in which only the back-end regression is performmed, each set of coefficients for the curves for each variable is independent from those of the other variables: e.g., the coefficients to the sets of polynomials that parameterize $p(q)$ do not depend on those for $T(q)$.
+This means that this step can be performmed in blocks, performing the linear step for each component independently, at a great reduction of cost.
+For the full system, 3/4 of the entries are zero.
+We did not bother performing this optimization, but note that the savings are quadratic as the number of coefficients grow for more complex equations of state.
+(This is not true for training the entire deep network: it is only true for this linear regression step on the sub-network.)
+
 In future work, we would want to continue appending new \(q\)s in a

### 2019-06-12 autoencoder.py
Really deep into debugging the newton step

@@ -53,13 +53,13 @@ class Autoencoder(object):
         W = self._var( "enc_W", (self.size_q, self.size_x),
                      initial_value=encoder_init_options[self.encoder_init])
         b = self._var( "enc_b", (self.size_q,) )
-        return tf.add(tf.matmul(W,x),b,name=name)
+        return tf.add(tf.matmul(W,x),b)
     
     def decode(self, q, name=None):
         W = self._var( "dec_W", (self.size_x, self.size_q) )
         b = self._var( "dec_b", (self.size_x,) )
         x = tf.matmul(W,q)+b
-        return tf.identity(x,name=name)
+        return tf.identity(x,)
     
     def make_goal(self, data):
         q = self.encode(data)
@@ -112,7 +112,9 @@ class Autoencoder(object):
              initial_value=None):
         try:
             v = self.vars[name]
+            print("Fetching ",name)
         except KeyError:
+            print("Making ",name)
             if not initial_value is None:
                 ini = tf.constant(initial_value,dtype=self.dtype) + tf.truncated_normal(shape=shape,
                        stddev=stddev, dtype=self.dtype)
@@ -188,14 +190,14 @@ class PolyAutoencoder(Autoencoder):
                         initial_value = encoder_init_options[self.encoder_init])
         be1 = self._var("enc_b", (self.size_q,) )
         q = tf.matmul( atu.polyexpand(x, self.Np_enc), We1 ) + be1
-        return tf.identity(q,name=name)
+        return tf.identity(q)
     
     def decode(self, q, name=None):
         N_coeff = atu.Npolyexpand( self.size_q, self.Np_dec )
         We1 = self._var("dec_W", (N_coeff, self.size_x) )
         be1 = self._var("dec_b", (self.size_x,) )
         x = tf.matmul( atu.polyexpand(q, self.Np_dec), We1 ) + be1
-        return tf.identity(x,name=name)
+        return tf.identity(x)
     
     
 class DeepPolyAutoencoder(Autoencoder):
@@ -225,13 +227,13 @@ class DeepPolyAutoencoder(Autoencoder):
         N_coeff = atu.Npolyexpand( self.size_x, self.Np_enc )
         nxt = atu.polyexpand(x, self.Np_enc)
         nxt = self._layers(nxt, self.enc_layers + [self.size_q], prefix="enc")
-        return tf.identity(nxt,name=name)
[…]

### 2019-06-12 autoencoder_trainer.py

@@ -119,6 +119,13 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
                 print("newt:",sess.run(ae.goal_all) )
                 for v in ae._get_hess_vars():
                     print(sess.run(v))
+                x=ae.data.eval(session=sess)
+                W=ae.vars['dec_W_curve'].eval(session=sess)
+                b=ae.vars["dec_b_curve"].eval(session=sess)
+                q = ae.o_q.eval(session=sess,feed_dict={ae.i_x:x[0:1,:]})
+                p_hand = ae.o_class.eval(session=sess,feed_dict={ae.i_q:q})
+                f_hand = np.einsum('jik,Qj->Qik',W,q)+b
+                x_hand = np.einsum('Qik,Qi->Qk',f_hand,p_hand)
                 from IPython import embed ; embed()
                 sess.run(ae.newt_step)
             print("newt:",sess.run([ae.goal_all]) )

### 2019-06-12 autoencoder.py
This training step now hard-selects the classes and robustly does the regression

@@ -276,7 +276,7 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         b = self._var("enc_b", (self.size_q,) )
         return tf.add(tf.matmul( atu.polyexpand(x, self.Np_enc), W ), b)
     
-    def classify(self, q, name=None):
+    def classify(self, q, name=None, phase_act="softmax"):
         qpoly = atu.polyexpand(q, self.Np_dec)
         N_coeff = atu.Npolyexpand( self.size_q, self.Np_dec )
 
@@ -287,13 +287,15 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         
         W_select = self._var("dec_W_select", (self.N_bound,self.N_curve))
         h_select = tf.tensordot(h_bound,W_select, axes=[-1,0])
-        if self.softmax_it:
+        if phase_act=="softmax":
             h_select = tf.nn.softmax(h_select)
         else:
-            h_select = tf.onehot(tf.argmax(h_select))
+            h_select = tf.one_hot(tf.argmax(h_select,axis=-1),depth=h_select.shape[-1],
+                                  dtype=h_select.dtype)
+            print("Did this.")
         return h_select
     
-    def decode(self, q, name=None):
+    def decode(self, q, name=None, phase_act="softmax"):
         qpoly = atu.polyexpand(q, self.Np_dec)
         N_coeff = atu.Npolyexpand( self.size_q, self.Np_dec )
         
@@ -301,7 +303,7 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         b1 = self._var("dec_b_curve", (self.N_curve, self.size_x) )
         h_curve = tf.tensordot(qpoly,W1,axes=[-1,0])+b1
         
-        h_select = self.classify(q)
+        h_select = self.classify(q,phase_act=phase_act)
         
         x = tf.einsum('ijk,ij->ik',h_curve,h_select)
         return x #tf.identity(x)
@@ -311,7 +313,37 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         classes = probs.argmax(axis=-1)
         return ",classes",classes
 
+    def make_goal(self, data, phase_act="softmax"):
+        q = self.encode(data)
+        pred = self.decode(q,phase_act=phase_act)
+        loss = tf.losses.mean_squared_error(data, pred)
+        if self.cae_lambda != 0:
+            dqdx = atu.vector_gradient_dep(q,data)
[…]

### 2019-06-12 autoencoder.py
Updated the training; put the names back in

@@ -74,9 +74,9 @@ class Autoencoder(object):
             return loss
 
     def _make_train_step(self, data):
-#         opt = tf.train.AdamOptimizer(1e-2)
-        opt = tf.train.GradientDescentOptimizer(1e-3)
-#         opt = tf.train.RMSPropOptimizer(1e-2)
+        opt = tf.train.AdamOptimizer(1e-2)
+        #opt = tf.train.GradientDescentOptimizer(1e-3)
+        #opt = tf.train.RMSPropOptimizer(1e-2)
         loss = self.make_goal(data)
         var_list = tf.get_collection(tf.GraphKeys.TRAINABLE_VARIABLES)
         ts = opt.minimize(loss,global_step=tf.train.get_or_create_global_step(),var_list=var_list)
@@ -105,6 +105,17 @@ class Autoencoder(object):
         self.newt_step = ops[0]
         return self.newt_step
     
+    def _do_hess_train_step(self,session):
+        G = self.newt_ops[1].eval(session=session)
+        H = self.newt_ops[2].eval(session=session)
+        idxs=np.where(~H.any(axis=1))[0]
+        for i in idxs: 
+            H[i,i]=1
+        DeltaW = np.linalg.solve(H,-G)
+        i_delta_W = self.newt_ops[-2]
+        delta_assign_op = self.newt_ops[-1]
+        session.run(delta_assign_op,feed_dict={i_delta_W:DeltaW.reshape(-1,1)})
+        
     def eval_q(self, i_x):
         return sess.eval( self.o_q, feed_dict={self.i_x:i_x} )
 
@@ -190,14 +201,14 @@ class PolyAutoencoder(Autoencoder):
                         initial_value = encoder_init_options[self.encoder_init])
         be1 = self._var("enc_b", (self.size_q,) )
         q = tf.matmul( atu.polyexpand(x, self.Np_enc), We1 ) + be1
-        return tf.identity(q)
+        return tf.identity(q,name=name)
     
     def decode(self, q, name=None):
         N_coeff = atu.Npolyexpand( self.size_q, self.Np_dec )
         We1 = self._var("dec_W", (N_coeff, self.size_x) )
         be1 = self._var("dec_b", (self.size_x,) )
         x = tf.matmul( atu.polyexpand(q, self.Np_dec), We1 ) + be1
-        return tf.identity(x)
+        return tf.identity(x,name=name)
     
     
[…]

### 2019-06-12 autoencoder_trainer.py

@@ -119,15 +119,16 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
                 print("newt:",sess.run(ae.goal_all) )
                 for v in ae._get_hess_vars():
                     print(sess.run(v))
-                x=ae.data.eval(session=sess)
-                W=ae.vars['dec_W_curve'].eval(session=sess)
-                b=ae.vars["dec_b_curve"].eval(session=sess)
-                q = ae.o_q.eval(session=sess,feed_dict={ae.i_x:x[0:1,:]})
-                p_hand = ae.o_class.eval(session=sess,feed_dict={ae.i_q:q})
-                f_hand = np.einsum('jik,Qj->Qik',W,q)+b
-                x_hand = np.einsum('Qik,Qi->Qk',f_hand,p_hand)
-                from IPython import embed ; embed()
-                sess.run(ae.newt_step)
+                # x=ae.data.eval(session=sess)
+                # W=ae.vars['dec_W_curve'].eval(session=sess)
+                # b=ae.vars["dec_b_curve"].eval(session=sess)
+                # q = ae.o_q.eval(session=sess,feed_dict={ae.i_x:x[0:1,:]})
+                # p_hand = ae.o_class.eval(session=sess,feed_dict={ae.i_q:q})
+                # f_hand = np.einsum('jik,Qj->Qik',W,q)+b
+                # x_hand = np.einsum('Qik,Qi->Qk',f_hand,p_hand)
+                # from IPython import embed ; embed()
+                # sess.run(ae.newt_step)
+                ae._do_hess_train_step(sess)
             print("newt:",sess.run([ae.goal_all]) )
             for v in ae._get_hess_vars():
                 print(sess.run(v))

### 2019-06-15 autoencoder.py
Added a small ramp up of softmax beta ; not sure what it does yet

@@ -277,6 +277,8 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         self.softmax_it = softmax_it
         self.softmax_beta = tf.Variable(1.0, dtype=data.dtype,
                                         trainable=False)
+        self.i_beta = tf.placeholder(name='beta',shape=None,dtype=data.dtype)
+        self.update_beta_op = tf.assign(self.softmax_beta,self.i_beta)
         Autoencoder.__init__(self,size_x, size_q, data, data_all,
                             encoder_init=encoder_init,
                             cae_lambda=cae_lambda)
@@ -300,7 +302,8 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         h_bound = act(tf.tensordot(qpoly,W2,axes=[-1,0])+b2)
         
         W_select = self._var("dec_W_select", (self.N_bound,self.N_curve))
-        h_select = tf.tensordot(h_bound,W_select, axes=[-1,0])
+        b_select = self._var("dec_b_select",(self.N_curve,))
+        h_select = tf.tensordot(h_bound,W_select, axes=[-1,0]) + b_select
         if phase_act=="softmax":
             h_select = tf.nn.softmax(self.softmax_beta*h_select
                                      ,name=name)
@@ -353,3 +356,9 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         self.newt_ops = ops
         self.newt_step = ops[0]
         return self.newt_step
+
+    def update_beta(self,session,inc=0.1):
+        bnow = self.softmax_beta.eval(session=session)
+        session.run(self.update_beta_op,
+                    feed_dict={self.i_beta:bnow+inc})
+        

### 2019-06-15 autoencoder_trainer.py

@@ -56,10 +56,9 @@ class DoStuffHook(tf.train.SessionRunHook):
 def train_autoencoder(name, dataname, outerdim, innerdim, 
                           hyper=default_hyper,
                       training_dir='',data_dir='',n_epoch=5000, image_freq=1500):
-    
     hyperpath = string_identifier(hyper)
     training_dir = training_dir+"/training_"+name+"/"+hyperpath
-    
+    newt_freq = n_epoch
     graph = tf.Graph()
     with graph.as_default():
         # Load the data
@@ -94,7 +93,7 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
         loghook = tf.train.SummarySaverHook(
             summary_op=[tf.summary.scalar("goaltrain",ae.goal_all),
                         tf.summary.scalar("goaltest",ae.goal_test)],
-            save_steps=100,output_dir=training_dir)
+            save_steps=10,output_dir=training_dir)
         # Print to screen hook
         @DoStuffHook(freq=1)
         def printgoalhook(ctx,run_values):
@@ -151,11 +150,16 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
             # Do the SGD rounds
             i=0
             while not sess.should_stop():
-                if i%(n_epoch/2)==n_epoch/2-1:
+                # Do two newt steps throughout the training period
+                if i%(newt_freq)==newt_freq-1:
                     newtstep(sess)
                 else:
+                    ae.update_beta(sess,inc=0.1)
+
                     print("loop:",sess.run([ae.goal_all,ae.train_step]
                                        +list(ae._get_hess_vars())) )
+                    # increment beta
+                    #try:
                 i+=1
 
 

### 2019-06-15 training_water_lg.py

@@ -1,35 +1,34 @@
 if __name__=="__main__":
     training_dir = "/Users/afq/Google Drive/networks/"
-#    data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
-    data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
+    if True: #imac
+        data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
+    else:
+        data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
     dataset = "water_lg_sharded/*.csv"
 #     caes = [ 0.0, 1.0 ]
-    inis = [ 'rand', ]# 'pT', 'rhoh' ]
+    inis = [ 'rand','pT', 'rhoh' ]
     caes = [ 0.0, ]
     extend_train_params = lambda arch : \
         [ dict(ini=ini,cae=cae,**arch) for ini in inis for cae in caes ]
     archs_to_try = [
-#          {'type':'Classifying','args':[1,1, 3,3,'sigmoid']},
-#          {'type':'Classifying','args':[1,1, 6,12,'sigmoid']},
+        #{'type':'Classifying','args':[1,1, 3,2,'sigmoid']},
+        #{'type':'Classifying','args':[1,1, 3,3,'sigmoid']},
+        #{'type':'Classifying','args':[1,1, 6,12,'sigmoid']},
+        {'type':'Classifying','args':[1,2, 6,12,'sigmoid']},
+        {'type':'Classifying','args':[1,3, 3,3,'sigmoid']},
+        {'type':'Classifying','args':[1,3, 3,4,'sigmoid']},
+        {'type':'Classifying','args':[1,3, 3,5,'sigmoid']},
+        {'type':'Classifying','args':[1,3, 6,12,'sigmoid']},
+
         #{'type':'Poly','args':[1,2]},
         #{'type':'Poly','args':[1,3]},
         #{'type':'Poly','args':[1,5]},
-        {'type':'Classifying','args':[1,1, 3,2,'sigmoid']},
-        
-        
-        # {'type':'Classifying','args':[1,3, 6,12,'sigmoid']},
-#         {'type':'Classifying','args':[1,4, 6,12,'sigmoid']},
-        #         {'type':'Classifying','args':[1,4, 3,3,'sigmoid']},
[…]

### 2019-06-15 paper/paper.tex
equation for hessian augmentation
added a few citations to the paper

@@
 to solve. \cite{queiruga2018millstone}
-(In this paper, we use use $\partial_t$, $\mathrm{d}/\mathrm{d}t$,
+(In this paper, we use $\partial_t$, $\mathrm{d}/\mathrm{d}t$,
 and $\dot{()}$ to denote differentiation with respect to time
-interchangably depending on context.
-The complixity is hidden inside the relationships between temperature
+interchangeably depending on context.
+The complexity is hidden inside the relationships between temperature
 $T$, and pressure,$p$, density, $\rho$, and internal energy $e$ (or
@@
 % plasticity, }
+One of the simulators with the widest-encompassing physical modules is
+the TOUGH family of solvers that have been developed at Lawrence
+Berkeley National Lab.
+The methodology is based on a primary variable switching method in 
+The simulator of THAT GUY WITH DUAL VARIABLES I MET AT AGU
+parameterizes a simplified hydrate system using density and enthalpy
+as the primary variables.
+Higher order representations of flow are also a problem to tackle
+anisotropy and increase the computational efficiency of the spatial
+discretization. CITE ME AND MICKEY AND STEPHAN
+Multiphase flow was done with finite elements in \cite{YANG2014417}.
+THAT MORIDIS FEM PAPER
+However, the complexity of the state representation restricts
+development to a 2-phase
 
@@
 
-We redescribe our problem as a differential algebraic equation
+We redescribed our problem as a differential algebraic equation
 with a constraint: the balances of mass and energy are constrained by
@@
  A program synthesis approach was used to search for and evaluate possible methods of Peridynamics in Queiruga, 2017 \cite{queiruga2017numerical}.
-(An open source implementation of the study can be found Periflakes)\cite{queiruga2018periflakes}.
+(An open source implementation of the study can be found Periflakes\cite{queiruga2018periflakes}.)
 
@@
 \partial_t \rho &  = \mathbf{k}_p(p_\infty - p) + r\\
-\partial_t (\rho h-p) & = \mathbf{k}_T(T_\infty-T) + s \\
+\partial_t (\rho h-p) & = \mathbf{k}_T(T_\infty-T)  + s + h\left(\mathbf{k}_p(p_\infty - p) + r\right) \\
 \text{such that they lie on the material EOS,}\\
@@
 not depend on \(W^{dec}\).)
+
+\begin{equation}
+\bar{H}_{ij} = H_{ij} + \left(\delta_{ij}\,\text{if}\,H_{ik}=0\forall k\right)
+  \end{equation}
 
@@
 
-The architecture, illustrated in Figure \ref{fig:classifyingnetwork}, has the form:
+The architecture is illustrated in Figure \ref{fig:classifyingnetwork}.
+It has the following form in index notation:
 \begin{align}
-  q &= W_{enc} x^n + b_{enc}\\
-  f &= W_{dec} q^m + b_{dec} \\
-  h &= \sigma\left(W_{bound}q^m+b_{bound}\right) \\
-  p &= \text{softmax}\left(W_{phase}q^m+b_{bound[phase}\right) \\
-  \bar{x}&= f * p
+  q_i &= W^{enc}_{ij} x^n_j + b^{enc}_i\\
+  f_{ij} &= W^{dec}_{ijk} q^m_k + b^{dec}_{ij} \\
+  h_i &= \mu\left(W^{bound}_{ik}q^m_k+b^{bound}_{i}\right) \\
+  p_i &= \sigma\left(W^{phase}_{ij}h_j+b^{phase}_i\right) \\
+  \bar{x}_i&= f_{ij} p_j
 \end{align}
+where $x$ is the original prediction, $x^n$ denotes a polynomial basis
+set on $x$ (which increases the number of channels), $q$ is the latent
+space, $f$ is a set of polynomial curves, $h$ is a hidden layer
+denoting sides of boundaries, $p$ is a classification layer depending
+on $h$, and $\bar{x}$ is the predicted feature. The activation $\mu$
+is chosen as a standard neuron, e.g. $\tanh$ or $\text{sigmoid}$. The
+activation $\sigma$ is a probability activation, which is chosen to be
+a softmax,
+\begin{equation}
+\sigma_i(x;\beta) = \frac{e^{\beta x_i}}{\sum_j e^{\beta x_j}}
+\end{equation}
+or, in the limit, the hard argmax:
+\begin{equation}
+\sigma_i \rightarrow \text{onehot}\arg\max z as  \beta\rightarrow\infty
+\end{equation}
 
@@
 \end{figure}
-Because the identity of the phases,
+Because the identity of the phases,XXXXXXXX
 For comparison purposes, we group together liquid-gas-supercritical as one contiguous label.
@@
 
+Examining the $T,p,\rho$ surface allows for human interpretability, but
+the decoder actually yields four new functions with two unknowns:
+$D(q)=\{T(q),p(q),\rho(q),h(q)\}$. These surfaces can be plotted as
+independent 2D functions: the surface of \ref{fig:classified} is
+unfolded into the new parameterization in Figure \ref{fig:foursurfaces}.
+\begin{figure}
+  \centering
+  \includegraphics{../slides/water_four_surfaces_plot.png}
+  \caption{\label{fig:foursurfaces}The four 2D surfaces that make up
+    the new parameterization. The learned surface captures the sharp
+    kinks and vertical fronts in Figure \ref{fig:classifyingresult},
+    but is an analytic and smooth function of the two learned unknowns.}
+  \end{figure}
+
 With the trained model, we solve for \(q_1(t)\) and \(q_2(t)\) such that:
 \begin{align}
-\partial_t \rho(q_1,q_2) & = \nabla \cdot \mathbf{k} \nabla p(q_1,q_2) + r \\
-\partial_t \rho h(q_1,q_2)-p & = \nabla \cdot \mathbf{k'}\nabla T(q_1,q_2) + s
+\partial_t \rho(q_1,q_2) & = \mathbf{k} \left( p_\infty-p(q_1,q_2) \right)  + r \\
+\partial_t \left(\rho h(q_1,q_2)-p\right) & =\mathbf{k}_T(T_\infty-T(q_1,q_2))  + s + h\left(\mathbf{k}_p(p_\infty - p) + r\right)
 \end{align}
-where \(\rho(q_1,q_2)\) etc. are the back of an autoencoder:
-
-Using the autoencoder to automatically remove the local constraint:
-
+where the expressions for  \(\rho(q_1,q_2)\) etc. are the back of the
+trained autoencoder,. The trained
+autoencoder automatically removes the local constraint, satisfying
 \begin{equation}
@@
 \end{equation}
-
-Only training the material representation, not the balance
-laws.
-
-First, we had a constrained DAE:
+Thus, the system will always satisfy the constraints as $q$ is
+evolved, even during intermediate iterative solution steps. Note again
+that the machine learning is only applied to the
+material representation: the balance laws are expressed directly.
+
+To develop the rest of the discretization, we introduce the compact form:
 \begin{align}
@@
 
-We obtain four new functions with two unknowns. We're actually solving the manifold of Figure X.
-\begin{figure}
-  \centering
-  \includegraphics{../slides/water_four_surfaces_plot.png}
-  \caption{\label{fig:foursurfaces}The four 2D surfaces that make up the new parameterization.
-The learned surface captures the sharp kinks and vertical fronts in Figure \ref{fig:classifyingresult}, but is ananalytic and smooth function of the two learned unknowns.}
-  \end{figure}
-
 The decoder part of the network is inserted into the balance law
-
 \begin{equation}
@@
 \end{equation}
-
 We can parameterize this equation with any time stepping scheme, .e.g. a
@@
 
-This equation is deceptively simple: the function \(R(q)\) has complex logic encoded in it; if-then
+This equation is deceptively simple: the function $R(q)$ has complex logic encoded in it; if-then
 statements appear if the model used activations such as rectifiers.
@@
 
+\subsection{Dimensionalization}
+
+Recall that the datasets are normalized to be zero-center and
+unit-ranged to improve training. Thus, the outputs of the decoder are not in true
+units.
+The ranges are saved in a separate file which are loaded by the
+simulation. The network is loaded, the $D(q)$ is extracted, and the following
+transformations are applied to the outputs:
+\begin{align}
+T(q)      &= \left(T_{max}-T_{min}\right) D_T(q) +T_{mean} \\
+p(q)      &= \exp\left(\left( p_{max}-p_{min}\right)
+            D_p(q) +p_{mean} \right) \\
+\rho(q) &= \left(\rho_{max}-\rho_{min}\right) D_{\rho}(q) +\rho_{mean}\\
+h(q)      &= \left( h_{max}-h_{min}\right) D_h(q)  +h_{mean}
+\end{align}
+The exponent transformation is only applied to the cases where
+$\log(p)$ was used to make the training set.
+The range values are in the proper units, so that equation rows in the
+simulation are in SI units.
+
 \hypertarget{header-n3339}{%
@@
 There is a caveat with representing the degrees of freedom of the system
-on a learned space: the user does not know the meaning of \(q\) as an
-input. Further, the values of \(q\) vary from network to network and
+on a learned space: the user does not know the meaning of $q$ as an
+input. Further, the values of $q$ vary from network to network and
 depend on the parameterization. The user needs to be able to specify the
@@
     9e5 20 Aqu
-    0.5 20 AqG
+    0.4 20 AqG
 \end{verbatim}
@@
 but saturation in the second line for the liquid-gas equilibrium.
-This makes sense inuitively, as this is how humans are taught, but
+This makes sense to human intuition but
 still leads to bugs and occasionally requires looking up values in
-textbooks or the manual. (Is it $S_{gas}$ or $S_{aqu}$ on the second
+textbooks or the software manual. (Is it $S_{gas}$ or $S_{aqu}$ on the second
 line?) A fully fledged solver needs to check the user specified
-initial conditions for erroneous inputs.
+initial conditions for erroneous inputs, but it cannot detect ambiguities.
+
+The extra process of determining the initial state means that more error checking is needed
+in verifying user inputs, but this is necessary in any
+multiphase/multicomponent reservoir simulator. LatentSim does not yet
+detect and report errors, and only reports divergence. In our study, the
+testing problem specifications are all correct, but not all network
+realizations are acceptable. Failure to determine a \(q\) for one of the
+test initial conditions is a common mode of detecting unacceptable
+network realizations; this is marked as a testing error during training
+which invalidates a model as a potential realization.
+
+The current implementation LatentSim does not support inputting via
+saturations on the equilibria, a commonly used primary variable, but
+this problem could be solved with future engineering work for a
+non-experimental simulator.
 
@@
 on the decoder:
-\[D(q^0)_i = s^0_i \quad \text{for specified}\; i\]
+\begin{equation}
+  D(q^0)_i = s^0_i \quad \text{for specified}\; i
+\end{equation}
 This equation has two rows indexing into the four state variables;
@@
 the tangent matrix to $D$,
-\[K_{ij} = \frac{\partial D_i}{\partial q_j}  \quad \text{for specified}\; i, \text{for}\;j=1,2\]
+\begin{equation}
+K_{ij} = \frac{\partial D_i}{\partial q_j}  \quad \text{for
+    specified}\; i, \text{for}\;j=1,2
+\end{equation}
+
 and the following system of equations is solved for an update to $q$,
-\[K_{ij}\Delta q_j = s^0_i-D(q)_i\]
-The iteration u $q$ is repeated until convergence, with under-relaxation used again to update
+\begin{equation}
+K_{ij}\Delta q_j = s^0_i-D(q)_i
+\end{equation}
+The iteration on $q$ is repeated until convergence, with under-relaxation used again to update:
  \begin{equation}
@@
 0.1$.
-The initial condition problem is not gaurunteed to converge, however,
+
+The initial condition problem is not guaranteed to converge, however,
 and often does not.
@@
 It is further possible that none of the initial guesses for this iteration
-converge. There are two possiblities: the target state point is
+converge. There are two possibilities: the target state point is
 incorrectly specified or out-of-range (e.g., negative density), or the
 network is incapable of representing the desired material properties
-(i.e., poorly trained).
-
-The weakness of this necessity means that more error checking is needed
-in verifying user inputs, but this is necessary in any
-multiphase/multicomponent reservoir simulator. LatentSim does not yet
-detect and report errors, and only reports divergence. In our study, the
-testing problem specifications are all correct, but not all network
-realizations are acceptable. Failure to determine a \(q\) for one of the
-test initial conditions is a common mode of detecting unacceptable
-network realizations; this is marked as a testing error during training
-which invalidates a model as a potential realization.
-
-The current implementation LatentSim does not support inputting via
-saturations on the equilibria, a commonly used primary variable, but
-this problem could be solved with future engineering work for a
-non-experimental simulator.
+(i.e., poorly trained). The first possibility shows an invalid
+problem specification for the chosen EOS. The second possibility is
+crucial to the automated training process. If a given network cannot
+find an initial condition for every possible test problem, it fails
+the tests and is marked as invalid. This is one way of automatically
+verifying the training process.
 
@@
   Make a database of \(T,p,\rho,h,X_1,...\)
-
   \begin{itemize}
@@
 The codebase is released open source at
-\url{https://github.com/afqueiruga/latentsim}. Helper methods that the
+\url{https://github.com/afqueiruga/latentsim}. Helper methods and
+custom optimizers that the
 author reuses in various TensorFlow-based projects are factored out in
@@
 
-Three different extents:
+The phase space of water is broken up into three different extents to
+change the complexity of the training process:
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\item
-  Linear Equation of State
-
-  \begin{itemize}
-  \item
-    \textbf{Test!}
-  \item
-    Reduces to single phase Darcy's law problem (slide 1)
-  \item
-    \( p = 10^5+[-10^3, 10^3] Pa,\quad T = [ 19, 21 ] ^o C\)
-  \end{itemize}
-\item
-  Water Liquid-Gas Regime
-
-  \begin{itemize}
-  \item
-    One phase boundary 
-  \item
-    \( p = [100,5\times 10^5] Pa, \quad T = [274,594] K\)
-  \end{itemize}
-\item
-  Water Solid-Liquid-Gas-Supercritical Regimes
-
-  \begin{itemize}
-  \item
-    No linear mapping to latent space 
-  \item
-    Entire span
-  \item
-    \( p = [6\times 10^{-6},3\times 10^8]Pa, \quad T = [150,1000] K \)
-  \end{itemize}
+\item Linear Equation of State: the range is $ p = 10^5+[-10^3, 10^3]
+  Pa,\quad T = [ 19, 21 ] ^o C$. This reduces to the single phase
+  Darcy's law problem where the material has constant compressibility
+  and heat capacity. This is the most basic test of the training
+  framework, and provides analytical solutions to test the simulation
+  framework.
+\item Liquid-Gas Regime: the range is  $ p = [100,5\times 10^5] Pa,
+  \quad T = [274,594] K$. This restricts the domain to only one phase
+  boundary. With this range, it is possible to find a linear
+  combination of $T,p,\rho,h$ that can make a 1-1 mapping. (The usual
+  $T-p$ combination does not, but $\rho-h$ would.)
+\item Solid-Liquid-Gas-Supercritical Regimes: this covers the full
+  extent of the parameterized equation of state of water,
+    $ p = [6\times 10^{-6},3\times 10^8]Pa, \quad T = [150,1000] K
+    $. This problem has no linear mapping to a unique latent space.
 \end{enumerate}
@@
 \begin{table}
-  \caption{\label{tag:problems}Enumeration of the problems used to test and evaluate each fo the generaed architectures}
+  \caption{\label{tag:problems}Enumeration of the problems used to test and evaluate each fo the generated architectures}
 \begin{tabular}{l|c|c|c}
@@
 Single-grid block simulations were set up for most of the cases. TOUGH+
-does not handle supercritical \(H_2O\), so these test problems were not
+does not handle supercritical \ch{H2O}, so these test problems were not
 used to benchmark. (The supercritical phase region is not practically
 relevant for water, but it is very important for other fluids such as
-\(CO_2\).) (Note that, as a major motivation of this work, adding
+\ch{CO2}.) (Note that, as a major motivation of this work, adding
 supercritical states to the existing Fortran simulator would require
@@
 learning models into existing well known computational methodologies.
-This allows us to enforce physically known quanities, such as the
+This allows us to enforce physically known quantities, such as the
 balance laws.
@@
 The TensorFlow system is not quite suited for this type of workflow -\/-
-the resulting simualtions are extremely slow due to repeated calls to
+the resulting simulations are extremely slow due to repeated calls to
 session.run and a very complicated graph structure for the polynomials.
 Currently, the system is being reimplemented in Julia using Flux and
-Zygote to seemlessly reincorporate into a fully-fledged new reservoir
+Zygote to seamlessly reincorporate into a fully-fledged new reservoir
 simulator. The pendulum has been reimplemented in Julia at .
@@
 Each implemented equation of state (three options in this work spanning
-different extents of \(H_2O\)) has a database of potential realizations.
+different extents of \ch{H2O}) has a database of potential realizations.
 The entries are model architectures, parameters, and training
@@
 
-In TOUGH, for one instance, all of the calculations are performmed in SI units because it is simply too much labor to normalize everything!
+In TOUGH, for one instance, all of the calculations are performed in SI units because it is simply too much labor to normalize everything!
 Adverse effects to iterative algorithms and floating point error propogation occur, especially because numerical differentiation is used.
 The learned primary variables are normalized to a unit range around zero, and the scaling matrix is explicitly in one step.
-The scaling matrix could be a way to systematically nondimensionalize the entire problem.
+The scaling matrix could be a way to systematically non-dimensionalize the entire problem.
 

### 2019-06-15 plotting_routines.py
Plotting the surfaces inside of a notebook

@@ -118,7 +118,7 @@ def make_simulation_plot_list(database,eos_name):
     colorkey = defaultdict(lambda : colorring.next())
     sdb = SimDataDB(database)
     problems = sdb.Query('select distinct problem from {0}'.format(eos_name))
-    print problems
+    print(problems)
     plots = [ plot_one_simulation(sdb,eos_name, p[0], colorkey=colorkey) 
              for p in problems ]
     return plots

### 2019-06-15 paper/paper.tex
New citations

@@
 anisotropy and increase the computational efficiency of the spatial
-discretization. CITE ME AND MICKEY AND STEPHAN
-Multiphase flow was done with finite elements in \cite{YANG2014417}.
+discretization\cite{hannon_enhanced_2018}. 
+Multiphase flow was done with finite elements in \cite{yang_fully_2014}.
 THAT MORIDIS FEM PAPER
@@
 
- A program synthesis approach was used to search for and evaluate possible methods of Peridynamics in Queiruga, 2017 \cite{queiruga2017numerical}.
-(An open source implementation of the study can be found Periflakes\cite{queiruga2018periflakes}.)
+ A program synthesis approach was used to search for and evaluate
+ possible methods of Peridynamics in Queiruga, 2017
+ \cite{queiruga_numerical_2017}. (An open source implementation of the
+ study can be found Periflakes\cite{queiruga_periflakes_2019} using the
+ Popcorn/Cornflakes system for symbolic code generation
+ \cite{queiruga_cornflakes:_2018,queiruga_popcorn:_2018}.)
 
@@
 
-There are many ways to solve this cannonical equation: Lagrange
+There are many ways to solve this canonical equation: Lagrange
 multipliers (\(f(t)\) is the multiplier in the way we have written it

### 2019-06-15 plotting_routines.py
making animations

@@ -53,6 +53,22 @@ def plot_qq_Tprhoh(D,simplices,colorby=-1,offset=(0,0), name='', **kwargs):
     # return subfig
     return plts
 
+def read_all_surfaces(arch_dir):
+    files = list_files(arch_dir+'/surf_*.csv')
+    surfs = []
+    for f in files:
+        dat = np.loadtxt(f,delimiter=",",skiprows=1)
+        surfs.append( ( dat, Delaunay(dat[:,4:6]).simplices ) )
+    return surfs
+
+def make_animation(surfs):
+    return go.Figure(data=[plot_ptrho(surfs[0][0],surfs[0][1])],
+                     layout=dict(
+                         updatemenus= [{'type': 'buttons',
+                           'buttons': [{'label': 'Play',
+                                        'method': 'animate',
+                                        'args': [None]}]}])
+              frames=[{"data":[plot_ptrho(d,s)]} for d,s in surfs[1::20]])
 
 def read_networks(training_dir):
     archs = os.listdir(training_dir)

### 2019-06-16 dask_viz.py
made a dask app now

@@ -1,0 +1,64 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+import os
+import dash
+import dash_core_components as dcc
+import dash_html_components as html
+import plotly.graph_objs as go
+
+external_stylesheets = ['https://codepen.io/chriddyp/pen/bWLwgP.css']
+
+app = dash.Dash(__name__, external_stylesheets=external_stylesheets)
+server = app.server
+app.config.suppress_callback_exceptions = True
+from dash.dependencies import Input, Output, State
+
+from plotting_routines import *
+
+if 'DYNO' in os.environ:
+    app_name = os.environ['DASH_APP_NAME']
+else:
+    app_name = 'dash-text-annotationsplot'
+
+# Configure the eos hub
+hub = "/Users/afq/Google Drive/networks/"
+eoses = [
+        "water_slgc_logp_64",
+        "water_lg",
+        "water_linear",
+]
+eos = eoses[1]
+# Surfs
+surfs = read_networks(hub+'training_'+eos)
+surfs.pop('.DS_Store',None) # lol
+
+all_archs = list(surfs.keys())
+
+
+layout = html.Div([
+    html.Div([html.H1("Your 3D surfaces")], style={'textAlign': "center"}),
+    html.Div([html.Div([dcc.Dropdown(id='value-selected', 
+                                     options=[{'label':k,'value':k} for k in all_archs],
+                                     value=all_archs[0:2], multi=True)],
+                       style={"display": "block", "margin-left": "auto", "margin-right": "auto", "width": "60%"})],
+             className="row"),
+    dcc.Graph(id="my-graph"),
+], className="six columns")
+
+
+@app.callback(
+    Output("my-graph", "figure"),
+    [Input("value-selected", "value")])
+def update_graph(selected):
+    dropdown = {k:k for k in all_archs}
+    ctx = dash.callback_context
+    figure = plot_networks({k:surfs[k] for k in selected})
+    return figure
+    
+    
+
+
+app.layout = layout
+
+if __name__ == '__main__':
+    app.run_server(debug=True)

### 2019-06-17 dask_viz.py
Better viz, improving the dash app

@@ -5,10 +5,11 @@ import dash
 import dash_core_components as dcc
 import dash_html_components as html
 import plotly.graph_objs as go
+import dash_table
 
-external_stylesheets = ['https://codepen.io/chriddyp/pen/bWLwgP.css']
+# external_stylesheets = ['https://codepen.io/chriddyp/pen/bWLwgP.css']
 
-app = dash.Dash(__name__, external_stylesheets=external_stylesheets)
+app = dash.Dash(__name__)#, external_stylesheets=external_stylesheets)
 server = app.server
 app.config.suppress_callback_exceptions = True
 from dash.dependencies import Input, Output, State
@@ -34,23 +35,34 @@ surfs.pop('.DS_Store',None) # lol
 
 all_archs = list(surfs.keys())
 
-
-layout = html.Div([
-    html.Div([html.H1("Your 3D surfaces")], style={'textAlign': "center"}),
-    html.Div([html.Div([dcc.Dropdown(id='value-selected', 
+multi_dropdown = html.Div([html.Div([dcc.Dropdown(id='value-selected', 
                                      options=[{'label':k,'value':k} for k in all_archs],
                                      value=all_archs[0:2], multi=True)],
                        style={"display": "block", "margin-left": "auto", "margin-right": "auto", "width": "60%"})],
-             className="row"),
-    dcc.Graph(id="my-graph"),
-], className="six columns")
+             className="row")
+
+
+table = dash_table.DataTable(
+    id = "select-table",
+    columns = [{"name": "Name","id":"id"}],
+    data = [{"name":k,"id":k} for k in all_archs],
+    row_selectable = "multi",
+    sorting=True,
+    sorting_type="multi",
+    selected_rows=[0,1,2,3])
+
+layout = html.Div([
+    dcc.Graph(id="3d-graph"),
+    table,
+], className="two columns")
 
 
 @app.callback(
-    Output("my-graph", "figure"),
-    [Input("value-selected", "value")])
+    Output("3d-graph", "figure"),
+    [Input("select-table", "selected_row_ids")])
 def update_graph(selected):
-    dropdown = {k:k for k in all_archs}
+    if selected is None: selected = []
+    print(selected)
     ctx = dash.callback_context
     figure = plot_networks({k:surfs[k] for k in selected})
     return figure

### 2019-06-17 plotting_routines.py

@@ -13,6 +13,7 @@ import plotly.figure_factory as FF
 
 from SimDataDB import SimDataDB
 
+d_off = 2.0
 
 def list_files(fpattern):
     files = glob.glob(fpattern)
@@ -21,7 +22,7 @@ def list_files(fpattern):
     return files
 
 def make_tri_plot(x,y,z, simplices, c=None, offset=(0,0), name='', **kwargs):
-    return go.Mesh3d(x=x+1.5*offset[0],y=y,z=z+1.5*offset[1],
+    return go.Mesh3d(x=x+d_off*offset[0],y=y,z=z+d_off*offset[1],
                      i=simplices[:,0],j=simplices[:,1],k=simplices[:,2],
                      intensity=c/np.max(c),
                      name=name,showscale = True,**kwargs)
@@ -67,7 +68,7 @@ def make_animation(surfs):
                          updatemenus= [{'type': 'buttons',
                            'buttons': [{'label': 'Play',
                                         'method': 'animate',
-                                        'args': [None]}]}])
+                                        'args': [None]}]}]),
               frames=[{"data":[plot_ptrho(d,s)]} for d,s in surfs[1::20]])
 
 def read_networks(training_dir):
@@ -95,7 +96,10 @@ def plot_networks(surfaces,prefix=''):
     
     ptrhos = [ plot_ptrho(d,simp,offset=o,name=n)
                for (n,(d,simp)),o in zip(surfaces.items(),offsets) ]
-    fig = go.Figure(data=ptrhos)
+    annotes = [ dict(text=n,x=d_off*o[0],y=0,z=d_off*o[1])
+              for n,o in zip(surfaces,offsets)]
+    fig = go.Figure(data=ptrhos,
+                    layout=go.Layout(scene=dict(annotations=annotes)) )
     return fig
 
 #     qqrhos = [ plot_qqrho(d,simp,offset=o,name=n) 

### 2019-06-19 batch_test_latent_sim.py
Plotting reference results

@@ -56,7 +56,18 @@ class Hot_Gas():
     answer = dict(T=550,p=5.0e5,
         rho= iapws97.density_region2(550.0,5.0e5),
         h  =iapws97.enthalpy_region2(550.0,5.0e5) )
-        
+
+class Hot_Gas_Fill():
+    t_max = 1000.0
+    initial = dict(T=450,p=5.0e5, phase="Gas")
+    params =  dict(mass_source=0.1,k_p=0.0,k_T=0.0, Dt=t_max/1000.0)
+    @staticmethod
+    def schedule(sim,t):
+        pass
+    answer = dict(T=450,p=5.0e5,
+        rho= iapws97.density_region2(450.0,5.0e5),
+        h  =iapws97.enthalpy_region2(450.0,5.0e5) )
+    
 class Transition_L2G():
     t_max = 10.0
     initial = dict(T=350,p=5.0e5, phase="Liquid")
@@ -81,7 +92,7 @@ class Transition_L2G_Drain():
     
 class Liquid_Drain():
     t_max = 2000.0
-    initial = dict(T=373.0,p=3.0e5, phase="Liquid")
+    initial = dict(T=373.15,p=3.0e5, phase="Liquid")
     params =  dict(mass_source=-0.1,k_p=0.0,k_T=0.0, Dt=t_max/1000.0)
     @staticmethod
     def schedule(sim,t):
@@ -129,7 +140,8 @@ eoses = {
     'water_lg':dict(
         scale_file = "data_files/surf_ranges.csv",
         logp=False,
-        problem_list=['Small_Liquid','Small_Gas','Hot_Gas','Transition_L2G',]
+        problem_list=['Small_Liquid','Small_Gas','Hot_Gas','Transition_L2G','Transition_L2G_Drain'
+                     'Liquid_Drain']
     ),
     'water_linear':dict(
         scale_file = "data_files/water_linear_ranges.csv",
@@ -180,9 +192,14 @@ def perform_tests_for_eos(eos, result_dir='.'):
         return {'series':time_series}
     
     for n in networks:
-        for p in problem_list:
-            solve_a_problem(p,n)
-
+        try:
+            for p in problem_list:
+                solve_a_problem(p,n)
+        except Exception as e:
+            print("The network", n, " threw an error: ", e)
 if __name__=="__main__":
     for k in eoses:
-        perform_tests_for_eos(k, hub+'test_databases/')
+        try:
[…]

### 2019-06-19 dask_viz.py
fixed missing surfaces

@@ -7,7 +7,6 @@ import dash_html_components as html
 import plotly.graph_objs as go
 import dash_table
 
-# external_stylesheets = ['https://codepen.io/chriddyp/pen/bWLwgP.css']
 external_stylesheets = ['https://codepen.io/anon/pen/mardKv.css']
 
 app = dash.Dash(__name__)#, external_stylesheets=external_stylesheets)
@@ -114,7 +113,7 @@ layout = html.Div([
 def update_graph(selected):
     if selected is None: selected = []
     ctx = dash.callback_context
-    figure = ls_plot.plot_networks({k:surfs[k] for k in selected})
+    figure = ls_plot.plot_networks({k:surfs[k] for k in selected if k in surfs.keys()})
     return figure
     
     

### 2019-06-19 plotting_routines.py

@@ -22,6 +22,9 @@ def list_files(fpattern):
     files.sort(key=lambda f: grab_digit(f) )
     return files
 
+#
+# 3D Plotting of the learned surfaces
+#
 def make_tri_plot(x,y,z, simplices, c=None, offset=(0,0), name='', **kwargs):
     return go.Mesh3d(x=x+d_off*offset[0],y=y,z=z+d_off*offset[1],
                      i=simplices[:,0],j=simplices[:,1],k=simplices[:,2],
@@ -83,8 +86,9 @@ def read_networks(training_dir):
             dat = np.loadtxt(files[-1],delimiter=",",skiprows=1)
             surfaces[arch] = ( dat, Delaunay(dat[:,4:6]).simplices )
         except:
-            dat = np.zeros((3,7))
-            surfaces[arch] = ( dat, [] )
+            print("No surface for ",arch)
+            #dat = np.zeros((3,7))
+            #surfaces[arch] = ( dat, [] )
     return surfaces
 
 def generate_trisurf_plots(surfaces):
@@ -109,6 +113,10 @@ def plot_networks(surfaces,prefix=''):
 #     py.plot(fig,filename=prefix+'networks_q.html')
     
 
+    
+#
+# Simulations
+#
 def plot_one_simulation(sdb, eos_name, problem, colorkey=None):
     networks = sdb.Query('select distinct network from {0}'.format(eos_name))
     legends = ['T','p','rho','h']
@@ -203,6 +211,11 @@ def plot_simulations(database,eos_name,prefix=''):
 #     from IPython import embed ; embed()
     return subfig
     
+    
+    
+#
+# Generate a be-all-end-all report
+#
 if __name__=='__main__':
     hub = "/Users/afq/Google Drive/networks/"
     eoses = [

### 2019-06-19 autoencoder_trainer.py
Option for skipping newton step

@@ -55,7 +55,8 @@ class DoStuffHook(tf.train.SessionRunHook):
     
 def train_autoencoder(name, dataname, outerdim, innerdim, 
                           hyper=default_hyper,
-                      training_dir='',data_dir='',n_epoch=5000, image_freq=1500):
+                      training_dir='',data_dir='',n_epoch=5000, image_freq=1500,
+                     UseNewt = True):
     hyperpath = string_identifier(hyper)
     training_dir = training_dir+"/training_"+name+"/"+hyperpath
     newt_freq = n_epoch
@@ -76,7 +77,8 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
         global_step = tf.train.get_or_create_global_step()
         onum = tf.Variable(0,name="csv_output_num")
         ae = AutoencoderFactory(hyper,outerdim,innerdim,stream_mini, stream_all)
-        ae._make_hess_train_step(stream_all)
+        if UseNewt: # Leave it out of the graph if so; it's heavy and expensive to compute
+            ae._make_hess_train_step(stream_all)
         ae.goal_test = ae.make_goal(stream_test)
         init = tf.global_variables_initializer()
         meta_graph_def = tf.train.export_meta_graph(filename=training_dir+"/final_graph.meta")
@@ -143,7 +145,7 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
             i=0
             while not sess.should_stop():
                 # Do two newt steps throughout the training period
-                if i%(newt_freq)==newt_freq-1:
+                if UseNewt and i%(newt_freq)==newt_freq-1:
                     newtstep(sess)
                 else:
                     ae.update_beta(sess,inc=0.1)

### 2019-06-19 training_slgc.py

@@ -20,7 +20,7 @@ if __name__=="__main__":
         {'type':'Classifying','args':[2,4, 12,24,'sigmoid']},
         {'type':'Classifying','args':[2,5, 12,24,'sigmoid']},
         {'type':'Classifying','args':[2,6, 18,36,'sigmoid']},
-        {'type':'Classifying','args':[2,6, 24,48,'sigmoid']},
+        #{'type':'Classifying','args':[2,6, 24,48,'sigmoid']},
     ]
     sets_to_try = []
     for arch in archs_to_try:
@@ -34,13 +34,14 @@ if __name__=="__main__":
         # Import of tensorflow has to happen *after* forking
         print(S)
         from autoencoder_trainer import train_autoencoder
-        train_autoencoder("water_lg",dataset, 4,2,S,
+        train_autoencoder("water_slgc",dataset, 4,2,S,
                           training_dir=training_dir,
                           data_dir = data_dir,
                           n_epoch = n_epoch,
-                          image_freq = 20)
+                          image_freq = 50,
+                         UseNewt=False)
     if True: # Toggle parallel
-        p = multi.Pool(processes=4)
+        p = multi.Pool(processes=2)
         p.map( job, sets_to_try )
     else:
         for S in sets_to_try:

### 2019-06-19 training_water_lg.py

@@ -40,7 +40,7 @@ if __name__=="__main__":
                           training_dir=training_dir,
                           data_dir = data_dir,
                           n_epoch = n_epoch,
-                          image_freq = 10)
+                          image_freq = 50)
     if True: # Toggle parallel
         p = multi.Pool(processes=4)
         p.map( job, sets_to_try )

### 2019-06-19 dask_viz.py
Fixed that table bug

@@ -42,12 +42,13 @@ eoses = [
 eos = eoses[1]
 def get_it_all(eos):
     "Load the data for a particular EOS into server memory"
+    directory = hub+'training_'+eos
     # Surfaces
-    surfs = ls_plot.read_networks(hub+'training_'+eos)
+    surfs = ls_plot.read_networks(directory)
     surfs.pop('.DS_Store',None) # lol
     # Training and results 
     table = ls_grade.prep_table(eos,hub)
-    all_archs = list(surfs.keys())
+    all_archs = os.listdir(directory)
     return all_archs, table, surfs
 
 all_archs, archs_table, surfs = get_it_all(eos)

### 2019-06-19 dask_viz.py
Added the EOS selector

@@ -34,8 +34,9 @@ else:
 
 # Configure the eos hub
 hub = "/Users/afq/Google Drive/networks/"
+# TODO: read this from list dir
 eoses = [
-        "water_slgc_logp_64",
+        "water_slgc",
         "water_lg",
         "water_linear",
 ]
@@ -51,26 +52,41 @@ def get_it_all(eos):
     all_archs = os.listdir(directory)
     return all_archs, table, surfs
 
-all_archs, archs_table, surfs = get_it_all(eos)
-
+all_archs, archs_table, surfs = {}, {}, {}
+for eos in eoses:
+    all_archs[eos], archs_table[eos], surfs[eos] = get_it_all(eos)
+    for row in archs_table[eos]:
+        row["id"]=row["name"]
+    print(archs_table[eos])
 ##
 # trash
-multi_dropdown = html.Div([html.Div([dcc.Dropdown(id='value-selected', 
-                                     options=[{'label':k,'value':k} for k in all_archs],
-                                     value=all_archs[0:2], multi=True)],
-                       style={"display": "block", "margin-left": "auto", "margin-right": "auto", "width": "60%"})],
-             className="row")
+#
 
-@app.callback(
-    [Output('datatable-interactivity-channel', "selected_rows"),],
-    [Input('all-button', 'n_clicks'),],
-    [State('datatable-interactivity-channel', "derived_virtual_data"),]
-)
-def select_all(n_clicks, selected_rows):
-    if selected_rows is None:
-        return [[]]
-    else:
-        return [[i for i in range(len(selected_rows))]]
+# multi_dropdown = html.Div([html.Div([dcc.Dropdown(id='value-selected', 
+#                                      options=[{'label':k,'value':k} for k in all_archs],
+#                                      value=all_archs[0:2], multi=True)],
+#                        style={"display": "block", "margin-left": "auto", "margin-right": "auto", "width": "60%"})],
+#              className="row")
+
+# @app.callback(
+#     [Output('datatable-interactivity-channel', "selected_rows"),],
+#     [Input('all-button', 'n_clicks'),],
+#     [State('datatable-interactivity-channel', "derived_virtual_data"),]
+# )
+# def select_all(n_clicks, selected_rows):
+#     if selected_rows is None:
+#         return [[]]
+#     else:
+#         return [[i for i in range(len(selected_rows))]]
+    
+# dcc.Slider(
[…]

### 2019-06-20 dask_viz.py
reading list of all eoses

@@ -1,6 +1,6 @@
 #!/usr/bin/env python3
 # -*- coding: utf-8 -*-
-import os
+import os, glob
 import dash
 import dash_core_components as dcc
 import dash_html_components as html
@@ -35,12 +35,13 @@ else:
 # Configure the eos hub
 hub = "/Users/afq/Google Drive/networks/"
 # TODO: read this from list dir
-eoses = [
-        "water_slgc",
-        "water_lg",
-        "water_linear",
-]
-eos = eoses[1]
+eos_dirs = glob.glob(hub+'training_*')
+
+eoses = [ k[(len(hub)+len('training_')):] for k in eos_dirs ]
+#         "water_slgc",
+#         "water_lg",
+#         "water_linear",
+# ]
 def get_it_all(eos):
     "Load the data for a particular EOS into server memory"
     directory = hub+'training_'+eos
@@ -68,16 +69,18 @@ for eos in eoses:
 #                        style={"display": "block", "margin-left": "auto", "margin-right": "auto", "width": "60%"})],
 #              className="row")
 
-# @app.callback(
-#     [Output('datatable-interactivity-channel', "selected_rows"),],
-#     [Input('all-button', 'n_clicks'),],
-#     [State('datatable-interactivity-channel', "derived_virtual_data"),]
-# )
-# def select_all(n_clicks, selected_rows):
-#     if selected_rows is None:
-#         return [[]]
-#     else:
-#         return [[i for i in range(len(selected_rows))]]
+button = html.Button('Select All', id='my-button')
+@app.callback(
+    [Output('select-table', "selected_rows"),],
+    [Input('my-button', 'n_clicks'),],
+    [State('select-table', "derived_virtual_data"),]
+)
+def select_all(n_clicks, selected_rows):
+    if selected_rows is None:
+        return [[]]
+    else:
+        if len(selected_rows)==0:
+            return [[i for i in range(len(selected_rows))]]
     
 # dcc.Slider(
 #     min=-5,
@@ -96,7 +99,7 @@ eos_dropdown = dcc.Dropdown(id='eos-selected',
 
 # Prep the EOS selector table
 columns = ls_grade.table_column_names.copy()
-columns.append("id")
+# columns.append("id")
 
 table = dash_table.DataTable(
     id = "select-table",
@@ -115,6 +118,7 @@ layout = html.Div([
         html.Div([eos_dropdown],className="ten columns"),
     ],className="row"),
[…]

### 2019-06-22 paper/paper.tex
edited the paper

@@
 \documentclass[]{article}
+\usepackage[margin=1in]{geometry}
+\usepackage{amssymb,amsmath}
+\usepackage{chemmacros}
+\usepackage{siunitx}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
-\usepackage{chemmacros}
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 \usepackage{lineno}
-\usepackage[margin=1in]{geometry}
 \usepackage{fixltx2e} % provides \textsubscript
@@
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\title{Solution of Differential Algebraic Equations with Manifold Constraints
+% \makeatletter
+% \def\fps@figure{tbph}
+% \makeatother
+
+\title{Solution of Differential Algebraic Equations with Local Constraints
 on the Latent Space of an Autoencoder}
@@
 
-Our
-proposed method is motivated by the problems of multiphase reaction and
+Our proposed method is motivated by the problems of multiphase reaction and
 transport, where the underlying constraints are the ill-behaved material
@@
 
-Representing complex physical systems with multiple speicies, phases,
+Representing complex physical systems with multiple species, phases,
 chemical reactions, and heterogenous scale effects poses an
@@
 these relations in the rest of our analyses and production-scale
-computer simulations?
-
+computer simulations
 Accurately simulating multispecies reaction and transport with
@@
 systems.
+
+Our fundamental goal is to develop the tools
+for unsupervised representation learning in way where the
+representations have physical and useful meaning.
+How can we find a latent space that is good for our tasks?
+How can we find regressions in a deep learning context of high enough
+accuracy with needed qualities?
+How do we include these types of machine-generated models into a
+scientific and engineering calculation without sacrificing accuracy?
+Can this improve our workflow?
+Can we then find paths in the learned space that follow physical laws?
 
@@
 % {\bf ML LIT REVIEW HERE} % TODO
+Representation learning methods such as StyleGAN and G|FLOW have shown
+the ability to find meaningful dimensions in the latent space,
+e.g. the popular ability to transform age, race, and gender of human
+faces in a smooth way.
+But, can this directions be used meaningfully?
+
 \cite{wu_physics-informed_2018}
@@
 
-% \begin{figure}
-%    CSTR
-%   \caption{ \label{fig:cstr} Illustration of the Continuous stirred-tank
-%   reactor. The material in the reactor is exposed to a reservoir with pressure
-%   $p_\inf$ and temperature $T_\inf$ via a conduit with permeability
-%   and thermal conductivities $k_p$ and $k_T$, direct forced mass
-%   injection/withdrawl $r$, or heating and cooling $s$.}
-% \end{figure}
-
 For this preliminary work, we consider only a single-beaker problem
-without flow, a continuous stirred-tank reactor (CSTR), as illustrated
+without flow, the idealization of the continuous stirred-tank reactor (CSTR), as illustrated
 in Figure \ref{fig:cstr}. Let \(\rho\) be the material density, \(T\) be the
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=3in]{../slides/cstr.png}
+  \caption{ \label{fig:cstr} Illustration of the Continuous
+    Stirred-Tank    
+  Reactor. The material in the reactor is exposed to a reservoir with
+  pressure
+  $p_\infty$ and temperature $T_\infty$ via a conduit with permeability
+  and thermal conductivities $k_p$ and $k_T$, direct forced mass
+  injection or withdrawal $r$ via a pump, or heating and cooling $s$
+  via a heater/heat-exchanger. Stirring and insulation keeps the
+  material at a uniform pressure and temperature. Gravity is
+  neglected (though is usually the motivation for putting a feeding inlet at
+the top and product outlet at the bottom.)}
+\end{figure}
+
 We also consider the phase equilibria: beyond the scale of resolving
@@
 
+{\bf Remark:} This interpretation breaks down with the existenceof super-saturated phases in non-stable regions: e.g., it is possible to
+heat water beyond it's boiling point in a microwave. We are concerned
+with environments where this would not happen (the very rough and
+heterogenous) pores of rocks, so we have not thought of how to
+integrate this problem.
+
+\subsection{Forms of Equations of State}
+
+This work focuses on the EOS of pure water. The most accurate
+formulations were developed by the International Association for the
+Properties of Water Steam (IAPWS) using a variety of experimental
+sources for multiple phases \cite{wagner_iapws_2000,
+  water_revised_2007, water_revised_2009}. We utilize the 1997
+formulation for water gas, liquid, and supercritical phases, and the
+2006 formulation for the Ice Ih phase to build the dataset.\footnote{Our
+implementation is publicly available at GITHUB:EOS.} (Their
+painstaking work in determining the forms of decent fits is a great
+motivation for a 21st century approach.)
 We can also assemble the surface of possible $T,p,\rho$ coordinates,
@@
 \centering
-\includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/phase_diagram.png}
+\includegraphics[width=4in]{../figures/phase_diagram.png}
 \caption{\label{fig:phaseboundaries}Phase boundaries for water on the $T-p$ diagram. The standard
@@
 a phase boundary. The supercritical regime has no clear boundaries,
-but requires a seperate curve fit in practice due to numerical challenges.}
+but requires a separate curve fit in practice due to numerical challenges.}
 \end{figure}
@@
 \centering
-\includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/water_eos.png}
+\includegraphics[width=4in]{../figures/water_eos.png}
 \caption{\label{fig:pTrho} The $\log{p}$, $T$, $\rho$ equation of
@@
 
-{\bf Remark:}There are even more known phases of crystal ice which for water
+{\bf Remark:} There are even more known phases of crystal ice which for water
 For water, the supercritical phase and exotic ices only occur in
@@
 regime at \si{10^8}{Pa}. For other materials these exotic phases can be
-applicable. For example, in geoengineered systems using $CO_2$
+applicable. Methane $CH_4$ is actually well beyond its supercritical
+point at T=\si{-82.59}{C}\cite{the_engineering_toolbox_methane_2008}, so only an appropriate gas law can do the trick, until
+the bi-species hydrate phase comes into play.
+The liquid-gas-equilibrium-supercritical state transition comes into
+play in, for example, in geoengineered systems using $CO_2$
 such as geothermal plants are carbon sequestration facilities, the
-working fluid is usually in the supercritical regime at depth. cite
+working fluid is often in the supercritical regime at depth, and the
+phase boundary. cite
 curt. cite hydrates.
@@
 
-{\bf Remark:} This interpretation breaks down with the existenceof super-saturated phases in non-stable regions: e.g., it is possible to
-heat water beyond it's boiling point in a microwave. We are concerned
-with environments where this would not happen (the very rough and
-heterogenous) pores of rocks, so we have not thought of how to
-integrate this problem.
+{\bf Remark:} Many popular forms of EOS do not even have closed form
+expressions. The family of cubic equations of state for nonideal
+gases, such as a Peng-Robinsion equation \cite{peng_new_1976}, are
+expressed as functions for $p$ of $V$ (a proxy for $\rho$) and
+$T$. Since $\rho$ is rarely a primary varibale, these expressions must be inverted for
+$\rho$ given $p$ by solving the cubic equation. This can be performed
+analytically, as is done in TOUGH+, but often polynomial roots can
+be best solved numerically. When the equations become more
+complicated, such as when there are gas mixtures, the equations of
+state must be solved numerically. TOUGH+ employs a Richardson
+iteration \cite{moridis_users_2014-2}.
+Note how this is a numerical
+iteration {\em inside} of another numerical iteration that is trying
+to differentiation this equation state. Approximations
+abound--the mathematical theory is not straight foward. Passing
+automatic differentiation objects through these computations is
+difficult in practice. Equations are
+often solved faster and more accurately by numerical
+approximations--why not completely rewrite the equations?
+
+{\bf Remark:} The relations inside of EOSes are often artificially
+smoothed. For example, the hydrate equilibrium surface has a kink at
+the ice-liquid phase boundary in water. This causes problems
+numerically when a material state is moving along this equilibrium, so
+it is artificially smoothed using splines or the hyperbolic tangent to
+improve convergence in TOUGH+Hydrate \cite{moridis_users_2014}. Approximations abound.
 
@@
 \end{align}
-The complicatation is that the functions \(\rho(p,T)\) and \(u(p,T)\)
+The complication is that the functions \(\rho(p,T)\) and \(u(p,T)\)
 are not well defined functions due to the presence of phase changes that
@@
   \begin{figure}
-    \includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/phase_4.png}
+    \centering
+    \includegraphics[width=4in]{../slides/phase_4.png}
     \caption{\label{fig:statemachine}States and possible transitions for water.}
@@
 \centering
-\includegraphics[width=5in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/pendulum_q.png}
+\includegraphics[width=5in]{../figures/pendulum_q.png}
 \caption{\label{fig:pendulum}Solution of the pendulum on an arbitrary latent space.}
@@
 \subsection{Unsupervised Phase Classification}\label{header-n3317}}
+
+\begin{figure}
+\centering
+\includegraphics{../slides/classifier_network.png}
+\caption{\label{fig:classifyingnetwork}The architecture of the classifier network designed for
+  unsupervised labeling of phases.}
+\end{figure}
 
@@
 \begin{equation}
-\sigma_i \rightarrow \text{onehot}\arg\max z as  \beta\rightarrow\infty
+\sigma_i \rightarrow \text{onehot}\arg\max z \quad as\quad  \beta\rightarrow\infty
 \end{equation}
@@
   \centering
-\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_network.png}
-\caption{\label{fig:classifyingnetwork}The architecture of the classifier network designed for
-  unsupervised labeling of phases.}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=3in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_good.jpg}
+  \includegraphics[width=3in]{../slides/classifier_good.jpg}
   \caption{\label{fig:classified}Result of the unsupervised phase classifier. Compare this plot to the original labeled dataset Figure \ref{fig:pTrho}.}
@@
 \begin{align}
-\frac{\mathrm{d}}{\mathrm{d}t} m(T,p,\rho,h) &= r(T,p,\rho,h)\\
+\partial_t m(T,p,\rho,h) &= r(T,p,\rho,h)\\
 eos(T,p,\rho,h) &= 0
 \end{align}
-where $m=\{\rho,\rho h -p\}$ and $r$ is the right hand side, with no good representation of \(eos\) everywhere.
+where $m=\{\rho,\rho h -p\}$ and $r$ is the right hand side, where
+$eos(T,p,\rho,h)=0$ is the material constraint.
 
@@
 \begin{equation}
-\frac{\mathrm{d}}{\mathrm{d}t} m(D(q)) = r(D(q))
+\partial_t (D(q)) = r(D(q))
 \end{equation}
@@
 this paper is not exploring new time stepping schemes. This yields the
-discretizated algebraic equation for \(q_i\) and \(t+\Delta t\) given the
+discretized algebraic equation for \(q_i\) and \(t+\Delta t\) given the
 value \(q^0\) at \(t\),
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=6in]{../slides/user_workflow.png}
+  \caption{\label{fig:workflow}Work flow of the simulation programming process.}
+\end{figure}
+This process yields are more complicated simulation software system
+than previous generations. The architecture and workflow from problem
+specification to user experience is illustrated in Figure
+\ref{fig:workflow}.
+Previous generations of simulators were entirely composed of human
+written computer code, with some inline parameters (such as those of
+the IAPWS fits typed into constants in dedicated fortran files)
+fits and the occasional parameter files that are fed into the system
+as inputs (such as reaction rates databases.)
+Now, the behavior of the software is not only defined by the human
+written code, it is also defined by the model it loads, and the
+dataset that produced it.
+
+In our vision, the theoretician or experimentalist specify a dataset
+defined the possible states of the material. Multiple architectures
+are trained using the autoencoder unsupervised representation learning
+goal. Then, the simulation is ran by loading in each of the models and tested against known problems that
+span the problem space. Some of the tests are written by the original
+developers, and some of them must be specified by the scientist to
+pass through the newly added physical regimes.
+This testing process detects models that may have faired well in the
+ML process and produced low values of the loss function on testing and
+validation, but cannot be actually used in the simulation.
+This testing process is also used to determine the model that is
+optimal in terms of compute time and satisfactorily accurate.
+The ``best'' model can then be ``shipped'' (i.e., marked for
+deployment in DVC) to be automatically selected by the final end user,
+who only specifies the Equation of State.
+
+What is not shown in this figure is the testing of the data, and the
+testing of the simulation framework, which are needed in a
+fully-fledged machine learning based software system \cite{google ML smells}.
+Testing of the data is performed before the simulation
+methodology. These are contained in the Equations\_Of\_State package in our case, and would
+need to be performed by the theoretician and experimentalist feed-side
+users.
+Integration testing of the train-deploy-load-simulation framework is
+performed using the linear equation of state as a trivial test dataset
+with a trivial linear model.
+
 \hypertarget{header-n3353}{%
@@
 \hypertarget{header-n3356}{%
-\section{Implementation}\label{header-n3356}}
+\section{Architecture}\label{header-n3356}}
+
+The use of ML models yields a new software structure paradigm. It is now much simpler to implement new equation of state modules in
+the simulation package, but managing the datasets and models adds
+another layer to the software management. The high-level architecture
+and low-level implementation details developed for the prototype
+LatentSim are described in this section.
+
+LatentSim is implemented entirely in Python 3 using the
+
+The original datasets are written out as .csv files. To prep the
+datasets, the numbers are scaled to a zero-centered unit-range, shuffled, and broken up into the
+2/3-1/6-1/6 subdivisions for train, test, and validate subsets. The
+sets are stored as simple .npy files, with an accompanying short csv file
+for the scaling (which is loaded by the simulation to de-scale the
+network outputs.)
+
+The ``hub'' has a directory structure:
+\begin{verbatim}
+  1. training\_[EOS]
+    - One folder per architecture
+      - MonitoredTraningSession checkpoints
+      - Tensorboard summaries
+      - .csv 3d plot renderings
+    - More architectures...
+  2. training\_[other EOS]...
+  3. test\_database
+    - One sqlite file per EOS containing all tests
+  4. report
+    - static renderings (superseded by Dash app)
+\end{verbatim}
+The code has the following structure:
+\begin{verbatim}
+  1. sources for data
+  2. prepped data\_files
+  3. reference solutions for testing
+  4. training code
+  5. LatentSim.py
+  6. testing code
+  7. Dash app
+\end{verbatim}
+For every EOS, there are multiple generated implementations.
+After using a combination of Tensorboard, [] Visdom, [] and Paraview [] during
+development, a custom Dash [] web interface was ultimately developed
+to asses the results. A screenshot interface is shown in Figure []. The
+interface allows the developer (i.e., the author) to browse all
+statistics, renderings, of the trained surfaces and their test
+problems.
+
+In the version of the software distributed to users, only one
+``frozen'' and optimized model needs to be included per equation of state.
+
+\subsection{Implementation details}
 
@@
 The codebase is released open source at
-\url{https://github.com/afqueiruga/latentsim}. Helper methods and
+\url{https://github.com/afqueiruga/latentsim}. Datasets and other
+presentation artifacts such as videos and interactive plots are stored
+in git-lfs. Helper methods and
 custom optimizers that the
@@
 \href{https://github.com/afqueiruga/equations_of_state}{https://github.com/afqueiruga/equations\_of\_state}.
+
+\subsection{Things we'll do differently}
+
+Adopt julia for at least the simulation
 
@@
 In TOUGH, for one instance, all of the calculations are performed in SI units because it is simply too much labor to normalize everything!
-Adverse effects to iterative algorithms and floating point error propogation occur, especially because numerical differentiation is used.
+Adverse effects to iterative algorithms and floating point error propagation occur, especially because numerical differentiation is used.
 The learned primary variables are normalized to a unit range around zero, and the scaling matrix is explicitly in one step.
@@
 
-Techniquessuch as L1 regularization would be needed to search for low
+Techniques such as L1 regularization would be needed to search for low
 cost models that also satisfy our desire for Occam's razor
-simplifications. (The IAPWS relations used a prelimary form of this
+simplifications. (The IAPWS relations used a preliminary form of this
 type of simplification, but are not optimal in the author's opinion.)

### 2019-06-22 dask_viz.py
improved plotting and rendering

@@ -13,12 +13,12 @@ app = dash.Dash(__name__)#, external_stylesheets=external_stylesheets)
 app.css.append_css({
     "external_url": "https://codepen.io/chriddyp/pen/bWLwgP.css"
 })
-theme =  {
-    'dark': True,
-    'detail': '#007439',
-    'primary': '#00EA64',
-    'secondary': '#6E6E6E',
-}
+# theme =  {
+#     'dark': True,
+#     'detail': '#007439',
+#     'primary': '#00EA64',
+#     'secondary': '#6E6E6E',
+# }
 
 server = app.server
 app.config.suppress_callback_exceptions = True
@@ -79,8 +79,11 @@ def select_all(n_clicks, selected_rows):
     if selected_rows is None:
         return [[]]
     else:
+        print(selected_rows)
         if len(selected_rows)==0:
             return [[i for i in range(len(selected_rows))]]
+        else:
+            return [[]]
     
 # dcc.Slider(
 #     min=-5,

### 2019-06-22 plotting_routines.py

@@ -95,16 +95,19 @@ def generate_trisurf_plots(surfaces):
     return [ (n,go.Figure(data=[plot_ptrho(d,simp,name=n)]))
              for n,(d,simp) in surfaces.items() ]
 
-def plot_networks(surfaces,prefix=''):
+def plot_networks(surfaces,prefix='',aspectratio=1):
     offs = range(int(np.ceil(len(surfaces)**0.5)))
     offsets = list(itertools.product(offs,offs))
     
-    ptrhos = [ plot_ptrho(d,simp,offset=o,name=n)
+    ptrhos = [ plot_ptrho(d,simp,offset=(o[1],o[0]),name=n)
                for (n,(d,simp)),o in zip(surfaces.items(),offsets) ]
-    annotes = [ dict(text=n,x=d_off*o[0],y=0,z=d_off*o[1])
-              for n,o in zip(surfaces,offsets)]
+    annotes = [ dict(text=n,x=d_off*o[1],y=0,z=d_off*o[0])
+                for n,o in zip(surfaces,offsets) ]
     fig = go.Figure(data=ptrhos,
-                    layout=go.Layout(scene=dict(annotations=annotes)) )
+                    layout=go.Layout(
+                        scene=dict(annotations=annotes,
+                                   aspectmode='data'),
+                        margin=dict(r=5, l=5,b=5, t=5)) )
     return fig
 
 #     qqrhos = [ plot_qqrho(d,simp,offset=o,name=n) 
@@ -181,12 +184,12 @@ def plot_simulations(database,eos_name,prefix=''):
     gridy = int(numproblems / gridx)+1
     positions = list(itertools.product(range(1,gridx+1),range(1,gridy+1)))
 
-#     titles, titles_trans = [], range(4*gridx*gridy)
-#     for prob in problems:
-#         titles.extend([prob[0]] + ['_' for _ in range(4-1) ])
-#     for i in range(4*gridx):
-#         for j in range(gridy):
-#             titles_trans[j*4*gridx+i] = titles[4*i*gridy+j]
+    #     titles, titles_trans = [], range(4*gridx*gridy)
+    #     for prob in problems:
+    #         titles.extend([prob[0]] + ['_' for _ in range(4-1) ])
+    #     for i in range(4*gridx):
+    #         for j in range(gridy):
+    #             titles_trans[j*4*gridx+i] = titles[4*i*gridy+j]
     titles = ['' for _ in range(4*gridx*gridy)]
[…]

### 2019-06-22 dask_viz.py
fixed select button and better coloscheme

@@ -73,15 +73,16 @@ button = html.Button('Select All', id='my-button')
 @app.callback(
     [Output('select-table', "selected_rows"),],
     [Input('my-button', 'n_clicks'),],
-    [State('select-table', "derived_virtual_data"),]
+    [State('select-table', "derived_virtual_data"),
+     State('select-table', "derived_virtual_selected_rows"),]
 )
-def select_all(n_clicks, selected_rows):
+def select_all(n_clicks, all_rows, selected_rows):
     if selected_rows is None:
         return [[]]
     else:
         print(selected_rows)
         if len(selected_rows)==0:
-            return [[i for i in range(len(selected_rows))]]
+            return [[i for i in range(len(all_rows))]]
         else:
             return [[]]
     

### 2019-06-22 plotting_routines.py

@@ -29,7 +29,9 @@ def make_tri_plot(x,y,z, simplices, c=None, offset=(0,0), name='', **kwargs):
     return go.Mesh3d(x=x+d_off*offset[0],y=y,z=z+d_off*offset[1],
                      i=simplices[:,0],j=simplices[:,1],k=simplices[:,2],
                      intensity=c/np.max(c),
-                     name=name,showscale = True,**kwargs)
+                     name=name,showscale = True,
+                     colorscale='Jet',
+                     **kwargs)
 
 def plot_ptrho(D,simplices,colorby=-1,offset=(0,0), name='', **kwargs):
     return make_tri_plot(D[:,0],D[:,1],D[:,2], simplices, c=D[:,colorby],

### 2019-06-24 training_water_iapws_lg.py
another training batch

@@ -1,0 +1,49 @@
+if __name__=="__main__":
+    training_dir = "/Users/afq/Google Drive/networks/"
+    if True: #imac
+        data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
+    else:
+        data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
+    dataset = "water_iapws_lg_scaled"
+    eosname = "water_iapws_lg_tighter"
+#     caes = [ 0.0, 1.0 ]
+    inis = [ 'rand','pT', 'rhoh' ]
+    caes = [ 0.0, ]
+    extend_train_params = lambda arch : \
+        [ dict(ini=ini,cae=cae,**arch) for ini in inis for cae in caes ]
+    archs_to_try = [
+        {'type':'Classifying','args':[1,1, 3,2,'sigmoid']},
+        {'type':'Classifying','args':[1,1, 3,3,'sigmoid']},
+        {'type':'Classifying','args':[1,1, 6,12,'sigmoid']},
+        {'type':'Classifying','args':[1,2, 6,12,'sigmoid']},
+        {'type':'Classifying','args':[1,3, 3,3,'sigmoid']},
+        {'type':'Classifying','args':[1,3, 3,4,'sigmoid']},
+        {'type':'Classifying','args':[1,3, 3,5,'sigmoid']},
+        {'type':'Classifying','args':[1,3, 6,12,'sigmoid']},
+        #{'type':'Poly','args':[1,2]},
+        #{'type':'Poly','args':[1,3]},
+        #{'type':'Poly','args':[1,5]},
+    ]
+    sets_to_try = []
+    for arch in archs_to_try:
+        sets_to_try.extend(extend_train_params(arch))
+
+    n_epoch = 5000
+
+    import multiprocessing as multi
+    import itertools
+    def job(S):
+        # Import of tensorflow has to happen *after* forking
+        print(S)
+        from autoencoder_trainer import train_autoencoder
+        train_autoencoder(eosname ,dataset, 4,2,S,
+                          training_dir=training_dir,
+                          data_dir = data_dir,
+                          n_epoch = n_epoch,
+                          image_freq = 50)
+    if True: # Toggle parallel
+        p = multi.Pool(processes=4)
+        p.map( job, sets_to_try )
+    else:
+        for S in sets_to_try:
+            job(S)

### 2019-06-24 training_slgc.py
new training runs for the rho*h fix

@@ -4,7 +4,8 @@ if __name__=="__main__":
         data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
     else:
         data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
-    dataset = "water_iapw_logp_scaled"
+    dataset = "water_iapws_slgc_logp_scaled"
+    eosname = "water_iapws_rh_slgc_logp"
 #     caes = [ 0.0, 1.0 ]
     inis = [ 'rand','pT', 'rhoh' ]
     caes = [ 0.0, ]
@@ -20,7 +21,7 @@ if __name__=="__main__":
         {'type':'Classifying','args':[2,4, 12,24,'sigmoid']},
         {'type':'Classifying','args':[2,5, 12,24,'sigmoid']},
         {'type':'Classifying','args':[2,6, 18,36,'sigmoid']},
-        #{'type':'Classifying','args':[2,6, 24,48,'sigmoid']},
+        {'type':'Classifying','args':[2,6, 24,48,'sigmoid']},
     ]
     sets_to_try = []
     for arch in archs_to_try:
@@ -34,14 +35,14 @@ if __name__=="__main__":
         # Import of tensorflow has to happen *after* forking
         print(S)
         from autoencoder_trainer import train_autoencoder
-        train_autoencoder("water_slgc",dataset, 4,2,S,
+        train_autoencoder(eosname,dataset, 4,2,S,
                           training_dir=training_dir,
                           data_dir = data_dir,
                           n_epoch = n_epoch,
                           image_freq = 50,
                          UseNewt=False)
     if True: # Toggle parallel
-        p = multi.Pool(processes=2)
+        p = multi.Pool(processes=4)
         p.map( job, sets_to_try )
     else:
         for S in sets_to_try:

### 2019-06-24 training_water_iapws_lg.py

@@ -5,7 +5,7 @@ if __name__=="__main__":
     else:
         data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
     dataset = "water_iapws_lg_scaled"
-    eosname = "water_iapws_lg_tighter"
+    eosname = "water_iapws_rh_lg"
 #     caes = [ 0.0, 1.0 ]
     inis = [ 'rand','pT', 'rhoh' ]
     caes = [ 0.0, ]

### 2019-06-24 autoencoder.py
put labeling of the encoder back in

@@ -127,7 +127,12 @@ class Autoencoder(object):
         except KeyError:
             print("Making ",name)
             if not initial_value is None:
-                ini = tf.constant(initial_value,dtype=self.dtype) + tf.truncated_normal(shape=shape,
+                padded_ini = np.pad(initial_value,
+                                   [[0,shape[0]-initial_value.shape[0]],
+                                    [0,0]],'constant')
+                print(initial_value)
+                print(padded_ini)
+                ini = tf.constant(padded_ini,dtype=self.dtype) + tf.truncated_normal(shape=shape,
                        stddev=stddev, dtype=self.dtype)
             else:
                 ini = tf.truncated_normal(shape=shape,
@@ -238,13 +243,13 @@ class DeepPolyAutoencoder(Autoencoder):
         N_coeff = atu.Npolyexpand( self.size_x, self.Np_enc )
         nxt = atu.polyexpand(x, self.Np_enc)
         nxt = self._layers(nxt, self.enc_layers + [self.size_q], prefix="enc")
-        return tf.identity(nxt)
+        return tf.identity(nxt, name=name)
     
     def decode(self, q, name=None):
         N_coeff = atu.Npolyexpand( self.size_q, self.Np_dec )
         nxt = atu.polyexpand(q, self.Np_dec)
         nxt = self._layers(nxt, self.dec_layers + [self.size_x], prefix="dec")
-        return tf.identity(nxt)
+        return tf.identity(nxt, name=name)
     
     
 class ClassifyingPolyAutoencoder(Autoencoder):
@@ -290,7 +295,7 @@ class ClassifyingPolyAutoencoder(Autoencoder):
         W = self._var("enc_W", (N_coeff, self.size_q), 
                      initial_value=encoder_init_options[self.encoder_init])
         b = self._var("enc_b", (self.size_q,) )
-        return tf.add(tf.matmul( atu.polyexpand(x, self.Np_enc), W ), b)
+        return tf.add(tf.matmul( atu.polyexpand(x, self.Np_enc), W ), b, name=name)
     
     def classify(self, q, name=None, phase_act="softmax"):
         qpoly = atu.polyexpand(q, self.Np_dec)

### 2019-06-24 batch_test_latent_sim.py
added a test problem

@@ -143,6 +143,13 @@ eoses = {
         problem_list=['Small_Liquid','Small_Gas','Hot_Gas','Transition_L2G','Transition_L2G_Drain'
                      'Liquid_Drain']
     ),
+     'water_iapws_lg_logp':dict(
+        scale_file = "data_files/water_iapws_lg_ranges.csv",
+        logp=False,
+        problem_list=['Small_Liquid','Small_Gas','Hot_Gas','Transition_L2G',
+                      'Transition_L2G_Drain'
+                     'Liquid_Drain','Hot_Gas_Fill']
+    ),
     'water_linear':dict(
         scale_file = "data_files/water_linear_ranges.csv",
         logp=False,

### 2019-06-24 autoencoder_trainer.py
removed a print

@@ -149,9 +149,9 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
                     newtstep(sess)
                 else:
                     ae.update_beta(sess,inc=0.1)
-
-                    print("loop:",sess.run([ae.goal_all,ae.train_step]
-                                       +list(ae._get_hess_vars())) )
+                    sess.run(ae.train_step)
+#                     print("loop:",sess.run([ae.goal_all,ae.train_step]
+#                                        +list(ae._get_hess_vars())) )
                     # increment beta
                     #try:
                 i+=1

### 2019-06-24 plotting_routines.py

@@ -28,7 +28,7 @@ def list_files(fpattern):
 def make_tri_plot(x,y,z, simplices, c=None, offset=(0,0), name='', **kwargs):
     return go.Mesh3d(x=x+d_off*offset[0],y=y,z=z+d_off*offset[1],
                      i=simplices[:,0],j=simplices[:,1],k=simplices[:,2],
-                     intensity=c/np.max(c),
+                     intensity=(1.0*c)/np.max(1.0*c),
                      name=name,showscale = True,
                      colorscale='Jet',
                      **kwargs)

### 2019-07-11 dash_viz.py
Plotting the entire training history

@@ -38,11 +38,13 @@ else:
 hubs = [
     "/Users/afq/Google Drive/networks/",
     "/Users/afq/Documents/Research/LBNL/eoshub/eoshub/networks/",
+    "/Users/afq/Research/eoshub/networks/",
+    
 ]
 
 class LazyLoad():
     "Lazily load the state of all of the eoses"
-    Entry = namedtuple('Entry', ['archs', 'table','surfs'])
+    Entry = namedtuple('Entry', ['archs', 'table','surfs','train_scores'])
     def __init__(self,hubs):
         self.eoses = []
         self.cache = {} # Cache processing results
@@ -66,13 +68,14 @@ class LazyLoad():
         surfs = ls_plot.read_networks(directory)
         surfs.pop('.DS_Store',None) # lol
         # Training and results 
-        table = ls_grade.prep_table(eos,self.hubs[eos])
+        
+        table, train_scores = ls_grade.prep_table(eos,self.hubs[eos])
         all_archs = os.listdir(directory)
         # TODO: Get the simulation results
         # TODO: Run the simulations in batch
         for row in table:
             row["id"]=row["name"]
-        return self.Entry(all_archs, table, surfs)
+        return self.Entry(all_archs, table, surfs, train_scores)
     
     def __getitem__(self, key):
         try:
@@ -84,6 +87,9 @@ class LazyLoad():
 
 loaded = LazyLoad(hubs)
 
+#
+# The components
+#
 # Prep the EOS selector table
 columns = ls_grade.table_column_names.copy()
 eos_dropdown = dcc.Dropdown(id='eos-selected',
@@ -96,7 +102,8 @@ graph_radio = dcc.RadioItems(
     options=[
         {'label': '3D rho', 'value': 'rho'},
         {'label': '3D rho_h', 'value': 'rho*h'},
-        {'label': 'simulations', 'value': 'simulations'}
+        {'label': 'Traning Loss', 'value': 'training'},
+        {'label': 'simulations', 'value': 'simulations'},
     ],
     value='rho',
     labelStyle={'display': 'inline-block'}
@@ -107,8 +114,9 @@ table = dash_table.DataTable(
     columns = [{"name":k,"id":k} for k in columns],
     #data = None, #filled by callback
     row_selectable = "multi",
-    sorting=True,
-    sorting_type="multi",
+    sort_action="native",
+    #sorting=True,
+    #sorting_type="multi",
     selected_rows=[0,1,2,3])
 
 
[…]

### 2019-07-11 grading_routines.py

@@ -15,7 +15,8 @@ from batch_test_latent_sim import problems as test_problems
 def extract_scalars(directory):
     # TODO: check if its a top-directory or the tf training session
     archs = os.listdir(directory)
-    scalars = {}
+    final_scores = {}
+    train_scores = {}
     for arch in archs:
         path = os.path.join(directory,arch)
         try:
@@ -23,10 +24,12 @@ def extract_scalars(directory):
             acc.Reload()
             tags = acc.Tags()['scalars']
             w_times, step_nums, vals = zip(*acc.Scalars('goaltest'))
-            scalars[arch] = np.mean(vals[-10:])
+            train_scores[arch] = np.stack([step_nums,vals]).T
+            final_scores[arch] = np.mean(vals[-10:])
         except:
             pass
-    return scalars
+    return final_scores, train_scores
+    
     
 def grade_simulations(database,eos_name):
     """Examine and distill the results for each of the architectures"""
@@ -76,12 +79,12 @@ table_column_names = [" ","name","loss","successes","total_run_time","red_flag",
 def prep_table(eos,hub):
     test_db = hub+'test_databases/'+eos+'_testing.db'
     training_dir = hub+'training_'+eos 
-    training_scores = extract_scalars(training_dir)
+    final_scores, train_scores = extract_scalars(training_dir)
     table = []
-    for k in training_scores:
-        row = {"name":k,"loss":training_scores[k]}
+    for k in train_scores:
+        row = {"name":k,"loss":final_scores[k]}
         table.append(row)
-    return table
+    return table, train_scores
 
 if __name__=='__main__':
     hub = "/Users/afq/Google Drive/networks/"

### 2019-07-11 plotting_routines.py

@@ -14,7 +14,7 @@ import plotly.figure_factory as FF
 from SimDataDB import SimDataDB
 
 color_ring = itertools.cycle(['black','orange','purple','red','blue','green'])
-colorkey = defaultdict(lambda : next(pr.color_ring)) # It should be global, no?
+colorkey = defaultdict(lambda : next(color_ring)) # It should be global, no?
 d_off = 2.0
 
 def list_files(fpattern):
@@ -22,6 +22,16 @@ def list_files(fpattern):
     grab_digit = lambda f : int(re.search("([0-9]*)\.[a-zA-Z]*$",f).groups()[-1])
     files.sort(key=lambda f: grab_digit(f) )
     return files
+
+#
+# Plotting of all of the scores
+#
+def make_training_plot(train_scores):
+    layout = dict(
+        yaxis=dict(title='Test loss',type='log'),
+        xaxis=dict(title='Step number'))
+    return go.Figure(data=[ go.Scatter(x=v[:,0],y=v[:,1],name=k)
+                            for k,v in train_scores.items() ], layout=layout)
 
 #
 # 3D Plotting of the learned surfaces
@@ -131,8 +141,8 @@ def plotly_simulation_t(sims,ref, showleg = False):
     legends=['T','p','rho','rho*h']
     traces = []
     for i,name in enumerate(legends):
-        trace_ref = go.Scatter(x=ref_arr_prep[:,0],
-                              y=ref_arr_prep[:,i+1],
+        trace_ref = go.Scatter(x=ref[:,0],
+                              y=ref[:,i+1],
                               name='ref',legendgroup='ref',
                               showlegend= (i==0 and showleg),
                               line=dict(dash='dash',color=colorkey['ref']))
@@ -173,14 +183,20 @@ def plotly_simulations_Tp(sims,ref):
 def plotly_simulations_Tp_ts(sims,ref):
     trace_Tp, layout_Tp = plotly_simulations_Tp(sims,ref)
     traces_ts = plotly_simulation_t(sims,ref)
-    subfig = pytools.make_subplots(rows=4,cols=2,
+    subfig = tools.make_subplots(rows=4,cols=2,
                 specs=[[{'rowspan':3,'colspan':1}, {}],
                 [None, {}],
                 [None, {}],
                 [None, {}],],
-                shared_xaxes=True)
+                shared_xaxes=False)
[…]

### 2019-07-11 paper/paper.tex
Edits to the paper

@@
 \usepackage{siunitx}
+\usepackage{algorithm}
 \usepackage{lmodern}
@@
 laws, where now the traditional partial differential equations are being
-solved on the learned latent space. This provides a more robust
+solved on the learned latent space.
+We train an autoencoder to determine the lower-dimensional
+representation of the EOS surface using modern deep learning techniques
+without using any phase labels. The decoder end of the model is then
+plugged back into the balance laws, eliminating the constraint equation.
+This provides a more robust
 numerical algorithm, while putting the onus of programming part of the
@@
 balance laws. Through deep learning approaches combined with
-domain-specific goals, we can learn feature sets that are both
+domain-specific goals, we can learn representations that are both
 interpretable and directly apply to analyzing larger systems.
-
-We train an autoencoder to determine the lower-dimensional
-representation of the EOS surface using modern deep learning techniques
-without using any phase labels. The decoder end of the model is then
-plugged back into the balance laws, eliminating the constraint equation.
+We illustrate the methodology on two traditional problems: 1) the
+pendulum, and 2) the continuous stirred tank reactor.
+
 \end{abstract}
@@
 problems.
-This work is motivated by advancing simulation techniques for subsurface energy systems such as shale
-and oceanic natural gas reservoirs in shales or engineered geothermal
+This work is motivated by addressing challenges in the simulation of subsurface energy systems, such as shale
+and oceanic natural gas reservoirs or engineered geothermal
 systems.
@@
 reactions, and phase transitions makes the equations extremely difficult
-to solve. \cite{queiruga2018millstone}
-(In this paper, we use $\partial_t$, $\mathrm{d}/\mathrm{d}t$,
+to solve \cite{queiruga_simulation_2019}. 
+(In this paper, we use $\partial_t$
 and $\dot{()}$ to denote differentiation with respect to time
-interchangeably depending on context.
+for compactness.)
 The complexity is hidden inside the relationships between temperature
@@
 Multiphase flow was done with finite elements in \cite{yang_fully_2014}.
-THAT MORIDIS FEM PAPER
-However, the complexity of the state representation restricts
-development to a 2-phase
+THAT MORIDIS FEM PAPER SAY MORE ABOUT IT
 
@@
 % {\bf ML LIT REVIEW HERE} % TODO
-Representation learning methods such as StyleGAN and G|FLOW have shown
+Representation learning methods such as StyleGAN
+\cite{karras_style-based_2018} and Flow/Glow \cite{kingma_glow:_2018} have shown
 the ability to find meaningful dimensions in the latent space,
@@
 But, can this directions be used meaningfully?
+
+Recent review papers have highlighted open applicaitons of machine
+learning to subsurface systems \cite{bergen_machine_2019} and
+climate scale systems  \cite{reichstein_deep_2019}.
+History matching on the latent space of a GAN was demonstrated on a
+synthetic geomorphology by Mosser, 2019 \cite{mosser_deepflow:_2019}.
+
+Many approaches try to use machine learning to learn the dynamics of a
+system.
+The method of Dynamic Mode Decomposition
+to learn a linear operator based on time series data \cite{tu_dynamic_2014,kutz_dynamic_2016}.
+
+Extending this space with known features, such as partial derivatives,
+\cite{ one of kutz's students}
+Our presented approach differs from this approach by saying that we
+know the dynamics {\em a priori}, but do not how to represent the
+current state in a useful way. (Our approach will be later extended to
+learn dynamics in a second step using the learned representation as a
+fixed input.)
 
@@
 human-written code with computer-written code. We are applying this
-mindset from industry problems in the Artificial
-Intelligence space into scientific computing.
+mindset from industry problems in the artificial
+intelligence space into scientific computing.
 
@@
 the simulator.
+
+There are infinitely many representations of a system. The simplest and clearist description is not neccessarily the best way
+to solve it, and vice versa.
 
@@
 
-We build a test dataset off of the IAPWS solutions. The proposed
+We build a dataset using the IAPWS solutions. The proposed
 methodology applies to experimental data directly, or preexisting
@@
 
-In section 2, the balance of mass and energy will be described for a
+In Section 2, the balance of mass and energy will be described for a
 single multiphase material. We will discuss the current state of the
@@
 constraint problem in th above form.
-In section 3, we will demonstrate applying the autoencoder learning
+In Section 3, we will demonstrate applying the autoencoder learning
 methodology to solve the pendulum,
 which can be considered the archetypical constraint problem.
-In section 4, the autoencoder methodology wil lbe applied to the
-equation of state of water. In section 5, the new simulation process
+In Section 4, the autoencoder methodology wil lbe applied to the
+equation of state of water. In Section 5, the new simulation process
 will be described, wherein the learned decoders are placed into the
@@
 representations. The software implementation of the training and
-simulation programs are detailed in section 6. The new simulation
+simulation programs are detailed in Section 6. The new simulation
 framework on sets of learned representations is evaluated on
 multiple datasets of varying extents of water with a
-hyperparameter search on a suite of test problems to prouce the discussion of section 7, as
+hyperparameter search on a suite of test problems to prouce the discussion of Section 7, as
 well as a systematic automated testing procedure of a
@@
 without flow, the idealization of the continuous stirred-tank reactor (CSTR), as illustrated
-in Figure \ref{fig:cstr}. Let \(\rho\) be the material density, \(T\) be the
-temperature, \(p\) be the material pressure, \(h\) be the material
-enthalpy, and \(u\) be the internal energy. There are two equations for
+in Figure \ref{fig:cstr}. 
+In the analysis of geological fluids, this idealization is used in the
+assumption of well-mixed pore spaces at the macroscale.  Let \(\rho\)
+be the material density, measured in \si{kg/m^3}; \(T\) be the
+temperature, measured in \si{K}; \(p\) be the material pressure,
+measured in \si{Pa}; \(h\) be the material
+specific enthalpy, measured in \si{J/kg}; and \(u\) be the internal
+energy, measured in \si{J/kg}, and related to the enthalpy by  \(\rho u = \rho h - p\). There are two equations for
 balance of mass and energy.
@@
 \partial_t \rho & = \mathbf{k}_p(p_\infty - p) + r\\
-\partial_t \rho u & = \mathbf{k}_T(T_\infty-T) + s 
+\partial_t \rho u & = \mathbf{k}_T(T_\infty-T) + s + h q_{mass}
 \end{align}
@@
 via a mixed boundary condition straw with transport and conduction
-coefficients \(\mathbf{k}_p\), \(\mathbf{k}_T\). Mass and heating
-sources and sinks are the terms \(r\) and \(s\).\footnote{We express the
-  material by enthalpy instead of internal energy; they are related by
-  the relation \(\rho u = \rho h - p\). Future work may work directly
-  with internal energy.}
+coefficients \(\mathbf{k}_p\), \(\mathbf{k}_T\). 
 We are neglecting upstreaming in the connection between the reservoir
-and the container.
+and the container. Mass and heating
+sources and sinks are the terms \(r\) and \(s\).
 
@@
 individual bubbles, it is possible
-for a control volume to be a mixture of two phase coexisting.
+for a control volume to be a mixture of multiple coexisting
+phases.
 (For a very fine scale simulation, e.g. at the pore scale, bubbles
@@
 properties of all phases at that point in the phase diagram:
-\begin{equation}
-  \rho = S_{gas} \rho_{gas} + S_{liq} \rho_{liq} + S_{ice}
-    \rho_{ice} + S_{crit} \rho_{crit}
-  \end{equation}
+\begin{align}
+  \rho = & S_{gas} \rho_{gas} + S_{liq} \rho_{liq} + S_{ice}
+  \rho_{ice} + S_{crit} \rho_{crit} \\
+  \rho h =& S_{gas} \rho_{gas} h_{gas} + S_{liq} \rho_{liq} h_{liq} + S_{ice}
+  \rho_{ice} h_{ice} + S_{crit} \rho_{crit} h_{crit}
+\end{align}
 Each phase
@@
 hydrates.
+
+Note that the definition of enthalpy as a per-mass quanitity means the mixture needs to be weighted by the
+density: $h_{mix}=\sum{\rho_a h_aS_a}/\sum{rho_aS_a}$.
 
@@
 expressions. The family of cubic equations of state for nonideal
-gases, such as a Peng-Robinsion equation \cite{peng_new_1976}, are
+gases, such as the Peng-Robinsion equation \cite{peng_new_1976}, are
 expressed as functions for $p$ of $V$ (a proxy for $\rho$) and
@@
 \partial_t \rho(p,T) & = \nabla \cdot \mathbf{k}\nabla p + r\\
-\partial_t \rho u(p,T) & = \nabla \cdot \mathbf{k'}\nabla T + s
+\partial_t \rho(p,T) h(p,T)-p & = \nabla \cdot \mathbf{k'}\nabla T + s
 \end{align}
@@
 water.
-
+Each state requires a potentially different set of primary
+variables. For a pure water system, this would have to be:
+\begin{center}
 \begin{tabular}{l|c|c}
@@
 \end{tabular}
-
-  This requires a statemachine. For the complete states of water, it is shown in Figure \ref{fig:statemachine}.
+\end{center}
+Solving the balance laws with this caveat requires the incorpoeration
+of a statemachine. Representing the material with the known
+representations thus requires two variables and an additional state
+tag. For the complete states of water, the states and their possible
+transitions is shown in Figure \ref{fig:statemachine}.
   \begin{figure}
@@
 \[phase^{k+1},X^{k+1} = statemachine(X^{k+1/2} ; phase^k)\]
-The phase switching logic has the form of a switch case:
-
+The phase switching logic has the form of a switch case illustrated in
+Algorithm \ref{alg:statemachine}.
+
+\begin{algorithm}
+  \caption{\label{alg:statemachine}Statemachine transition logic}
 \begin{Shaded}
@@
 \end{Shaded}
+\end{algorithm}
 
@@
 \begin{align}
-\text{Solve for}\, \rho(t), \, h(t), \, p(t),\, \text{and}\, T(t)\, \text{satisfying:}\\
+\text{Solve for}\, \rho(t), \, h(t), \, p(t),\, \text{and}\, T(t)\,
+  \text{satisfying:} \nonumber
+                       \\
 \partial_t \rho &  = \mathbf{k}_p(p_\infty - p) + r\\
 \partial_t (\rho h-p) & = \mathbf{k}_T(T_\infty-T)  + s + h\left(\mathbf{k}_p(p_\infty - p) + r\right) \\
-\text{such that they lie on the material EOS,}\\
-\rho,h,p,T & \in \{\text{EOS}\}
+\text{such that they lie on the material EOS,}\nonumber
+                                                \\
+  \rho,h,p,T & \in \{\text{EOS}\}\\
+  \text{given }\rho,h,p,T \text{ at } t=0. \nonumber
 \end{align}
@@
 \begin{align}
-\text{Solve for}\, x(t), \, y(t), \, f(t) \, \text{satisfying:} \\
+\text{Solve for}\, x(t), \, y(t), \, f(t) \, \text{satisfying:} \nonumber \\
 m \ddot{x} & = f x/\sqrt{x^2+y^2} \\
@@
   x^2 + y^2 & = R^2
+  \text{given } x,y, \dot{x},\dot{y} \text{ at } t=0. \nonumber
 \end{align}
@@
 \begin{equation}
-L = \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right) - m g y
+\mathcal{L} = \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right) - m g y
 \end{equation}
@@
 \begin{align}
-\text{Solve for}\, \theta(t) \ \text{satisfying:} \\
-\frac{\mathrm{d}}{\mathrm{d}t} \frac{\partial L}{\partial
+\text{Solve for}\, \theta(t) \ \text{satisfying:} \nonumber \\
+\frac{\mathrm{d}}{\mathrm{d}t} \frac{\partial \mathcal{L}}{\partial
   \dot{\theta}} =
-  \frac{\partial L}{\partial \theta}
+  \frac{\partial \mathcal{L}}{\partial \theta}
+   \text{given } \theta,\dot{\theta} \text{ at } t=0. \nonumber
 \end{align}
@@
 \begin{equation}
-X = \left\{ (R\cos\theta_i, R\sin\theta_i)\; | \;\theta_i \in [-5\pi/4,\pi/4] \right\}
+y = \left\{ (R\cos\theta_i, R\sin\theta_i)\; | \;\theta_i \in [-5\pi/4,\pi/4] \right\}
 \end{equation}
 We also try a ``goofier'' looking curve with no known closed form
-solution to exercise the methodology,
-
-(For the equation of state problem, this is analogous to the dataset
-of $p,T,\rho,h$ 4-tuples that make the plot in Figure \ref{fig:pTrho}.)
+solution to exercise the methodology.
+For the equation of state problem, this is analogous to the dataset
+of $p,T,\rho,h$ 4-tuples that make the plot in Figure \ref{fig:pTrho}.
 
@@
 We'll use the autoencoder to generate the parameterization of \(x(q)\)
-and \(y(q)\). To do this, we optimize the loss function,
-\begin{equation}
-L(x) = \left\| x-D(E(x)) \right\|
-\end{equation}
-
+and \(y(q)\). To do this, we optimize the loss function $L$ (a
+distinct symbol from the Lagrangian $\mathcal{L}$),
+\begin{equation}
+L(x) = \left\| x-D(E(x)) \right\|_2^2
+\end{equation}
+
+\begin{figure}
+  \center
+\includegraphics[width=3in]{../figures/autoencoder_balance.png}
+  \caption{\label{fig:balance}Illustration of the insert of the encoder to form the
+  description of the system of equation. The equations
+  of motion are generated in a declarative programming methodology,
+  where the trained decoder can be extracted and inserted into the abstract
+  syntax tree, which can then be differentiated and inserted into a
+  time stepping loop.}
+\end{figure}
+
+The process of learning and generating the code is illustrated in
+Figure \ref{fig:balance}.
 When we want to solve the dynamics, we can treat the autoencoder the
-same as \(x(\theta)\), plugging in
-\[L(x(q),v(q,\dot{q}))\]
+same as \(x(\theta)\), plugging it into the equation for the Lagrangian,
+\begin{equation}
+  \mathcal{L}(x(q),v(q,\dot{q}))
+\end{equation}
 Taking the partial derivatives with respect to the feature variable and
 its rate, the components to the equation of motion are build the DAE we know from physics:
-\[\frac{\mathrm{d}}{\mathrm{d}t}\frac{\partial L}{\partial \dot{q}} = \frac{\partial L}{\partial q}\].
+\begin{equation}
+  \frac{\mathrm{d}}{\mathrm{d}t}\frac{\partial \mathcal{L}}{\partial \dot{q}} =
+  \frac{\partial \mathcal{L}}{\partial q}.
+\end{equation}
 By the chain rule, the velocities are related to the
@@
 method:
-\[\frac{\partial L}{\partial \dot{q}}_i - \Delta t \frac{\partial L}{\partial q}_i = \frac{\partial L}{\partial \dot{q}}_0 + \Delta t \frac{\partial L}{\partial q}_0\]
-
-Soltutions are shown in Figure \ref{fig:pendulum}.
+\begin{equation}
+\frac{\partial \mathcal{L}}{\partial \dot{q}}_i - \Delta t
+\frac{\partial \mathcal{L}}{\partial q}_i = \frac{\partial
+  \mathcal{L}}{\partial \dot{q}}_0 + \Delta t \frac{\partial
+  \mathcal{L}}{\partial q}_0
+\end{equation}
+
+Solutions are shown in Figure \ref{fig:pendulum}.
 \begin{figure}
@@
 (This is not true for training the entire deep network: it is only true for this linear regression step on the sub-network.)
-
-In future work, we would want to continue appending new \(q\)s in a
-systemic way to derive, e.g., an \(H_2O + NaCl\) equation of state with
-3 degrees of freedom, bootstraping from the pre-trained \(H_2O\)
-architecture with 2 degrees of freedom.
 
@@
 where \([\sigma]\) is the random initialization and \(…\) denotes
-padding the coefficients to the nonlinear terms by 0s. Recall that the
+padding the coefficients to the higher mononomial terms by zeros. Recall that the
 values for \(s\) have already been rescaled to be around 0 with a range
@@
 \centering
-\includegraphics{../slides/classifier_network.png}
+\includegraphics{../figures/classifier_network.png}
 \caption{\label{fig:classifyingnetwork}The architecture of the classifier network designed for
@@
 
+\subsection{Training Behavior}
+
+I need to show A) learning rate graphs of all of them
+B) describe the algorithm better. How many epochs?
+
+C) How do I quantify how well it classified?
+
+One of the surfaces colored by class is shown in
+\ref{fig:classified}. 
+Because the identity of the phases,XXXXXXXX
+For comparison purposes, we group together liquid-gas-supercritical as one contiguous label.
+
 \begin{figure}
@@
   \includegraphics[width=3in]{../slides/classifier_good.jpg}
-  \caption{\label{fig:classified}Result of the unsupervised phase classifier. Compare this plot to the original labeled dataset Figure \ref{fig:pTrho}.}
+  \caption{\label{fig:classified}Result of the unsupervised phase
+    classifier. Compare this plot to the original labeled dataset
+    Figure \ref{fig:pTrho}. Note the axes are in the
+    nondimensionalized scaling.}
 \end{figure}
-Because the identity of the phases,XXXXXXXX
-For comparison purposes, we group together liquid-gas-supercritical as one contiguous label.
 
@@
 \partial_t \rho(q_1,q_2) & = \mathbf{k} \left( p_\infty-p(q_1,q_2) \right)  + r \\
-\partial_t \left(\rho h(q_1,q_2)-p\right) & =\mathbf{k}_T(T_\infty-T(q_1,q_2))  + s + h\left(\mathbf{k}_p(p_\infty - p) + r\right)
+\partial_t \left(\rho h(q_1,q_2)-p\right) &
+                                            =\mathbf{k}_T(T_\infty-T(q_1,q_2))  + s + \frac{\rho h}{h}\left(\mathbf{k}_p(p_\infty - p) + r\right)
 \end{align}
@@
 evolved, even during intermediate iterative solution steps. Note again
-that the machine learning is only applied to the
-material representation: the balance laws are expressed directly.
+that the balance laws are directly incorporated into the simulation
+after loading the trained model.
 
@@
 \rho(q) &= \left(\rho_{max}-\rho_{min}\right) D_{\rho}(q) +\rho_{mean}\\
-h(q)      &= \left( h_{max}-h_{min}\right) D_h(q)  +h_{mean}
+\rho h(q)      &= \left(\rho h_{max}-\rho h_{min}\right) D_h(q)  + \rho h_{mean}
 \end{align}
@@
 Then, the simulator must determine $q(t=0)$ from these two
-values. This sets up a system of two equations.
-
-Note that the encoder only necessarily works when all four values are
-correct. Thus, it is necessary to actually solve the nonlinear problem
-on the decoder:
+values, but the econder only works when all four values are on the
+surface. Let $s=\{T,p,\rho,\rho h\}$ be a four dimensional
+variable of the stacked physical quantities. It is necessary to
+actually solve the following nonlinear problem
+on the decoder for the indices $i$ of the two specified values:
 \begin{equation}
@@
 e.g. only the rows associated with $p$ and $T$ are solved if those are
-specified, which is usually the case. The above equation is solved
-with Newton's method. The two corresponding rows are cut out of
+specified, which is usually the case. The resulting system of
+equations is solved with Newton's method. The two corresponding rows are cut out of
 the tangent matrix to $D$,
@@
 \end{equation}
-
-and the following system of equations is solved for an update to $q$,
+and the $2\times2$ linear system is solved for an update to $q$,
 \begin{equation}
@@
 \end{equation}
-with the under-relaxation parameter usually set to $\alpha \approx
+with the under-relaxation parameter that is usually set to $\alpha \approx
 0.1$.
@@
 \bar{\rho}_{gas}\\
-\bar{h}_{gas}
+\bar{\rho h}_{gas}
 \end{array}\right),\left(\begin{array}{c}
@@
 \bar{\rho}_{liq}\\
-\bar{h}_{liq}
+\bar{\rho h}_{liq}
 \end{array}\right),\left(\begin{array}{c}
@@
 \bar{\rho}_{ice}\\
-\bar{h}_{ice}
+\bar{\rho h}_{ice}
 \end{array}\right)... \right\}
@@
 \item
-  Make a database of \(T,p,\rho,h,X_1,...\)
+  Make a database of \(T,p,\rho,\rho h, ,...\)
   \begin{itemize}
@@
 
-\hypertarget{header-n3353}{%
-\subsection{Pseudo-code}\label{header-n3353}}
-
+\begin{algorithm}
+  \caption{Psuedo code of the simulation}
 \begin{verbatim}
+specify EOS and a which network
 load network from database
@@
 \end{verbatim}
+  \end{algorithm}
 
@@
 \end{table}
+
+The linear problem is only of interest for integration testing. Only one
+hyperparameter set for each architecture-generating class was trained
+on the dataset. The simulation code is then executed on the two linear
+problems, verifying the implementation of the multi-stage process and
+generation of the balance laws. In a fully-fledged simulator including flow, this
+dataset would yield a two-phase Darcy's law equation of state, which
+would allow for verifcation against that set of analytical solutions.
+
 The simulator was benchmarke against the long standing TOUGH+ multiphase
-multicomponent
-
-Single-grid block simulations were set up for most of the cases. TOUGH+
+multicomponent simulator. Single-grid block simulations were set up for most of the cases. TOUGH+
 does not handle supercritical \ch{H2O}, so these test problems were not
@@
 
+In future work, we would want to continue appending new \(q\)s in a
+systemic way to derive, e.g., an \(H_2O + NaCl\) equation of state with
+3 degrees of freedom, bootstraping from the pre-trained \(H_2O\)
+architecture with 2 degrees of freedom.
+
 \section*{Acknowlegements}

### 2019-07-11 autoencoder.py
bugfixes; new training

@@ -209,10 +209,11 @@ class PolyAutoencoder(Autoencoder):
         return tf.identity(q,name=name)
     
     def decode(self, q, name=None):
+        qpoly = atu.polyexpand(q, self.Np_dec)
         N_coeff = atu.Npolyexpand( self.size_q, self.Np_dec )
         We1 = self._var("dec_W", (N_coeff, self.size_x) )
         be1 = self._var("dec_b", (self.size_x,) )
-        x = tf.matmul( atu.polyexpand(q, self.Np_dec), We1 ) + be1
+        x = tf.matmul( atu.polyexpand(qpoly, self.Np_dec), We1 ) + be1
         return tf.identity(x,name=name)
     
     

### 2019-07-11 autoencoder_trainer.py

@@ -29,7 +29,7 @@ def AutoencoderFactory(hyper, outerdim, innerdim, stream_mini, stream_all):
                      *hyper['args'],
                      data_all=stream_all,
                      encoder_init=hyper['ini'],
-                     cae_lambda=hyper['cae'],)
+                     cae_lambda=hyper['cae'])
 
 
 class SaveAtEndHook(tf.train.SessionRunHook):

### 2019-07-11 training_slgc.py

@@ -1,11 +1,13 @@
 if __name__=="__main__":
-    training_dir = "/Users/afq/Google Drive/networks/"
+    # training_dir = "/Users/afq/Google Drive/networks/"
+    training_dir = "/Users/afq/Research/eoshub/networks/"
+
     if True: #imac
         data_dir = "/Users/afq/Dropbox/ML/primaryautoencoder/data_files/"
     else:
         data_dir = "/Users/afq/Documents/Dropbox/ML/primaryautoencoder/data_files/"
-    dataset = "water_iapws_slgc_logp_scaled"
-    eosname = "water_iapws_rh_slgc_logp"
+    dataset = "water_iapws_logp_scaled"
+    eosname = "water_iapws_slgc_logp"
 #     caes = [ 0.0, 1.0 ]
     inis = [ 'rand','pT', 'rhoh' ]
     caes = [ 0.0, ]
@@ -25,7 +27,7 @@ if __name__=="__main__":
         [2,6, 18,36,'sigmoid'],
         [2,6, 24,48,'sigmoid'],
     ]
-    beta_incs = [0.0,]
+    beta_incs = [0.0,0.05]
     archs_to_try = []
     for arch in class_archs:
         for beta_inc in beta_incs:
@@ -37,7 +39,7 @@ if __name__=="__main__":
     for arch in archs_to_try:
         sets_to_try.extend(extend_train_params(arch))
 
-    n_epoch = 5000
+    n_epoch = 10000
 
     import multiprocessing as multi
     import itertools

### 2019-07-11 batch_test_latent_sim.py
Factored out listing and loading of EOS and networks

@@ -120,10 +120,7 @@ class Cycle_sgclg():
         h  =iapws97.enthalpy_region2(400.0,5.0e3) )
     
 
-problems = dict(inspect.getmembers(sys.modules[__name__], inspect.isclass))
-
-
-hub = "/Users/afq/Google Drive/networks/"
+all_test_problems = dict(inspect.getmembers(sys.modules[__name__], inspect.isclass))
 
 import numpy as np
 import os
@@ -131,19 +128,13 @@ from latent_sim import LatentSim
 
 from SimDataDB import SimDataDB
 
-eoses = {
-    'water_slgc_logp_64':dict(
+eos_test_cfg = {
+    'water_slgc_logp':dict(
         scale_file = "data_files/water_iapw_logp_ranges.csv",
         logp=True,
-        problem_list=problems.keys() # TODO watch it!
+        problem_list=all_test_problems.keys() # TODO watch it!
     ),
-    'water_lg':dict(
-        scale_file = "data_files/surf_ranges.csv",
-        logp=False,
-        problem_list=['Small_Liquid','Small_Gas','Hot_Gas','Transition_L2G','Transition_L2G_Drain'
-                     'Liquid_Drain']
-    ),
-     'water_iapws_lg_logp':dict(
+    'water_iapws_lg_logp':dict(
         scale_file = "data_files/water_iapws_lg_ranges.csv",
         logp=False,
         problem_list=['Small_Liquid','Small_Gas','Hot_Gas','Transition_L2G',
@@ -204,6 +195,7 @@ def perform_tests_for_eos(eos, result_dir='.'):
                 solve_a_problem(p,n)
         except Exception as e:
             print("The network", n, " threw an error: ", e)
+            
 if __name__=="__main__":
     for k in eoses:
         try:

### 2019-07-11 dash_viz.py

@@ -1,7 +1,6 @@
 #!/usr/bin/env python3
 # -*- coding: utf-8 -*-
-from collections import namedtuple
-import os, glob
+
 
 import dash
 import dash_core_components as dcc
@@ -26,66 +25,11 @@ server = app.server
 app.config.suppress_callback_exceptions = True
 from dash.dependencies import Input, Output, State
 
+from eoshub import EOSHub
 import plotting_routines as ls_plot
 import grading_routines as ls_grade
 
-if 'DYNO' in os.environ:
-    app_name = os.environ['DASH_APP_NAME']
-else:
-    app_name = 'dash-text-annotationsplot'
-
-# Configure the eos hub
-hubs = [
-    "/Users/afq/Google Drive/networks/",
-    "/Users/afq/Documents/Research/LBNL/eoshub/eoshub/networks/",
-    "/Users/afq/Research/eoshub/networks/",
-    
-]
-
-class LazyLoad():
-    "Lazily load the state of all of the eoses"
-    Entry = namedtuple('Entry', ['archs', 'table','surfs','train_scores'])
-    def __init__(self,hubs):
-        self.eoses = []
-        self.cache = {} # Cache processing results
-        self.hubs = {} # Mapping from eos to where to find it
-        for hub in hubs:
-            eos_dirs = glob.glob(hub+'/training_*')
-            eoses = [ k[(len(hub)+len('training_')):] 
-                      for k in eos_dirs ]
-            self.eoses.extend(eoses)
-            for e in eoses:
-                self.hubs[e] = hub
-        print(self.hubs)
-        
-    def _eos_dir(self,eos):
-        return self.hubs[eos]+'/training_'+eos
-    
-    def get_it_all(self,eos):
-        "Load the data for a particular EOS into server memory"
-        directory = self._eos_dir(eos)
-        # Surfaces
-        surfs = ls_plot.read_networks(directory)
-        surfs.pop('.DS_Store',None) # lol
-        # Training and results 
-        
-        table, train_scores = ls_grade.prep_table(eos,self.hubs[eos])
-        all_archs = os.listdir(directory)
-        # TODO: Get the simulation results
-        # TODO: Run the simulations in batch
-        for row in table:
-            row["id"]=row["name"]
-        return self.Entry(all_archs, table, surfs, train_scores)
-    
-    def __getitem__(self, key):
-        try:
-            return self.cache[key]
-        except KeyError:
[…]

### 2019-07-11 eoshub.py

@@ -1,0 +1,56 @@
+from collections import namedtuple
+import os, glob
+
+import grading_routines as ls_grade
+
+# Configure the eos hub
+hubs = [
+    "/Users/afq/Google Drive/networks/",
+    "/Users/afq/Documents/Research/LBNL/eoshub/eoshub/networks/",
+    "/Users/afq/Research/eoshub/networks/",   
+]
+
+class LazyLoad():
+    "Lazily load the state of all of the eoses"
+    Entry = namedtuple('Entry', ['archs', 'table','surfs','train_scores'])
+    def __init__(self,hubs):
+        self.eoses = []
+        self.cache = {} # Cache processing results
+        self.hubs = {} # Mapping from eos to where to find it
+        for hub in hubs:
+            eos_dirs = glob.glob(hub+'/training_*')
+            eoses = [ k[(len(hub)+len('training_')):] 
+                      for k in eos_dirs ]
+            self.eoses.extend(eoses)
+            for e in eoses:
+                self.hubs[e] = hub
+        print(self.hubs)
+        
+    def _eos_dir(self,eos):
+        return self.hubs[eos]+'/training_'+eos
+    
+    def get_it_all(self,eos):
+        "Load the data for a particular EOS into server memory"
+        directory = self._eos_dir(eos)
+        # Surfaces
+        surfs = ls_plot.read_networks(directory)
+        surfs.pop('.DS_Store',None) # lol
+        # Training and results 
+        
+        table, train_scores = ls_grade.prep_table(eos,self.hubs[eos])
+        all_archs = os.listdir(directory)
+        # TODO: Get the simulation results
+        # TODO: Run the simulations in batch
+        for row in table:
+            row["id"]=row["name"]
+        return self.Entry(all_archs, table, surfs, train_scores)
+    
+    def __getitem__(self, key):
+        try:
+            return self.cache[key]
+        except KeyError:
+            rez = self.get_it_all(key)
+            self.cache[key] = rez
+            return rez
+        
+EOSHub = LazyLoad(hubs)

### 2019-07-12 dash_viz.py
added filtering

@@ -59,6 +59,7 @@ table = dash_table.DataTable(
     #data = None, #filled by callback
     row_selectable = "multi",
     sort_action="native",
+    filter_action="native",
     #sorting=True,
     #sorting_type="multi",
     selected_rows=[0,1,2,3])

### 2019-07-12 grading_routines.py

@@ -8,7 +8,7 @@ import numpy as np
 from tensorboard.backend.event_processing.event_accumulator import EventAccumulator
 from SimDataDB import SimDataDB
 
-from batch_test_latent_sim import problems as test_problems
+from batch_test_latent_sim import all_test_problems as test_problems
 
 
 

### 2019-07-12 paper/paper.tex
new more citations

@@
   in the Appendix. (Every point on this plot is a point in the
-  generated dataset used for the optimization)}
+  generated dataset used for the optimization.) An interactive version
+of this figure is included online.}
 \end{figure}
@@
 {\bf Remark:} There are even more known phases of crystal ice which for water
-For water, the supercritical phase and exotic ices only occur in
-laboratory conditions or naturally near the
-crust \cite{}. For our case, we only handle the Ice I and cutoff the
+For water, the supercritical phase only occurs at extreme conditions
+near the mantel or near magma chambers, which could be potential sites
+for geothermal plants\cite{elders_science_2010}. The exotic ices only occur in
+laboratory conditions\cite{water_revised_2009}. For our case, we only handle the Ice I and cutoff the
 regime at \si{10^8}{Pa}. For other materials these exotic phases can be
@@
 point at T=\si{-82.59}{C}\cite{the_engineering_toolbox_methane_2008}, so only an appropriate gas law can do the trick, until
-the bi-species hydrate phase comes into play.
-The liquid-gas-equilibrium-supercritical state transition comes into
-play in, for example, in geoengineered systems using $CO_2$
-such as geothermal plants are carbon sequestration facilities, the
+the bi-species hydrate phase comes into play. Handling the
+liquid-gas-equilibrium-supercritical state transition also comes into
+play with other fluids; for example, in geoengineered systems using
+$CO_2$ such as geothermal plants and carbon sequestration facilities, the
 working fluid is often in the supercritical regime at depth, and the
-phase boundary. cite
-curt. cite hydrates.
+phase boundary \cite{pruess_feasibility_2006,oldenburg_migration_2006}. 
 When binary mixtures are involved, the number of possible phases and
-equilibria increases again.
+equilibria increases again; methane hydrate simulators have 15
+states \cite{moridis_simulation_2019}.
 

### 2019-08-06 batch_test_latent_sim.py
refactoring cyclic dependency; new script for tests

@@ -1,156 +1,13 @@
 from __future__ import print_function
-import sys,inspect
-from equations_of_state import iapws97
-
-class Linear_Liquid():
-    t_max = 1.0
-    initial = dict(p = 1.0e5, T = 20.0+273.15, phase="Liquid")
-    params =  dict(k_p=1.0e-4,k_T=1.0e4)
-    @staticmethod
-    def schedule(sim,t):
-        sim.set_params(T_inf=293.15,p_inf=1.0e5+10.0)
-    answer = dict(T=293.15,p=1.0e5+10.0,
-        rho=iapws97.density_region1(293.15,1.0e5+10.0),
-        h  =iapws97.enthalpy_region1(293.15,1.0e5+10.0) )
-    
-class Linear_Liquid_Heat():
-    t_max = 10.0
-    initial = dict(p = 1.0e5, T = 20.0+273.15, phase="Liquid")
-    params =  dict(k_p=1.0e-4,k_T=1.0e6)
-    @staticmethod
-    def schedule(sim,t):
-        sim.set_params(T_inf=294.15,p_inf=1.0e5)
-    answer = dict(T=294.15,p=1.0e5,
-        rho=iapws97.density_region1(294.15,1.0e5),
-        h  =iapws97.enthalpy_region1(294.15,1.0e5) )
-    
-class Small_Liquid():
-    t_max = 1000.0
-    initial = dict(T=350,p=5.0e5, phase="Liquid")
-    params =  dict(k_p=1.0e-4,k_T=1.0e4,Dt=10.0)
-    @staticmethod
-    def schedule(sim,t):
-        sim.set_params(T_inf=450,p_inf=5.0e5)
-    answer = dict(T=450,p=5.0e5,
-        rho=iapws97.density_region1(450.0,5.0e5),
-        h  =iapws97.enthalpy_region1(450.0,5.0e5) )
-
-class Small_Gas():
-    t_max = 1000.0
-    initial = dict(T=350,p=5.0e3, phase="Gas")
-    params =  dict(k_p=1.0e-4,k_T=-1.0e4,Dt=10.0)
-    @staticmethod
-    def schedule(sim,t):
-        sim.set_params(T_inf=450,p_inf=5.0e3)
-    answer = dict(T=450,p=5.0e3,
-        rho= iapws97.density_region2(450.0,5.0e3),
-        h  =iapws97.enthalpy_region2(450.0,5.0e3) )
-
[…]

### 2019-08-06 directory_parsing.py

@@ -1,0 +1,25 @@
+import glob, re
+# Configure the eos hub. 
+# This is factored out to not do imports before forking.
+hubs = [
+    "/Users/afq/Google Drive/networks/",
+    "/Users/afq/Documents/Research/LBNL/eoshub/eoshub/networks/",
+    "/Users/afq/Research/eoshub/networks/",   
+]
+
+test_dir = "/Users/afq/Research/eoshub/test_db/"
+
+def get_list_of_eoses():
+    eoses = []
+    for hub in hubs:
+        eos_dirs = glob.glob(hub+'/training_*')
+        hub_eoses = [ k[(len(hub)+len('training_')):] 
+                      for k in eos_dirs ]
+        eoses.extend(hub_eoses)
+    return eoses
+
+def list_files(fpattern):
+    files = glob.glob(fpattern)
+    grab_digit = lambda f : int(re.search("([0-9]*)\.[a-zA-Z]*$",f).groups()[-1])
+    files.sort(key=lambda f: grab_digit(f) )
+    return files

### 2019-08-06 eoshub.py

@@ -1,19 +1,12 @@
 from collections import namedtuple
 import os, glob
 
+from directory_parsing import hubs
+
 import grading_routines as ls_grade
 import plotting_routines as ls_plot
-import batch_test_latent_sim as ls_test
+import test_cfg
 from latent_sim import LatentSim
-
-# Configure the eos hub
-hubs = [
-    "/Users/afq/Google Drive/networks/",
-    "/Users/afq/Documents/Research/LBNL/eoshub/eoshub/networks/",
-    "/Users/afq/Research/eoshub/networks/",   
-]
-
-
 
 class LazyLoad():
     "Lazily load the state of all of the eoses"
@@ -69,10 +62,10 @@ class LazyLoad():
         try:
             ls = self.sims[(eos,network)]
         except KeyError:
-            for dataset_match in ls_test.eos_test_cfg:
+            for dataset_match in test_cfg.eos_test_cfg:
                 if dataset_match in eos:
-                    scale_file = ls_test.eos_test_cfg[dataset_match]['scale_file']
-                    logp = ls_test.eos_test_cfg[dataset_match]['logp']
+                    scale_file = test_cfg.eos_test_cfg[dataset_match]['scale_file']
+                    logp = test_cfg.eos_test_cfg[dataset_match]['logp']
                     break
             ls = LatentSim(self._eos_dir(eos)+'/'+network,scale_file,logp)
             self.sims[(eos,network)] = ls

### 2019-08-06 grading_routines.py

@@ -8,7 +8,7 @@ import numpy as np
 from tensorboard.backend.event_processing.event_accumulator import EventAccumulator
 from SimDataDB import SimDataDB
 
-from batch_test_latent_sim import all_test_problems as test_problems
+from test_cfg import all_test_problems as test_problems
 
 
 

### 2019-08-06 plotting_routines.py

@@ -17,11 +17,7 @@ color_ring = itertools.cycle(['black','orange','purple','red','blue','green'])
 colorkey = defaultdict(lambda : next(color_ring)) # It should be global, no?
 d_off = 2.0
 
-def list_files(fpattern):
-    files = glob.glob(fpattern)
-    grab_digit = lambda f : int(re.search("([0-9]*)\.[a-zA-Z]*$",f).groups()[-1])
-    files.sort(key=lambda f: grab_digit(f) )
-    return files
+from directory_parsing import list_files
 
 #
 # Plotting of all of the scores
@@ -137,7 +133,7 @@ def plot_networks(surfaces,aspectratio=1, z='rho'):
 #
 # Simulations
 #
-def plotly_simulation_t(sims,ref=None, showleg = False):
+def plotly_simulation_t(sims,ref=None, showlegend=False):
     legends=['T','p','rho','rho*h']
     traces = []
     for i,name in enumerate(legends):
@@ -145,19 +141,19 @@ def plotly_simulation_t(sims,ref=None, showleg = False):
             trace_ref = go.Scatter(x=ref[:,0],
                                   y=ref[:,i+1],
                                   name='ref',legendgroup='ref',
-                                  showlegend= (i==0 and showleg),
+                                  showlegend= (i==0 and showlegend),
                                   line=dict(dash='dash', color=colorkey['ref']))
             traces.append((trace_ref,i+1))
         for n,time_series in sims.items():
             trace = go.Scatter(x=time_series[:,0],y=time_series[:,i+3],
                                name=n,legendgroup=n,
                                line=dict(color=colorkey[n]),
-                               showlegend=(i==0 and showleg))
+                               showlegend=(i==0 and showlegend))
             traces.append((trace,i+1))
     return traces
 
 
-def plotly_simulations_Tp(sims,ref=None):
+def plotly_simulations_Tp(sims,ref=None,showlegend=True):
     from equations_of_state.iapws_boundaries \
         import plot_boundaries_plotly
     traces_bound = plot_boundaries_plotly()
@@ -167,14 +163,14 @@ def plotly_simulations_Tp(sims,ref=None):
         trace_ref = go.Scatter(x=ref[:,1],
                        y=ref[:,2],
                        name='ref',legendgroup='ref',
-                       showlegend=True,
[…]

### 2019-08-06 run_batch_tests.py

@@ -1,0 +1,14 @@
+import test_cfg
+import directory_parsing
+
+def job(eos):
+    import batch_test_latent_sim as bt
+    
+    cfg = test_cfg.find_eos_test_cfg(eos)
+    for problem in cfg['problem_list']:
+        bt.solve_a_problem(problem, eos, result_dir=directory_parsing.test_dir)
+    
+if __name__=='__main__':
+    # Get a list of eoses in the hubs without importing anything
+    #eoses = list_eos_directories.get_list_of_eoses()
+    job('water_iapws_slgc_logp')

### 2019-08-06 test_cfg.py

@@ -1,0 +1,151 @@
+import sys, inspect
+from equations_of_state import iapws97
+
+# TODO the meaning of the "answers" has changed now
+
+class Linear_Liquid():
+    t_max = 1.0
+    initial = dict(p = 1.0e5, T = 20.0+273.15, phase="Liquid")
+    params =  dict(k_p=1.0e-4,k_T=1.0e4)
+    @staticmethod
+    def schedule(sim,t):
+        sim.set_params(T_inf=293.15,p_inf=1.0e5+10.0)
+    answer = dict(T=293.15,p=1.0e5+10.0,
+        rho=iapws97.density_region1(293.15,1.0e5+10.0),
+        h  =iapws97.enthalpy_region1(293.15,1.0e5+10.0) )
+    
+class Linear_Liquid_Heat():
+    t_max = 10.0
+    initial = dict(p = 1.0e5, T = 20.0+273.15, phase="Liquid")
+    params =  dict(k_p=1.0e-4,k_T=1.0e6)
+    @staticmethod
+    def schedule(sim,t):
+        sim.set_params(T_inf=294.15,p_inf=1.0e5)
+    answer = dict(T=294.15,p=1.0e5,
+        rho=iapws97.density_region1(294.15,1.0e5),
+        h  =iapws97.enthalpy_region1(294.15,1.0e5) )
+    
+class Small_Liquid():
+    t_max = 1000.0
+    initial = dict(T=350,p=5.0e5, phase="Liquid")
+    params =  dict(k_p=1.0e-4,k_T=1.0e4,Dt=10.0)
+    @staticmethod
+    def schedule(sim,t):
+        sim.set_params(T_inf=450,p_inf=5.0e5)
+    answer = dict(T=450,p=5.0e5,
+        rho=iapws97.density_region1(450.0,5.0e5),
+        h  =iapws97.enthalpy_region1(450.0,5.0e5) )
+
+class Small_Gas():
+    t_max = 1000.0
+    initial = dict(T=350,p=5.0e3, phase="Gas")
+    params =  dict(k_p=1.0e-4,k_T=-1.0e4,Dt=10.0)
+    @staticmethod
+    def schedule(sim,t):
+        sim.set_params(T_inf=450,p_inf=5.0e3)
+    answer = dict(T=450,p=5.0e3,
+        rho= iapws97.density_region2(450.0,5.0e3),
+        h  =iapws97.enthalpy_region2(450.0,5.0e3) )
+
+class Hot_Gas():
+    t_max = 1000.0
[…]

### 2019-08-06 autoencoder_trainer.py
writing a trimmed meta graph; didn't test this yet

@@ -33,12 +33,18 @@ def AutoencoderFactory(hyper, outerdim, innerdim, stream_mini, stream_all):
 
 
 class SaveAtEndHook(tf.train.SessionRunHook):
-    def __init__(self,fname):
+    def __init__(self,fname, graph):
         self.fname = fname
+        self.graph = graph
     def begin(self):
         self._saver = tf.train.Saver()
     def end(self, session):
-        # TODO freeze it
+        atu.write_trimmed_pb_graph(self.graph,session,
+                              ["decode","encode"],
+                              self.fname+"_frz")
+        atu.write_trimmed_meta_graph(self.graph,session,
+                              ["decode","encode"],
+                              self.fname+"_frz_meta")
         self._saver.save(session, self.fname)
         
 class DoStuffHook(tf.train.SessionRunHook):
@@ -93,7 +99,7 @@ def train_autoencoder(name, dataname, outerdim, innerdim,
         def printgoalhook(ctx,run_values):
             print(ctx.sessiormn.run(ae.goal_all))
         stophook = tf.train.StopAtStepHook(last_step=n_epoch)
-        saverhook = SaveAtEndHook(training_dir+"/final_variables")
+        saverhook = SaveAtEndHook(training_dir+"/final_variables",graph)
         # Make a closure into a hook
         @DoStuffHook(freq=image_freq)
         def extrahook(ctx,run_values):

### 2019-08-06 training_water_iapws_slgc.py

@@ -39,7 +39,7 @@ if __name__=="__main__":
     for arch in archs_to_try:
         sets_to_try.extend(extend_train_params(arch))
 
-    n_epoch = 10000
+    n_epoch = 16001
 
     import multiprocessing as multi
     import itertools

### 2019-08-06 dash_viz.py
Added a radio button for simulation ; does nothing yet

@@ -139,6 +139,11 @@ def update_graph(eos,selected,radio):
         lines_to_plot = {k:loaded[eos].train_scores[k] for k in selected 
                          if k in loaded[eos].train_scores.keys() }
         figure = ls_plot.make_training_plot(lines_to_plot)
+    elif radio == 'simulation':
+        lines_to_plot = {k:loaded[eos].train_scores[k] for k in selected 
+                         if k in loaded[eos].train_scores.keys() }
+        figure = ls_plot.make_training_plot(lines_to_plot)
+        
     else:
         figure = ls_plot.make_training_plot(loaded[eos].train_scores)
     return figure

### 2019-08-06 paper/paper.tex
paper edits

@@
 simulation itself onto the computer: tens of thousands of lines of human
-written code consisting of empirical fits and branching logic are
+written code consisting of theoretical relations, empirical fits, and branching logic are
 replaced by code generated by the learned model. The methodology will

### 2019-08-07 autoencoder_trainer.py
loading from the frozen protobuf

@@ -41,10 +41,10 @@ class SaveAtEndHook(tf.train.SessionRunHook):
     def end(self, session):
         atu.write_trimmed_pb_graph(self.graph,session,
                               ["decode","encode"],
-                              self.fname+"_frz")
+                              self.fname+"_frz.pb")
         atu.write_trimmed_meta_graph(self.graph,session,
                               ["decode","encode"],
-                              self.fname+"_frz_meta")
+                              self.fname+"_frz.meta")
         self._saver.save(session, self.fname)
         
 class DoStuffHook(tf.train.SessionRunHook):

### 2019-08-07 latent_sim.py

@@ -3,6 +3,8 @@ import numpy as np
 import tensorflow as tf
 import afqstensorutils as atu
 
+LOAD_PROTOBUF = True
+
 def unshift(x, scale):
     return (scale[2,:]-scale[1,:])*x + scale[0,:]
 def shift(x,scale):
@@ -23,16 +25,24 @@ class LatentSim():
         """Read the autoencoder from a tensorflow file"""
         self._graph = tf.Graph()
         self._sess = tf.Session(graph=self._graph)
-#         with self._sess as sess:
-        with self._graph.as_default():
-            saver = tf.train.import_meta_graph(location+'/final_graph.meta')
-        saver.restore(self._sess,location+'/final_variables')
+        if LOAD_PROTOBUF:
+            with tf.gfile.GFile(location+'/final_variables_frz','rb') as f:
+                gdef = tf.GraphDef()
+                gdef.ParseFromString(f.read())
+            with self._graph.as_default():
+                tf.import_graph_def(gdef)
+            pfx='import/'
+        else:
+            with self._graph.as_default():
+                saver = tf.train.import_meta_graph(location+'/final_variables.meta')
+            saver.restore(self._sess,location+'/final_variables')
+            pfx=''
         # hooks to the decoder
-        self.i_q = self._graph.get_tensor_by_name('i_q:0') 
-        self.o_x = self._graph.get_tensor_by_name('decode:0')
+        self.i_q = self._graph.get_tensor_by_name(pfx+'i_q:0') 
+        self.o_x = self._graph.get_tensor_by_name(pfx+'decode:0')
         # hooks to the encoder
-        self.i_x = self._graph.get_tensor_by_name('i_x:0')
-        self.o_q = self._graph.get_tensor_by_name('encode:0')
+        self.i_x = self._graph.get_tensor_by_name(pfx+'i_x:0')
+        self.o_q = self._graph.get_tensor_by_name(pfx+'encode:0')
         # The simulation inputs and outputs in the unscaled space
         # (the prior conditioning of the data put it on [-1,1])
         if scale_file:

### 2019-08-07 plotting_routines.py
Parsing and plotting everything

@@ -178,7 +178,7 @@ def plotly_simulations_Tp(sims,ref=None,showlegend=True):
     return traces, layout
 
 
-def plotly_simulations_Tp_ts(sims,ref,showlegend=True):
+def plotly_simulations_Tp_ts(sims,ref=None,showlegend=True):
     trace_Tp, layout_Tp = plotly_simulations_Tp(sims,ref,showlegend=False)
     traces_ts = plotly_simulation_t(sims,ref,showlegend=showlegend)
     subfig = tools.make_subplots(rows=4,cols=2,

### 2019-08-21 paper/paper.tex
Fit it to the ijnme format

@@
 %
-\documentclass[]{article}
-\usepackage[margin=1in]{geometry}
-\usepackage{amssymb,amsmath}
+\documentclass[AMA,STIX1COL]{WileyNJD-v2}
+% \usepackage[margin=1in]{geometry}
+%\usepackage[numbers]{natbib}
+
+\usepackage{amsmath}
+%amssymb
 \usepackage{chemmacros}
@@
 \usepackage{algorithm}
-\usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\usepackage[numbers]{natbib}
+% \usepackage{lmodern}
+% \usepackage{ifxetex,ifluatex}
 \usepackage{lineno}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
+% \usepackage{fixltx2e} % provides \textsubscript
+% \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+%   \usepackage[T1]{fontenc}
+%   \usepackage[utf8]{inputenc}
+%   \usepackage{textcomp} % provides euro and other symbols
+% \else % if luatex or xelatex
+%   \usepackage{unicode-math}
+%   \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+% \fi
 % use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 % use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% \IfFileExists{microtype.sty}{%
+% \usepackage[]{microtype}
+% \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+% }{}
+% \IfFileExists{parskip.sty}{%
+% \usepackage{parskip}
+% }{% else
+% \setlength{\parindent}{0pt}
+% \setlength{\parskip}{6pt plus 2pt minus 1pt}
+% }
 \usepackage{hyperref}
@@
 \usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
+% \usepackage{fancyvrb}
+% \newcommand{\VerbBar}{|}
+% \newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+% \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% % Add ',fontsize=\small' for more characters per line
+% \newenvironment{Shaded}{}{}
+% \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
+% \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
+% \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
+% \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
+% \newcommand{\BuiltInTok}[1]{#1}
+% \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
+% \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
+% \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
+% \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
+% \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
+% \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
+% \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
+% \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
+% \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
+% \newcommand{\ExtensionTok}[1]{#1}
+% \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
+% \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
+% \newcommand{\ImportTok}[1]{#1}
+% \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
+% \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
+% \newcommand{\NormalTok}[1]{#1}
+% \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
+% \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
+% \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
+% \newcommand{\RegionMarkerTok}[1]{#1}
+% \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
+% \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
+% \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
+% \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
+% \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
+% \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
 \usepackage{graphicx,grffile}
@@
 on the Latent Space of an Autoencoder}
-\author{Alejandro Francisco Queiruga}
-\date{\today}
+\author[1]{Alejandro Francisco Queiruga*}
+\address[1]{
+  \orgdiv{Energy Geosciences Division}, \orgname{Lawrence Berkeley
+    National Laboratory}, \orgaddress{\state{CA}}, \country{USA}}
+\corres{*\email{afqueiruga@lbl.gov}}
+
+\received{\today}
+\revised{\today}
+
+\jnlcitation{\cname{Queiruga, A.}\ctitle{Solution of DAEs on Latent Spaces},\cjournal{I. J. Num. Meth. Eng.}, \cvol{2019;00:1--2}}
 
 \begin{document}
-\maketitle
-\begin{abstract}
+
+\abstract[Summary]{
   Multiphase simulation is hard.
-
 Our proposed method is motivated by the problems of multiphase reaction and
@@
 properties which are difficult to describe.
-
 Our proposed approach is to build upon modern machine learning
@@
 We illustrate the methodology on two traditional problems: 1) the
-pendulum, and 2) the continuous stirred tank reactor.
-
-\end{abstract}
+pendulum, and 2) the continuous stirred tank reactor.}
+
+\maketitle
 
@@
   \center
-\includegraphics[width=3in]{../figures/autoencoder_balance.png}
+\includegraphics[width=3in]{../figures/autoencoder_balance_detailed.png}
   \caption{\label{fig:balance}Illustration of the insert of the encoder to form the

### 2019-08-23 autoencoder.py
Bug in polyautoencoder

@@ -202,8 +202,11 @@ class PolyAutoencoder(Autoencoder):
         
     def encode(self, x, name=None):
         N_coeff = atu.Npolyexpand( self.size_x, self.Np_enc )
-        We1 = self._var("enc_W", (N_coeff, self.size_q),
-                        initial_value = encoder_init_options[self.encoder_init])
+        if self.encoder_init in encoder_init_options.keys():
+            We1 = self._var("enc_W", (N_coeff, self.size_q),
+                            initial_value = encoder_init_options[self.encoder_init])
+        else:
+            We1 = self._var("enc_W", (N_coeff, self.size_q) )
         be1 = self._var("enc_b", (self.size_q,) )
         q = tf.matmul( atu.polyexpand(x, self.Np_enc), We1 ) + be1
         return tf.identity(q,name=name)
@@ -213,7 +216,7 @@ class PolyAutoencoder(Autoencoder):
         N_coeff = atu.Npolyexpand( self.size_q, self.Np_dec )
         We1 = self._var("dec_W", (N_coeff, self.size_x) )
         be1 = self._var("dec_b", (self.size_x,) )
-        x = tf.matmul( atu.polyexpand(qpoly, self.Np_dec), We1 ) + be1
+        x = tf.matmul( qpoly, We1 ) + be1
         return tf.identity(x,name=name)
     
     

### 2019-08-23 paper/paper.tex
Updated the paper; added the abstract I submitted to agu

@@
 \author[1]{Alejandro Francisco Queiruga*}
-\address[1]{
-  \orgdiv{Energy Geosciences Division}, \orgname{Lawrence Berkeley
+\address[1]{\orgdiv{Energy Geosciences Division}, \orgname{Lawrence Berkeley
     National Laboratory}, \orgaddress{\state{CA}}, \country{USA}}
@@
 
-\jnlcitation{\cname{Queiruga, A.}\ctitle{Solution of DAEs on Latent Spaces},\cjournal{I. J. Num. Meth. Eng.}, \cvol{2019;00:1--2}}
-
 \begin{document}
 
-\abstract[Summary]{
+\abstract[Abstract]{
   Multiphase simulation is hard.
@@
 pendulum, and 2) the continuous stirred tank reactor.}
+
+\jnlcitation{\cname{Queiruga, A.}\ctitle{Solution of DAEs on Latent Spaces},\cjournal{I. J. Num. Meth. Eng.}, \cvol{2019;00:1--2}}
+
+\keywords{Class file; \LaTeXe; \emph{Wiley NJD}}
 
@@
 \begin{algorithm}
-  \caption{\label{alg:statemachine}Statemachine transition logic}
-\begin{Shaded}
-\begin{Highlighting}[]
-\ControlFlowTok{switch}\NormalTok{ phase_old:}
-  \ControlFlowTok{case}\NormalTok{ gas:}
-    \ControlFlowTok{if}\NormalTok{ p,T crossing boundary:}
-\NormalTok{      phase_new = liquid_gas}
-  \ControlFlowTok{case}\NormalTok{ liquid:}
-    \ControlFlowTok{if}\NormalTok{ p,T crossing boundary:}
-\NormalTok{      phase_new = liquid_gas}
-  \ControlFlowTok{case}\NormalTok{ liquid_gas:}
-    \ControlFlowTok{if}\NormalTok{ S_gas >= }\DecValTok{1}\NormalTok{:}
-\NormalTok{      phase_new = gas}
-    \ControlFlowTok{if}\NormalTok{ S_liquid >= }\DecValTok{1}\NormalTok{:}
-\NormalTok{      phase_new = gas}
-\end{Highlighting}
-\end{Shaded}
+\begin{lstlisting}[caption={Finite State Machine transition logic}, label={alg:statemachine}]
+switch phase_old:
+  case gas:
+    if p,T crossing boundary:
+      phase_new = liquid_gas
+  case liquid:
+    if p,T crossing boundary:
+      phase_new = liquid_gas
+  case liquid_gas:
+    if S_gas >= 1:
+      phase_new = gas
+    if S_liquid >= 1:
+      phase_new = gas
+\end{lstlisting}
 \end{algorithm}
+%  \begin{verbatim}
+
+%\end{verbatim}
 
@@
 \section*{Acknowlegements}
+%\acks
 

### 2019-08-27 latent_sim.py
I want to see what k is

@@ -189,7 +189,7 @@ class LatentSim():
                 'Trap':0.5}[method]
         with self._graph.as_default():
             aii = self.regvar('aii',aii)
-            Dt  = self.regvar('Dt',0.1)
+            Dt  = self.regvar('Dt',1.0)
 
             T,p,rho,rho_h = tf.split(self.o_s,4,axis=-1)
             m,r = self.m_and_r( T,p,rho,rho_h )
@@ -237,7 +237,7 @@ class LatentSim():
         regard."""
         qi = q0.copy()
         rhs_0 = self._sess.run(self.rhs,feed_dict={self.i_q:q0})
-        for k in range(500):
+        for k in range(100):
             K_k,lhs_k = self._sess.run([self.K_lhs,self.lhs], feed_dict ={self.i_q:qi})
             R = rhs_0 - lhs_k
             Dq = np.linalg.solve(K_k[0,:,:],R[0,:])
@@ -278,6 +278,6 @@ class LatentSim():
                 break
             if verbose:
                 print(t,k,nDq)
-            timeseries.append(np.c_[t,q,self.decode(q)])
+            timeseries.append(np.c_[t,q,self.decode(q),k])
         return np.vstack(timeseries)
         

### 2019-08-27 latent_sim.py
A flux function

@@ -227,6 +227,35 @@ class LatentSim():
             rho_h/rho*mflux + k_T*(T_inf-T)+heat_source],axis=-1)
         return m, rate
     
+    def build_flux(self):
+        self.i_XA = tf.placeholder(name='i_XA',shape=(None,2),dtype=self.dtype)
+        self.i_XB = tf.placeholder(name='i_XB',shape=(None,2),dtype=self.dtype)
+        self.i_q2 = tf.placeholder(name='i_q2',shape=(None,2),dtype=self.dtype)
+        self.o_s2 = replicate_subgraph(self.o_s,{self.i_q:i_q2})
+        TA,pA,rhoA,rho_hA = tf.split(self.o_s,4,axis=-1)
+        TB,pB,rhoB,rho_hB = tf.split(self.o_s2,4,axis=-1)
+        F = self.flux(TA,pA,rhoA,rho_hA,
+                      TB,pB,rhoB,rho_hB,
+                     self.i_XA,self.i_XB)
+        return F
+    
+    def flux(self,TA,pA,rhoA,rho_hA,
+                  TB,pB,rhoB,rho_hB,
+                  XA,XB):
+        k_p =   self.regvar( "k_p", 10.0)
+        k_T =   self.regvar( "k_T", 10.0)
+        g = self.regvar("g",[0,-9.81])
+        L = tf.norm(XA-XB)
+        mflux = k_p * (pB - pA) / L - tf.einsum('ij,ij->i',g,(XB-XA)/L)
+        qflux = k_T * (TB - TA) / L
+        F = tf.concat([
+            mflux,
+            qflux + mflux*rho_hbar/rhobar,
+            - mflux,
+            - (qflux + mflux*rho_hbar/rhobar),
+        ])
+        return F
+    
     def set_params(self, **kwargs):
         """Set a value k = v in the current session."""
         for k,v in kwargs.items():

### 2019-08-28 latent_sim.py
fixed the flux graph

@@ -10,6 +10,8 @@ def unshift(x, scale):
 def shift(x,scale):
     return (x-scale[0,:])/(scale[2,:]-scale[1,:])
 
+
+
 class LatentSim():
     """Object for storing a simulation state"""
     def __init__(self, model_loc,scale_file="", logp=True, method="BWEuler"):
@@ -228,32 +230,38 @@ class LatentSim():
         return m, rate
     
     def build_flux(self):
-        self.i_XA = tf.placeholder(name='i_XA',shape=(None,2),dtype=self.dtype)
-        self.i_XB = tf.placeholder(name='i_XB',shape=(None,2),dtype=self.dtype)
-        self.i_q2 = tf.placeholder(name='i_q2',shape=(None,2),dtype=self.dtype)
-        self.o_s2 = replicate_subgraph(self.o_s,{self.i_q:i_q2})
-        TA,pA,rhoA,rho_hA = tf.split(self.o_s,4,axis=-1)
-        TB,pB,rhoB,rho_hB = tf.split(self.o_s2,4,axis=-1)
-        F = self.flux(TA,pA,rhoA,rho_hA,
-                      TB,pB,rhoB,rho_hB,
-                     self.i_XA,self.i_XB)
-        return F
-    
+        with self._graph.as_default():
+            self.i_XA = tf.placeholder(name='i_XA',shape=(None,2),dtype=self.dtype)
+            self.i_XB = tf.placeholder(name='i_XB',shape=(None,2),dtype=self.dtype)
+            self.i_q2 = tf.placeholder(name='i_q2',shape=(None,2),dtype=self.dtype)
+            self.o_s2 = atu.replicate_subgraph(self.o_s,{self.i_q:self.i_q2})
+            TA,pA,rhoA,rho_hA = tf.split(self.o_s,4,axis=-1)
+            TB,pB,rhoB,rho_hB = tf.split(self.o_s2,4,axis=-1)
+            self.o_F = self.flux(TA,pA,rhoA,rho_hA,
+                          TB,pB,rhoB,rho_hB,
+                          self.i_XA,self.i_XB)
+
+            #self.o_KF = atu.vector_gradient_dep(self.o_F,tf.stack([self.i_q,self.i_q2]))
+        self._sess.run(tf.variables_initializer(self._vars.values()))
+        
     def flux(self,TA,pA,rhoA,rho_hA,
[…]

### 2019-08-30 latent_sim.py
Assembling the fluxes

@@ -240,8 +240,9 @@ class LatentSim():
             self.o_F = self.flux(TA,pA,rhoA,rho_hA,
                           TB,pB,rhoB,rho_hB,
                           self.i_XA,self.i_XB)
-
-            #self.o_KF = atu.vector_gradient_dep(self.o_F,tf.stack([self.i_q,self.i_q2]))
+            
+            self.o_KF = tf.concat([atu.vector_gradient_dep(self.o_F,self.i_q),
+                                   atu.vector_gradient_dep(self.o_F,self.i_q2)],axis=-1)
         self._sess.run(tf.variables_initializer(self._vars.values()))
         
     def flux(self,TA,pA,rhoA,rho_hA,
@@ -252,7 +253,8 @@ class LatentSim():
         g = self.regvar("g",[0,-9.81])
         L = tf.norm(XA-XB,axis=-1)
         n = tf.einsum('ij,i->ij',(XB-XA),(1.0/L))
-        mflux = k_p * ( tf.einsum('ij,i->ij',(pB - pA),(1.0/L) )  - tf.expand_dims(tf.einsum('j,ij->i',g, n),axis=-1 ) )
+        grav_forcing = tf.expand_dims(tf.einsum('j,ij->i',g, n),axis=-1 )
+        mflux = k_p * ( tf.einsum('ij,i->ij',(pB - pA),(1.0/L) )  - grav_forcing )
         qflux = k_T * ( tf.einsum('ij,i->ij',(pB - pA),(1.0/L) ) )
         rhobar = (rhoB + rhoA)/2.0
         rho_hbar = (rho_hB + rho_hA)/2.0

### 2019-11-28 LatentFlow.py
There wasn't actually a bug.

@@ -1,5 +1,6 @@
 import numpy as np
 import scipy.sparse.linalg
+from matplotlib import pylab as plt
 
 import cornflakes as cf
 import husk_identity
@@ -12,12 +13,12 @@ class LatentFlow(LatentSim):
     def __init__(self, model_loc, scale_file="", logp=True, method="BWEuler"):
         super(LatentFlow,self).__init__(model_loc,scale_file,logp,method)
         self.build_flux()
-        self.mesh = None
+        # self.mesh = None
         
     def make_line_mesh(self,Ny,L):
         """Makes a new mesh of elements in a line."""
         # The centers of the volumes
-        self.X = np.c_[np.zeros((Ny,)),np.linspace(0,Ny,L)]
+        self.X = np.c_[np.zeros((Ny,)),np.linspace(0,L,Ny)]
         # A graph of the volumes
         self.H_vol = cf.Hypergraph()
         for i in range(Ny):
@@ -45,47 +46,49 @@ class LatentFlow(LatentSim):
             plt.xlabel(leg)
     
     def set_uniform(self,T=None, p=None, rho=None, rho_h=None):
+        q = np.zeros(2*self.X.shape[0])
         q0 = self.find_point(T=T,p=p,rho=rho,rho_h=rho_h)
         for e in self.H_vol:
             q[self.dm_q.Get_List(e)] = q0
-        q0 = q.copy()
-        
+        return q
+    
     def build_system(self,q,q0):
         """Build the matrix system of equations for the next Newton step."""
-        DT = self._sess.run(ls._vars['Dt'])
+        DT = self._sess.run(self._vars['Dt'])
         R0_vol_arr = self._sess.run([self.rhs], feed_dict ={self.i_q:q0.reshape(-1,2)})
         # The mass component at t
         vol_R0, = cf.Assemble(husk_identity.kernel_idty_R,
                                 self.H_vol,
                                 {'iR':(R0_vol_arr,self.dm_q),},
                                 {'R':(self.dm_q,),},
-                                ndof=self.X.shape[1]*2)
+                                ndof=self.X.shape[0]*2)
         # Assemble the mass component at t+Dt:
-        K_vol_arr,R_vol_arr = ls._sess.run([self.K_lhs,self.lhs],
+        K_vol_arr,R_vol_arr = self._sess.run([self.K_lhs,self.lhs],
                                            feed_dict ={self.i_q:q.reshape(-1,2)})
         vol_R,vol_K = cf.Assemble(husk_identity.kernel_idty_RK,
[…]

### 2019-11-28 eoshub.py
bug in refactorization of the hub

@@ -77,7 +77,7 @@ class LazyLoad():
         try:
             ls = self.sims[(eos,network)]
         except KeyError:
-            args = eos_loadout(eos, network)
+            args = self.eos_loadout(eos, network)
             ls = LatentSim(*args)
             self.sims[(eos,network)] = ls
         return ls

### 2019-11-28 latent_sim.py

@@ -230,6 +230,7 @@ class LatentSim():
         return m, rate
     
     def build_flux(self):
+        """This should really go into the child LatentFlow."""
         with self._graph.as_default():
             self.i_XA = tf.placeholder(name='i_XA',shape=(None,2),dtype=self.dtype)
             self.i_XB = tf.placeholder(name='i_XB',shape=(None,2),dtype=self.dtype)

### 2019-11-28 paper/paper.tex
Work on the paper
old edits to the paper

@@
 
+\begin{document}
+
 \title{Solution of Differential Algebraic Equations with Local Constraints
@@
 \author[1]{Alejandro Francisco Queiruga*}
-\address[1]{\orgdiv{Energy Geosciences Division}, \orgname{Lawrence Berkeley
-    National Laboratory}, \orgaddress{\state{CA}}, \country{USA}}
+\address[1]{\orgdiv{Energy Geosciences Division}, \orgname{Lawrence Berkeley National Laboratory}, \orgaddress{\state{CA}}, \country{USA}}
 \corres{*\email{afqueiruga@lbl.gov}}
@@
 
-\begin{document}
-
-\abstract[Abstract]{
-  Multiphase simulation is hard.
+\abstract[Abstract]{ Multiphase simulation is hard.
 Our proposed method is motivated by the problems of multiphase reaction and
@@
 
-\jnlcitation{\cname{Queiruga, A.}\ctitle{Solution of DAEs on Latent Spaces},\cjournal{I. J. Num. Meth. Eng.}, \cvol{2019;00:1--2}}
-
-\keywords{Class file; \LaTeXe; \emph{Wiley NJD}}
+\jnlcitation{} %\cname{Queiruga, A.}\ctitle{Solution of DAEs on Latent Spaces},\cjournal{I. J. Num. Meth. Eng.}, \cvol{2019;00:1--2}
+
+\keywords{}
 
@@
   \centering
-  \includegraphics[width=3in]{../slides/cstr.png}
+  \includegraphics[width=3in]{../figures/cstr_vector.pdf}
   \caption{ \label{fig:cstr} Illustration of the Continuous
@@
 
-Note that the definition of enthalpy as a per-mass quanitity means the mixture needs to be weighted by the
+Note that the definition of enthalpy as a per-mass quantity means the mixture needs to be weighted by the
 density: $h_{mix}=\sum{\rho_a h_aS_a}/\sum{rho_aS_a}$.
@@
 Figure \ref{fig:phaseboundaries}. The use of intensive properties breaks down:\footnote{https://en.wikipedia.org/wiki/Phase\_transition}
-
 \begin{quote}
@@
 \end{quote}
-
-{\bf Remark:} This interpretation breaks down with the existenceof super-saturated phases in non-stable regions: e.g., it is possible to
+{\bf Remark:} This interpretation breaks down with the existence of super-saturated phases in non-stable regions: e.g., it is possible to
 heat water beyond it's boiling point in a microwave. We are concerned
 with environments where this would not happen (the very rough and
-heterogenous) pores of rocks, so we have not thought of how to
-integrate this problem.
+heterogenous pores of rocks), so we have not thought of how to
+integrate this behavior.
 
@@
 formulation for water gas, liquid, and supercritical phases, and the
-2006 formulation for the Ice Ih phase to build the dataset.\footnote{Our
-implementation is publicly available at GITHUB:EOS.} (Their
+2006 formulation for the Ice Ih phase to build the dataset.\footnote{The
+implementation is publicly available as a stand-alone library at \url{https://github.com/afqueiruga/equations\_of\_state}.} (Their
 painstaking work in determining the forms of decent fits is a great
@@
 {\bf Remark:} Many popular forms of EOS do not even have closed form
-expressions. The family of cubic equations of state for nonideal
-gases, such as the Peng-Robinsion equation \cite{peng_new_1976}, are
+expressions. The family of cubic equations of state for non-ideal
+gases, such as the Peng-Robinson equation \cite{peng_new_1976}, are
 expressed as functions for $p$ of $V$ (a proxy for $\rho$) and
-$T$. Since $\rho$ is rarely a primary varibale, these expressions must be inverted for
+$T$. Since $\rho$ is rarely a primary variable, these expressions must be inverted for
 $\rho$ given $p$ by solving the cubic equation. This can be performed
@@
 to differentiation this equation state. Approximations
-abound--the mathematical theory is not straight foward. Passing
+abound--the mathematical theory is not straight forward. Passing
 automatic differentiation objects through these computations is
@@
 \end{center}
-Solving the balance laws with this caveat requires the incorpoeration
+Solving the balance laws with this caveat requires the incorporation
 of a statemachine. Representing the material with the known
@@
 \end{figure}
-Formulate balance equations as
+The discretized balance equations become a nonlinear function of
+continuous variables $X$ and discrete indices $phase$:
 \begin{equation}
@@
 \end{equation}
-Solve the differentiable part:
-\[\frac{\partial R}{\partial X}\Delta X^{k+1/2} = -R(X^k ; phase^k)\]
-and iterate on the nondifferential part:
-\[phase^{k+1},X^{k+1} = statemachine(X^{k+1/2} ; phase^k)\]
+where the meaning of $X$ is dictated by $phase$. Solving this equation
+requires first iterating on the differentiable part by solving the
+Jacobian matrix,
+\begin{equation}
+  \frac{\partial R}{\partial X}\Delta X^{k+1/2} = -R(X^k ; phase^k)
+\end{equation}
+and the iterating the non-differentiable state machine part:
+\begin{equation} \label{eq:fsm}
+  phase^{k+1},X^{k+1} = statemachine(X^{k+1/2} ; phase^k)
+\end{equation}
 The phase switching logic has the form of a switch case illustrated in
-Algorithm \ref{alg:statemachine}.
+Algorithm \ref{alg:statemachine}. This converges quadratically in
+situations where no phases changes occur, but execution of
+Eq. \ref{eq:fsm} breakdown into linear convergence or even divergence.
+This often contains the presence of ``tricks'' such as overshoots to
+push the grid blocks over the edge, $P_{new}=(1+10^{-6} )P_{old}$, or
+artificially smoothing the phase boundaries
+\citep{moridis_simulation_2019}. 
+ Besides the computational difficulties, from a software
+engineering standpoint, the long and winding logical branches are very
+difficult to maintain and extend with new physics. The presence of the state machine and implicit equations inside of the
+material descriptions also requires numerical differentiation, which is
+a further numerical challenge. The engineering
+difficulty further makes it difficult to improve the efficiencies of the implementations.
 
@@
 problem with a Lagrange-multipler-esque centripetal force, it
-has three unknowns wtih two second-order equations and a constraint,
-
+has three unknowns with two second-order equations and a constraint,
 \begin{align}
@@
   \center
-\includegraphics[width=3in]{../figures/autoencoder_balance_detailed.png}
+\includegraphics[width=5in]{../figures/autoencoder_balance_detailed.png}
   \caption{\label{fig:balance}Illustration of the insert of the encoder to form the
@@
 \section{Representation Learning for
-Reparameterization}\label{header-n3294}}
+Re-parameterization}\label{header-n3294}}
 
@@
 parameterization of the equation of state constraint. We want to select a
-\(q_1\) and \(q_2\) to require no auxilary phase index to define the system
+\(q_1\) and \(q_2\) to require no auxiliary phase index to define the system
 and no additional logic in the code. The sharp kinks and possible
@@
 the equation of state by learning and autoencoder. We have an encoding
-phase, \(E(\rho,p,e,T; a)\) and a decoding phase, \(D(u,v; b)\) that
+phase, $E(\rho,p,e,T; a)$ and a decoding phase, $D(u,v; b)$ that
 forms an identity function with a compressed subspace:
@@
 
-\[L(x)=\left\|x-D(E(x))\right\|^2_2+\lambda\left\|\frac{\partial E}{\partial x}\right\|_F^2\]
+\begin{equation}
+L(x)=\left\|x-D(E(x))\right\|^2_2+\lambda\left\|\frac{\partial E}{\partial x}\right\|_F^2
+\end{equation}
 
@@
 manufacture more datapoints. We train by alternating phases. The first
-step is minibatched stochastic descent on the entire deep network, assembling the losses on minibatches of the data, $X_{mini}$,
+step is mini-batched stochastic gradient descent on the entire deep network, assembling the losses on minibatches of the data, $X_{mini}$,
 \begin{equation}
@@
 
-\[[\mathbf{H}] \Delta W^{dec} = -\{\mathbf{G}\}\]
+\begin{equation}
+[\mathbf{H}] \Delta W^{dec} = -\{\mathbf{G}\}\]
+\end{equation}
 
@@
 not depend on \(W^{dec}\).)
-
 \begin{equation}
@@
 \end{equation}
-
-For this step in which only the back-end regression is performmed, each set of coefficients for the curves for each variable is independent from those of the other variables: e.g., the coefficients to the sets of polynomials that parameterize $p(q)$ do not depend on those for $T(q)$.
-This means that this step can be performmed in blocks, performing the linear step for each component independently, at a great reduction of cost.
+For this step in which only the back-end regression is performed, each set of coefficients for the curves for each variable is independent from those of the other variables: e.g., the coefficients to the sets of polynomials that parameterize $p(q)$ do not depend on those for $T(q)$.
+This means that this step can be performed in blocks, performing the linear step for each component independently, at a great reduction of cost.
 For the full system, 3/4 of the entries are zero.
@@
 
-Examining the $T,p,\rho$ surface allows for human interpretability, but
-the decoder actually yields four new functions with two unknowns:
-$D(q)=\{T(q),p(q),\rho(q),h(q)\}$. These surfaces can be plotted as
-independent 2D functions: the surface of \ref{fig:classified} is
-unfolded into the new parameterization in Figure \ref{fig:foursurfaces}.
+\subsection{Single Grid Block}
+
 \begin{figure}
@@
 
+Examining the $T,p,\rho$ surface allows for human interpretability, but
+the decoder actually yields four new functions with two unknowns:
+$D(q)=\{T(q),p(q),\rho(q),h(q)\}$. These surfaces can be plotted as
+independent 2D functions: the surface of \ref{fig:classified} is
+unfolded into the new parameterization in Figure
+\ref{fig:foursurfaces}. These surfaces were designed to be smooth
+enough to solve effectively.
 With the trained model, we solve for \(q_1(t)\) and \(q_2(t)\) such that:
 \begin{align}
-\partial_t \rho(q_1,q_2) & = \mathbf{k} \left( p_\infty-p(q_1,q_2) \right)  + r \\
+\partial_t \rho(q_1,q_2) & = \mathbf{k} \left( p_\infty-p(q_1,q_2) \right)  + r \label{eq:massbalanceq}\\
 \partial_t \left(\rho h(q_1,q_2)-p\right) &
-                                            =\mathbf{k}_T(T_\infty-T(q_1,q_2))  + s + \frac{\rho h}{h}\left(\mathbf{k}_p(p_\infty - p) + r\right)
+                                            =\mathbf{k}_T(T_\infty-T(q_1,q_2))+ s + \frac{\rho h}{h}\left(\mathbf{k}_p(p_\infty - p) + r\right) \label{eq:heatbalanceq}
 \end{align}
@@
 where $m=\{\rho,\rho h -p\}$ and $r$ is the right hand side, where
-$eos(T,p,\rho,h)=0$ is the material constraint.
-
-The decoder part of the network is inserted into the balance law
+$eos(T,p,\rho,h)=0$ is the material constraint. The decoder part of the network is inserted into the balance law
 \begin{equation}
@@
 \end{equation}
-We can parameterize this equation with any time stepping scheme, .e.g. a
+We can parameterize this equation with any time stepping scheme, e.g. a
 Runge-Kutta scheme. For subsurface flow, backward Euler is popular due
-to its simplicity and L-stability, which we follow since the topic of
+to its simplicity and L-stability. We make the same choice because the topic of
 this paper is not exploring new time stepping schemes. This yields the
@@
 value \(q^0\) at \(t\),
-
-\begin{equation}
-0 = R(q^i) = m(q^{i}) - m(q^{0}) - \Delta t \, r(q^i)
+\begin{equation}
+0 = R(q^i) = m(q^{i}) - m(q^{0}) - \Delta t \, r(q^i) \label{eq:bweuler}
 \end{equation}
@@
 \begin{equation}
-\frac{\partial R}{\partial q} \Delta q^{k+1} = - R(q^k)
+\frac{\partial R}{\partial q} \Delta q^{k+1} = - R(q^k) \label{eq:jacobianq}
 \end{equation}
@@
 the vast majority of programming was implemented automatically by the
-models; implementing equations XX, XX, and XX requires only a few lines
-of human-written code.
+models; implementing equations \ref{eq:massbalanceq}, \ref{eq:heatbalanceq}, and \ref{eq:bweuler } requires only a few lines
+of human-written code. The iteration for \ref{eq:jacobianq} can be constructed entirely with automatic differentiation because the model is fully differentiable.
 
@@
 
-The current implementation LatentSim does not support inputting via
-saturations on the equilibria, a commonly used primary variable, but
-this problem could be solved with future engineering work for a
-non-experimental simulator.
-
 For LatentSim, the input system is not yet as robust as a
@@
 \end{verbatim}
+The current implementation LatentSim does not support inputting via
+saturations on the equilibria, a commonly used primary variable, but
+this problem could be solved with future engineering work for a
+non-prototype simulator.
 Then, the simulator must determine $q(t=0)$ from these two
-values, but the econder only works when all four values are on the
+values, but the encoder only works when all four values are on the
 surface. Let $s=\{T,p,\rho,\rho h\}$ be a four dimensional
@@
 For example, if the initial guess was in the liquid phase and the target point
-in the EOS space is in a solid, the iteration will fail.
-To address this issue, we perform multiple starts with an initial condition in each
-phase and equilibrium, discard points that diverge, and then pick the
+in the EOS space is in a solid, the iteration will diverge.
+To address this issue, the solver performs multiple starts with an initial condition in each
+phase and equilibrium, discard points that diverge, and then picks the
 \(q\) that is closest to the two specified components. The multiple
 start points are specified at EOS development time to form a set of
-values that represent middle points for each phase,
+values that represent middle points for each phase, \footnote{The multistart method could be specified without any prior knowledge
+with random points, but picking start points with prior knowledge was necessary to test the cause of the divergence.}
 \begin{equation}
@@
   \end{equation}
-The multistart method could be specified without any prior knowledge
-with random points, but this is a sufficiently robust method for
-now. Each of the possible guess points is used to fill in the two
+Each of the possible guess points is used to fill in the two
 unspecified values as guesses and then passed into the encoder.
@@
   \centering
-  \includegraphics[width=6in]{../slides/user_workflow.png}
+  \includegraphics[width=4in]{../figures/workflow_vector.pdf}
   \caption{\label{fig:workflow}Work flow of the simulation programming process.}
@@
 optimal in terms of compute time and satisfactorily accurate.
-The ``best'' model can then be ``shipped'' (i.e., marked for
-deployment in DVC) to be automatically selected by the final end user,
+The ``best'' model can then be ``shipped''  to be automatically selected by the final end user,
 who only specifies the Equation of State.
@@
 
-\subsection{Things we'll do differently}
-
-Adopt julia for at least the simulation
-
 \hypertarget{header-n3359}{%
@@
 
-\begin{itemize}
-\item
-  Deep learning to replace and improve hand-baked equations and
-  algorithms
-\item
-\item
-  \textbf{Only training the material representation, not the balance
-  laws}
-\item
-  Extend to more complicated materials
-\item
-  Put into a flow simulation
-\item
-  Close loop on testing with reinforcement learning
-\end{itemize}
-
-The ultimate goal for this work is to automate integration of data
-analysis into simulation.
-
-We hypothesize this end-to-end approach will allow the system to develop
-\emph{better} simulations \emph{faster}.
-
-We demonstrated a specially crafted architecture whose parameters can be
-human interpretted and perform unsupervised phase/equilibria
-classification.
-
-We illustrated a programming model that is able to incorporate machine
-learning models into existing well known computational methodologies.
-This allows us to enforce physically known quantities, such as the
-balance laws.
-
-The TensorFlow system is not quite suited for this type of workflow -\/-
-the resulting simulations are extremely slow due to repeated calls to
-session.run and a very complicated graph structure for the polynomials.
-Currently, the system is being reimplemented in Julia using Flux and
-Zygote to seamlessly reincorporate into a fully-fledged new reservoir
-simulator. The pendulum has been reimplemented in Julia at .
-
-We are also demonstrating a new programming paradigm applied to
-scientific computing. The entire system has another level of complexity
-to it. An offline training phase searches for a new representation, i.e.
-for new equations with new unknown degrees of freedom, that are good to
-solve.
-
 Each implemented equation of state (three options in this work spanning
@@
 
-We can also think about this work as offline preconditioning.
-
-In TOUGH, for one instance, all of the calculations are performed in SI units because it is simply too much labor to normalize everything!
-Adverse effects to iterative algorithms and floating point error propagation occur, especially because numerical differentiation is used.
-The learned primary variables are normalized to a unit range around zero, and the scaling matrix is explicitly in one step.
-The scaling matrix could be a way to systematically non-dimensionalize the entire problem.
+We are also demonstrating a new programming paradigm applied to
+scientific computing.
+Approaching this from a building-block problem, we can craft models
+with baked in physical knowledge.
+The entire system has another level of complexity
+to it. An offline training phase searches for a new representation, i.e.
+for new equations with new unknown degrees of freedom, that are good to
+solve.
+This allows us to enforce physically known quantities, such as the
+balance laws.
+The ultimate goal for this work is to automate integration of data
+analysis into simulation.
 
@@
 
+We foresee this end-to-end approach will allow the system to develop
+\emph{better} simulations \emph{faster}.
+We demonstrated a specially crafted architecture whose parameters can be
+human interpretted and perform unsupervised phase/equilibria
+classification.
+
+We can also think about this work as offline preconditioning.
+In TOUGH, for one instance, all of the calculations are performed in SI units because it is simply too much labor to normalize everything!
+Adverse effects to iterative algorithms and floating point error propagation occur, especially because numerical differentiation is used.
+The learned primary variables are normalized to a unit range around zero, and the scaling matrix is explicitly in one step.
+The scaling matrix could be a way to systematically non-dimensionalize the entire problem.
+
 Techniques such as L1 regularization would be needed to search for low
@@
 simplifications. (The IAPWS relations used a preliminary form of this
-type of simplification, but are not optimal in the author's opinion.)
-
+type of simplification.)
 A fully automated testing system can verify which realizations worked
@@
 
+The TensorFlow system is not quite suited for this type of workflow ---
+the resulting simulations are extremely slow due to repeated calls to
+session.run and a very complicated graph structure for the
+polynomials. To implement a fully-fledged reservoir simulator, new programming libraries are being explored, such as
+Julia with Flux and Zygote for machine learning, or the JAX library in Python.
+
 \section*{Acknowlegements}

### 2019-11-28 kernels_pop.py
Assembly moved over

@@ -30,7 +30,7 @@ o_R = pc.Output('R',(Vol_2Vec,),1)
 o_K = pc.Output('K',(Vol_2Vec,),2)
 pc.Kernel('idty_2_R',listing=[
     pc.Asgn(o_R,i_R),
-])
+])a
 pc.Kernel('idty_2_RK',listing=[
     pc.Asgn(o_R,i_R),
     pc.Asgn(o_K,i_K.reshape(4,4))
